### 2012-04-15 front-matter.tex
new files

\title{%
AA Distributed Software Development Methodology
}

\author{%
Renato~Fabbri \and Ricardo~Fabbri \and Vilson Vieira \and Alexandre Negrao \and Lucas Zambianchi
}

\maketitle
%\thispagestyle{empty}

\begin{abstract}
Sweet paper about the aa!
\end{abstract}

### 2012-04-16 front-matter.tex
fig
new abstract

@@
 Renato~Fabbri \and Ricardo~Fabbri \and Vilson Vieira \and Alexandre Negrao \and Lucas Zambianchi
+\and Marcos Mendonca
 }
@@
 \begin{abstract}
-Sweet paper about the aa!
+Algorithmic Autoregulation is a new self-regulating methodology for coordinating
+distributed teamwork,  based on collaboration and
+individual merit. Team members take on an egalitarian role, and stay
+voluntarily logged into so-called AA sessions for part of their time (e.g.,
+2h/day), during which they check for open tickets and microblog periodical
+"tweets" or status messages they wish to share about their activity with the
+team. These status logs are publicly aggregated and are peer-validated,
+as in code review, and a public videolog is recorded summarizing daily experiences.
+This methodology is well-suited for increasing the efficiency of distributed
+teams through asynchronous on-demand communication, reducing the need for
+central management, unproductive meetings or time-consuming reports.  The AA
+methodology also legitimizes the activities of a distributed software team.  It
+thus enables entities to have a solid means to fund these activities, allowing
+for new and concrete business models to emerge for very distributed software
+development.  
 \end{abstract}

### 2012-04-16 front-matter.tex
talked about self-replication

@@
 \begin{abstract}
-Algorithmic Autoregulation is a new self-regulating methodology for coordinating
-distributed teamwork,  based on collaboration and
-individual merit. Team members take on an egalitarian role, and stay
+We present a new self-regulating methodology for coordinating
+distributed teamwork called Algorithmic Autoregulation (AA),  based on recent social
+networking concepts and individual merit. Team members take on an egalitarian role, and stay
 voluntarily logged into so-called AA sessions for part of their time (e.g.,
@@
 for new and concrete business models to emerge for very distributed software
-development.  
+development. AA as a methodology is also at the core of having self-replicating
+hacker initiatives. These claims are discussed in a real case-study of running a distributed
+free software hacker team called Lab Macambira.
 \end{abstract}

### 2012-04-16 paper.tex
paper wasnt in repo

%\documentclass[letterpaper,peerreview,12pt,compsoc,draftcls]{IEEEtran}
%\documentclass[10pt,twocolumn,letterpaper]{article}
\documentclass[letterpaper]{article}

%
% Final Checklist TODO XXX 
% - Spellcheck
% - check xxx's todo's etc
%
%\input{preamble}
\usepackage{url}
\usepackage{graphicx}
\usepackage[tight,footnotesize]{subfigure}
\usepackage{color}
\usepackage{verbatim}
\usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% You have two versions of the macro
% \draftnote{My note}. The first version puts notes (e.g. My note in the example)
% into the margin of your document. The second formats the note as nothing. You
% 'comment out' the version of the macro you don't want (using a % at the
% beginning of the line).
%\newcommand{\draftnote}[1]{\marginpar{\tiny\raggedright\textsf{\hspace{0pt}#1}}}
\newcommand{\draftnote}[1]{\marginpar{\tiny\raggedright\textsf{\hspace{0pt}#1}}}
%\newcommand{\draftnote}[1]{}

% This one is just for the comments for in-line text.
\newcommand{\indraftnote}[1]{\textcolor{blue}{\texttt{\footnotesize[#1]}}}
\newcommand{\todo}[1]{\indraftnote{todo: #1}}
%\newcommand{\indraftnote}[1]{}

\begin{document}
\input{front-matter}

\section{Introduction}

One of the defining features of modern times is the widening geographical
distribution of software teams~\cite{}. One example is the free software
movement. \todo{citar alguns exemplos}.

However, we have noticed how difficult it is to coordinate and fund free
software on a larger scale than currently available, when teams are very
heterogeneous containing not only volutaries and very experienced developers,
but also contractors from different backgrounds and cultures.

Another problem faced by modern software companies and other collectives are frequent
ineffective meetings, which are seldom focused to the interest to any
attendant. The result is that it has become the norm to participate in too many
meetings with the laptop open, which is very unproductive at the very least.
Coders like to code, to be productive, to have their hands-on their project to
do what they're best at. They hate to have to stop for too many meetings,
and they hate to have to write lengthy reports to justify their funding.
\todo{ler mais. cacm, etc}

AA is a methodology and an associated software system for coordinating
distributed teamwork.  Team members take on an egalitarian role, and stay
voluntarily logged in the system for part of their time (e.g., 2h/day), during
which they microblog periodical "tweets" about their activity. These microblog
logs are publicly aggregated and are validated by their own peers. Through AA,
we have a methodology and an associated system to validate and enable the
activities of a distributed software team. It implicitly legitimizes financing
the expansions of the team's activity. The AA methodology is specially useful
for coordinating distributed and decentralized teamwork, providing effective
means to asyncrhonously update different team members without the need for
syncrhonous unproductive meetings.

Figure~\ref{fig:mm} summarizes our methodology.

\begin{figure}
\begin{center}
   \includegraphics[width=0.8\linewidth,keepaspectratio=true]{figs/aa-mm.png}
\end{center}
   \caption{
   Mindmap of our methodology
   }
\label{fig:mm}
\end{figure}

\todo{alerts work for greater consciousness of time}

\section{Related Work}
\todo{survey other methodologies such as agile etc}

\section{The AA Methodology}

\section{The AA Methodology}

\section{Conclusion}

%\section*{Acknowledgments}
%The authors would like to thank NSF and CNPq.

%{\small
%\input{paper-draft.bbl}
%%%%\bibliographystyle{splncs}
% this is a good style for drafting
%\bibliographystyle{abstract}
% this is a good style for finals
\bibliographystyle{acm}
\bibliography{personal}
%}

\end{document}

### 2012-04-17 paper.tex
AA Methodology - Session
Added an overlook of how an aa session works and why, probably good expand it,
also added the link for agile manifesto for referencing later

@@
 \todo{survey other methodologies such as agile etc}
+\todo{probably a good source http://agilemanifesto.org/}
 
 \section{The AA Methodology}
+
+\section{AA Session}
+
+From the developer perspective, the AA methodology is based on creating pretty
+small high perspective reports of what they're doing in a specific timeframe,
+that can be something between 5 to 15 minutes, depending in what is more
+confortable for the developer, an AA Session would be a period of at least 2hrs
+doing these reports. The developer can set reminders to show up when its time
+to make a new report.
+
+The objective of the flexible timeframe and reminders is to don't generate much
+overhead for the developer to be in an AA session, so he can make the reports
+and also concentrate in his code.
+
+Each report can be sent directly to the an online server, or cached locally in
+a log for sending later, this is necessary to make possible for developers that are
+offline to use AA without problems.
+
+The developer can also add a screencast in the end of session to make a summary
+of what has been done in the session, explaining with his words and showing his
+results, possibly making some points more easily understandable then with only
+the reports from the AA session.
 

### 2012-04-18 paper.tex

@@
 
+\section{AA Site Report}
+
+\begin{figure}
+\begin{center}
+   \includegraphics[width=0.8\linewidth,keepaspectratio=true]{figs/aa-0.1.png}
+\end{center}
+   \caption{
+   AA Version 0.1
+   }
+\label{fig:mm}
+\end{figure}
+
+All AA reports made by the developers are sent to a web site and become public,
+being possible then for a manager or another developer to follow very closely
+the work of a developer, nearly real-time, reading each of his small reports of
+what he is doing.
+
+Another possibility is to check older sessions to see when sometimes was done
+and the comments of the developer about it, since each AA post happens in a
+very small timeframe of work, the information about what was done become very
+easy to understand, instead of only of report in the end of a session.
+
+The site is where the developer can add to his session a link for his
+screencast about that session, a video summary of what was done to complement
+the reports, useful specially on cases where the small reports were done in a
+hurry, because the developer didn't want to lose his focus on something
+important at that moment.
+
+\secion{Validation}
+
+Each developer AA Session must be validated by another developer, it means that
+all reports are read by someone that will consider then valid as reports for
+that session of work, so the session doesn't gets lost on the site, the
+developer in charge of validating a session is decided randomly.
+
 \section{The AA Methodology}

### 2012-06-26 front-matter.tex
add recent works, results, discussions and conclusions. whole paper reviewed.

--- 2012-04-15 front-matter.tex
@@
-\title{%
+\title{
 AA Distributed Software Development Methodology
@@
 Renato~Fabbri \and Ricardo~Fabbri \and Vilson Vieira \and Alexandre Negrao \and Lucas Zambianchi
+\and Marcos Mendonca \and Danilo Shiga
 }
@@
 \begin{abstract}
-Sweet paper about the aa!
+We present a new self-regulating methodology for coordinating
+distributed team work called Algorithmic Auto-regulation (AA), based
+on recent social networking concepts and individual merit. Team
+members take on an egalitarian role, and stay voluntarily logged into
+so-called AA sessions for part of their time (e.g.\ 2 hours per day),
+during which they create periodical logs --- short text sentences ---
+they wish to share about their activity with the team. These logs are
+publicly aggregated in a Website and are peer-validated, as in code
+review. A video is generally recorded by the members to make their
+work logs more understandable. This methodology is well-suited for
+increasing the efficiency of distributed teams working on what is
+called Global Software Development (GSD) through asynchronous
+on-demand communication, reducing the need for central management,
+unproductive meetings or time-consuming reports. The AA methodology
+also legitimizes the activities of a distributed software team.  It
+thus enables entities to have a solid means to fund these activities,
+allowing for new and concrete business models to emerge for very
+distributed software development. AA as a methodology is also at the
+core of having self-replicating hacker initiatives. These claims are
+discussed in a real case-study of running a distributed free software
+hacker team called Lab Macambira.
 \end{abstract}

### 2012-06-26 paper.tex
subsection-ized some sections
renamed figure cuz of apparent texlive bugs with image extensions
add recent works, results, discussions and conclusions. whole paper reviewed.

--- 2012-04-16 paper.tex
@@
 \usepackage{verbatim}
+\usepackage{multirow}
 \usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}
@@
 
-One of the defining features of modern times is the widening geographical
-distribution of software teams~\cite{}. One example is the free software
-movement. \todo{citar alguns exemplos}.
+One of the defining features of modern times is the widening
+geographical distribution of software teams~\cite{last2003} creating
+what is called Global Software Development (henceforth
+GSD)~\cite{german2003}. An example is the free software
+movement. Projects and institutions like Mozilla Foundation has many
+employees and thousands of voluntary developers distributed around
+many countries. The same is true for GNOME~\cite{german2003}, OpenBSD,
+MySQL or Apache Software Foundation, to enumerate just the most active
+projects\footnote{Ohloh, the open source network, have a more complete
+  and constantly updated list of the most active projects on-line at
+  \url{http://www.ohloh.net/}}. Along the free and open software
+projects, GSD is growing popularity in every niche of software
+industry as a whole, even on those distributing their software with
+proprietary licenses. This phenomenon is attributed to various factors
+like larger labor pool, natural globalization of software companies
+and foundations or even the premise of cheaper cost of
+production~\cite{komi2005}.
 
-However, we have noticed how difficult it is to coordinate and fund free
-software on a larger scale than currently available, when teams are very
-heterogeneous containing not only volutaries and very experienced developers,
-but also contractors from different backgrounds and cultures.
+However, dispite the advantages of GCD, we have noticed how
+difficult it is to coordinate and fund free software on a larger scale
+than currently available, when teams are very heterogeneous containing
+not only volunteers and very experienced developers, but also
+contractors from different backgrounds and cultures. Our observations
+found base on the factors suggested by Carmel~\cite{carmel1999} as
+main difficulties for GCD: distance, time and cultural differences. In
+the case of free or open software projects, all these factors are
+involved.
 
-Another problem faced by modern software companies and other collectives are frequent
-ineffective meetings, which are seldom focused to the interest to any
-attendant. The result is that it has become the norm to participate in too many
-meetings with the laptop open, which is very unproductive at the very least.
-Coders like to code, to be productive, to have their hands-on their project to
-do what they're best at. They hate to have to stop for too many meetings,
-and they hate to have to write lengthy reports to justify their funding.
-\todo{ler mais. cacm, etc}
+Another problem faced by modern software companies and other
+collectives are frequent ineffective meetings, which are seldom
+focused to the interest to any attendant. The result is that it has
+become the norm to participate in too many meetings with the ``laptop
+open'', which is unproductive at the very least. Software developers
+like to code, to be productive, to have their hands on their project,
+to do what they are best at. They dislike to have to stop for meetings
+or to write lengthy reports to justify their funding.
+%\todo{ler mais. cacm, etc}
 
-AA is a methodology and an associated software system for coordinating
-distributed teamwork.  Team members take on an egalitarian role, and stay
-voluntarily logged in the system for part of their time (e.g., 2h/day), during
-which they microblog periodical "tweets" about their activity. These microblog
-logs are publicly aggregated and are validated by their own peers. Through AA,
-we have a methodology and an associated system to validate and enable the
-activities of a distributed software team. It implicitly legitimizes financing
-the expansions of the team's activity. The AA methodology is specially useful
-for coordinating distributed and decentralized teamwork, providing effective
-means to asyncrhonously update different team members without the need for
-syncrhonous unproductive meetings.
+Here we propose AA, a methodology and an associated software system
+for coordinating distributed team work dealing with the disadvantages
+of GSD. Team members take on an egalitarian role, and stay
+voluntarily \textit{logged} in the system for part of their time
+(e.g.\ 2 hours per day), during which they log a periodical short text
+sentence --- similar to a ``tweet'' in this Twitter era --- as the
+status of their activity using an easy to use command. These
+``microblog sentences'' are publicly aggregated and validated by other
+team members. Through AA, we have a methodology and an associated
+system to validate and enable the activities of a distributed software
+team. It implicitly legitimizes financial support for the expansion of
+the activity of the developer team. The AA methodology is specially
+useful for coordinating distributed and decentralized team work,
+providing effective means to asynchronously update different team
+members without the need for synchronous unproductive meetings.
 
-Figure~\ref{fig:mm} summarizes our methodology.
+%\todo{alerts work for greater consciousness of time}
+%\todo{Integrate notes from Ricardo+Renato april 15/16 meeting}
+%\todo{integrate IRC chat notes}
+
+A brief overview of current papers about GSD methodologies related with
+AA is presented in Section~\ref{related-work}, while in
+Section~\ref{aa-methodology} we describe the most relevant
+characteristics of the AA methodology followed by the description of
+our experience using AA in a team of 9 developers since July of
+2011. Finally, in Section~\ref{conclusions}, we draw some final
+conclusions and indicate future possibilities to the
+practical use of AA on other teams of software developers or
+individuals working on non-software activities.
+
+\section{Related Work}
+\label{related-work}
+
+%% aqui tomei como ponto de vista as metodologias associadas a GSD
+%% (Global Software Development) que considero foco do AA, e sua
+%% principal vantagem
+
+%\todo{survey other methodologies such as agile etc}
+%\todo{probably a good source http://agilemanifesto.org/}
+
+There has been a large amount of research done in the area of
+methodologies to deal with distributed teams of developers. We are
+focusing in GSD here, however some principles involved on those
+methodologies could be used on smaller teams of developers working in
+the same place, time and with minor cultural differences. Moreover we
+generally think on ``distributed development'' being global which is
+not totally true. We even applied AA to a team that work at the same
+city but on different times (more details on
+Section~\ref{results}). Even smaller groups of developers working on
+the same building could use GSD methodologies. A thorough survey of
+these methodologies is beyond the scope of this paper. Here we present
+a brief overview.
+
+Various methodologies for GSD were built around the factors that
+affect distributed team works, proposed by Carmel~\cite{carmel1999}
+and comprising three distances: geographical, cultural and
+temporal. Geographical distance prejudice \emph{coordination}, being
+the act of integrating all the tasks distributed between
+units~\cite{carmel2001}; \emph{control}, or the process to maintain
+specific goals, policies or quality levels; and
+\emph{communication}. All these factors are correlated, for example, a
+team need to have clearly communication to work on tasks of a specific
+problem. Cultural distance encompass differences on organizational and
+natural culture. Spoken language, unit and ethnic values are common
+forms of this distance. Some companies prefer to situate development
+units in foreign locations with minimal cultural distance (e.g.\ an
+American firm prefer Ireland, because of spoken language
+similarity~\cite{carmel2001}). And finally the temporal distance that
+prejudices synchronous communications like telephone or
+videoconferences. Units of developers working on different time-zones
+are concerned with managing of their agenda guided by this temporal
+distance.
+
+Targeting geographical distance, Carmel~\cite{carmel2001} suggests a
+tactic to reduce intensive collaboration. His approach divides the
+whole software life-cycle on levels of complexity. Each level has a
+degree of collaboration. For example, some developers working on a
+project with high collaboration level should use the follow-the-sun
+approach: when concluding the work day, they pass their work to the
+team working on another time-zone. Other tactics are suggested by the
+same author to deal with the three distances, like separating foreign
+units of developers in time-zone bands.
+
+Battin et al.~\cite{battin2001} propose and argues about their
+experiments using specific methodologies created for the distributed
+development centers of the Motorola Company (which has over 25
+software development centers worldwide). These methodologies included
+constant communication with critical units, incremental integration
+and scheduled based on time-zones distributed to developers on 6
+countries from 3 continents.
+
+While considering free software projects instead of companies, the
+same factors are present and some methodologies
+arises. German~\cite{german2003} gives a concise review of
+methodologies used by the GNOME project, one of the most active free
+software projects and used by companies like Sun Microsystems. It is
+interesting to note a difference on viewpoint present on his paper:
+German focuses the methodology description on code. He start explaining
+that GNOME is separated into modules (76 on version 2.4, to be
+precise) and each module has one maintainer who divide his modules
+into separated parts in which other developers can work on independent
+tasks, along other responsibilities. As like others free software
+projects, all the development was done using a \emph{bugtracker} to
+bugs and issues management, mail lists and Internet Relay Chat (IRC)
+to discussion and communication and a software configuration
+management like SVN or Git. Periodical (commonly yearly) conferences
+like GUADEC is common on free and open source projects to face-to-face
+meeting and is based each time on a different place.
+
+\section{The AA Methodology}
+\label{aa-methodology}
+
+As noted, some strategies for GSD is based on complex methodologies
+and many of those were built for a specific company or software
+center. Here we propose an alternative methodology based on a simple
+idea: small working sessions logged by a computational
+tool. Figure~\ref{fig:mm} summarizes our methodology.
 
@@
 
-\todo{alerts work for greater consciousness of time}
+\subsection{AA Session}
 
-\section{Related Work}
-\todo{survey other methodologies such as agile etc}
+From the developer perspective, the AA methodology is based on
+creating pretty small high perspective reports of what they are doing
+in a specific time frame, that can be something between 5 to 15
+minutes, depending in what is more comfortable for the developer. An
+AA Session would be a period of at least 2 hours doing these
+reports. The developer can set reminders to show up when its time to
+make a new report. The objective of the flexible time frame and
+reminders is to minimize developer overhead during his AA session. In
+this way he can make the reports while staying concentrated in his
+code. Each report can be sent directly to an on-line server, or stored
+locally in a temporarily database for sending later. This make
+possible the use of AA while offline.
 
-\section{The AA Methodology}
+Developers can also record a video screencast in the end of the
+session summarizing what has been done, explaining with his words and
+showing his most important results. This, combined with the textual
+log of his session, makes the whole report more understandable for
+himself or other people searching for information about his
+production.
 
-\section{The AA Methodology}
+\subsection{AA Website Report}
 
-\section{Conclusion}
+\begin{figure}
+\begin{center}
+   \includegraphics[width=0.8\linewidth,keepaspectratio=true]{figs/aa-0_1.png}
+\end{center}
+   \caption{AA Version 0.1}
+\label{fig:aaserver}
+\end{figure}
 
-%\section*{Acknowledgments}
+All AA reports made by the developers are sent to a Web server and
+become public in a Website, being possible then for a manager or
+another developer to follow very closely the work of a developer,
+nearly real-time, reading each of his small reports of what he is
+doing right now.
+
+Another possibility is to check older sessions to see when sometimes
+was done and the comments of the developer about it, since each AA
+post happens in a very small time frame of work, the information about
+what was done become very easy to understand, instead of a long
+report in the end of a session.
+
+The site is where the developer can add to his session a link for his
+screencast about that session, a video summary of what was done to
+complement the reports, useful specially on cases where the small
+reports were done in a hurry, because the developer did not want to
+lose his focus on something important at that moment.
+
+\subsection{Validation}
+
+Each AA session must be validated by another developer, it means that
+all reports are read by someone that will consider then valid or not
+and will even write commentaries about the specific session. The
+developer in charge of validating a session is decided randomly by the
+AA Web server, which send an email to the chosen developer with an
+URL to a validation interface.
+
+% TODO: adicionar figura aa cliente (ou tabela com os principais comandos)
+% TODO: adicionar figura com arquitetura do AA cliente + AA web + validação
+% TODO: melhorar screenshot do AA web
+
+\section{Results and Discussion}
+\label{results}
+
+The easy and effective management of teams working on GSD is the main
+purpose of the AA methodology. We applied this methodology to a group
+of 9 developers in July of 2011 on what we called Lab
+Macambira~\footnote{LabMacambira.sf.net:
+  \url{http://labmacambira.sf.net}}. The main objective of the team
+was to work on different free software projects, contributing directly
+to their development, sending bug corrections or proposing new
+features on their source code.
+
+All the team members had different levels of knowledge on software
+development as part of large and distributed free software projects
+like Scilab or Mozilla. In this way, one month of training was
+conducted by two experienced developers, teaching the basics of use of
+development support tools like bugtrackers, programming languages and
+version control systems. After this period, a challenge was proposed
+for the new developers: send a bug correction or a new feature to a
+large free software project and you will work with us, you will be a
+``Macambira'' developer. Table~\ref{tabela:contribuicoes} summarizes
+the contributions of each ``Macambira'' to free software projects.
+
+%% adicionar tabela de contribuições
+
+\begin{table}
+  \caption{Free and open software projects that received contributions
+    from ``Macambiras''. On the first column we can see a list of
+    applications of those projects. At right, the pseudonym of the
+    ``Macambiras'' who sent \textit{commits} to the application. At
+    ``Lab Macambira'', and at free software community in general, is a
+    common practice to use pseudonym as identification.}
+    \begin{tabular}{|l|l|}
+        \hline
+        Application           & ``Commiters''                       \\
+        \hline \hline
+        Mozilla Firefox       & daneoshiga, bzum                    \\
+        Evince                & hick209, bzum, marcicano, mquasar   \\
+        BePDF / Xpdf          & marcicano                           \\
+        Ekiga                 & flecha                              \\
+        Empathy               & fefo                                \\
+        Lib Folks (Telepathy) & kamiarc                             \\
+        Scilab                & v1z, humannoise                     \\
+        VxL                   & v1z                                 \\
+        ImageMagick           & v1z                                 \\
+        OpenOffice            & hick209                             \\
+        Puredata              & v1z, automata, greenkobold, gilson, bzum \\
+        Puredata OpenCV       & v1z                                 \\
+        Puredata GEM          & v1z, fefo, hick209                  \\
+        Puredata PDP          & v1z, fefo, hick209                  \\
+        ChucK                 & rfabbri, automata                   \\
+        ChucK MiniAudicle     & rfabbri, automata                   \\
+        WebRTC                & automata                            \\
+        OSC-Web               & automata                            \\
+        Web-PD-GUI            & automata                            \\
+        Live-Processing       & automata                            \\
+        ChucK-Wiimote         & automata                            \\
+        Audiolet              & automata                            \\
+        Extempore             & automata                            \\
+        \hline
+    \end{tabular}
+    \label{tabela:contribuicoes}
+\end{table}
+
+In one month, each developer contributed to many very large free software
+projects. Many of the developers started the training with no
+knowledge of what is free software and ended that period becoming a
+free software developer.
+
+During that month, the same team developed the first version of the AA
+system and used AA to manage their activities. Even while developing
+the system. All the source code of AA --- the client that sends the
+logs and the AA Web server --- is public available~\footnote{AA source
+  code:
+  \url{http://labmacambira.git.sourceforge.net/git/gitweb.cgi?p=labmacambira/aa}}
+and all the AA sessions log of the whole team of ``Lab Macambira'' is
+also on-line~\footnote{Logs of AA sessions:
+  \url{http://labmacambira.git.sourceforge.net/git/gitweb.cgi?p=labmacambira/paainel}}.
+
+After the training period, during more 6 months, the ``Macambiras''
+worked on a large range of free software projects, distributed on work
+groups --- each work group focusing on a specific theme like video,
+audio and web --- and financed by contracts and support of the
+``Pont\~{a}o N\'{o}s Digitais''. In Table~\ref{tabela:criados} we can
+see a list of the free software created by ``Lab Macambira'' since
+July of 2011.
+
+\begin{table}
+    \caption{Software projects created by ``Lab Macambira'' since July
+      of 2011 with a short description and the technologies --- like
+      programming languages or frameworks --- involved. It is
+      interesting to note the heterogeneity of projects and its areas
+      of application.}
+    \begin{tabular}{|l|p{5cm}|l|}
+        \hline
+        Application & Description & Technologies involved \\ 
+        \hline \hline
+        AA            & Algorithmic Auto-regulation      & Python, PHP \\
+        \hline
+        \'{A}gora Communs & System for on-line deliberations & PHP \\
+        \hline
+        SIP           & Scilab Image Processing toolbox & C, Scilab \\
+        \hline
+        animal        & An Imaging Library              & C \\
+        \hline
+        TeDi          & Test Framework for Distance Transform
+        Algorithms & C, Shellscript, Scilab \\
+        \hline
+        Macambot      & Multi-use IRC Bot               & Python \\
+        \hline
+        ``Confer\^{e}ncia Permanente'' & Platform for the permanent
+        conference of the rights of minors & PHP, JavaScript \\
+        \hline
+        CPC           & Center for accounting of the Brazilian culture
+        representation groups & Python, Django \\
+        \hline
+        Timeline      & Interactive time lines on the Web & JavaScript
+        \\
+        \hline
+        Imagemap      & Interactive marking for on-line photos &
+        JavaScript \\
+        \hline
+        ABT           & Program for real-time execution and musical
+        rhythmic analysis & Python \\
+        \hline
+        EKP           & Emotional Kernel Panic & Python, ChucK \\
+        \hline
+        SOS           & Aggregation and diffusion of popular and native
+        knowledge about health & Python, Django \\
+        \hline
+        Creative Economy & Platform for creative, collaborative and
+        solidarity economy of the culture hubs and cultural entities &
+        Python, Django \\
+        \hline
+        OpenID Integration & Adaptations to existing software for
+        unified login through OpenID & PHP \\
+        \hline
+        pAAinel & Dashboard for the real-time visualization of Lab
+        Macambira activity & Python, Django \\
+        \hline
+        Georef & Collection of scripts to be used as reference, which
+        aims to be a GIS platform to map public data of use to
+        citizens & Python, Django \\
+        \hline
+        AirHackTable & Software for an instrument which generates
+        sound from flying origami tracked by webcams & Puredata,
+        C/C++, Scilab \\
+        \hline
+        \end{tabular}
+    \label{tabela:criados}
+\end{table}
+
+As of this writing the ``Lab Macambira'' have many software
+developers, and some of the trained developers continue to work
+voluntarily in the project.
+
+\section{Conclusions}
+\label{conclusions}
+
+%% brief introduction
+In a scenario where GCD is growing as a popular form of software
+development, not just on free software projects but in the whole
+software industry, we need methodologies to deal with its
+disadvantages and at the same time to amplify its advantages.
+
+This paper has presented a methodology to GCD, being the development
+conducted on large or small groups of software developers, working on
+different countries or even at the same room. The AA methodology
+implements a simple system where each developer take notes of his work
+generating a periodical log of small text sentences. The sum of those
+sentences, along an entire session of work, results in a complete
+report. The report is made public available through a Website and be
+validated by other developers sorted aleatory by the AA Web server.
+
+Instead of a merely work-management tool, AA act as a methodology to
+improve the time sense of individuals, dividing their work on small
+sessions, and also reducing the need of extensive
+reports or unnecessary meetings. By asking users to write a minimal
+text sentence as a continuously log, AA does not disturb developers
+concentrated on programming: developers just have to type some
+characters, hit \textit{enter} and go back to coding.
+
+AA application is not restricted to software development. As of this
+writing there is a comic book studio~\footnote{Pula pirata:
+  \url{http://pulapirata.com}} starting to use AA to manage their
+activities. People with non-software background, like social
+scientists, musicians and activists has also using AA and contributing
+for its improvement.
+
+For developer teams, we have experienced the use of AA to
+auto-regulate the work of ``Lab Macambira'', a group of free software
+developers from Brazil. Since July of 2011 the group have contributed
+and created new free and open source software for a vast number of
+applications.
+
+There are many aspects of the work which remain unfinished. New ways
+to report logs --- the ``Twitter like'' messages --- from different
+interfaces like IRC, Internet Messaging services and email can make
+the use of AA easy and widespread, turning AA an ubiquitous system,
+presented on everyday communication channels. Even the work logs
+generated since July of 2011 could be vastly statistically analyzed
+aiming to recognize patterns in the behavior of individuals and their
+productions.
+
+We would like to conclude setting an important role of AA: being a
+free software system and an open methodology, AA could be used to
+auto-manage groups of individuals working on software or other kinds
+of activities. In this way, we are interested to spread AA for those
+groups, to have even more developers contributing in a collaborative
+way.
+
+\section*{Acknowledgments}
 %The authors would like to thank NSF and CNPq.
+
+We would like to also thanks AA: this whole research and even this
+paper was written using AA. The complete log is on-line at
+\url{http://hera.ethymos.com.br:1080/paainel/casca/}.
 
@@
 \bibliographystyle{acm}
-\bibliography{personal}
+\bibliography{aa.bib}
 %}

### 2012-08-08 paper.tex
added new reference to wired mag

@@
 individuals working on non-software activities.
+
+\indraftnote{
+A very good article on the value of asynchronous communication for personal
+and group productivity, related to the key necessity of having moments of
+introversion to avoid daily pressures of forced socialization. The way we work
+on the digital age enables people to be very productive, the article also
+mentions linux as a hallmark example~\cite{Thompson:Wired:2012}
+}
 
@@
 
+\nocite{last2003}
+\nocite{german2003}
+\nocite{carmel1999}
+\nocite{carmel2001}
+\nocite{komi2005}
+\nocite{battin2001}
+
 %{\small

### 2013-02-18 paper.tex
todo to discuss CIA.vc bots
figsize aspectratio fixed
improved 1st paragraph

@@
 GSD)~\cite{german2003}. An example is the free software
-movement. Projects and institutions like Mozilla Foundation has many
-employees and thousands of voluntary developers distributed around
+movement. Projects and institutions like Mozilla Foundation has several
+employees and thousands of voluntary developers distributed across
 many countries. The same is true for GNOME~\cite{german2003}, OpenBSD,
-MySQL or Apache Software Foundation, to enumerate just the most active
-projects\footnote{Ohloh, the open source network, have a more complete
+MySQL or Apache Software Foundation, to cite just a few of the most active
+projects.\footnote{Ohloh, the open source network, have a more complete
   and constantly updated list of the most active projects on-line at
-  \url{http://www.ohloh.net/}}. Along the free and open software
-projects, GSD is growing popularity in every niche of software
+  \url{www.ohloh.net}} Along the free and open software
+projects, GSD is growing popularity in every niche of the software
 industry as a whole, even on those distributing their software with
-proprietary licenses. This phenomenon is attributed to various factors
-like larger labor pool, natural globalization of software companies
+proprietary licenses. This phenomenon is attributed to a variety of factors
+such as a larger labor pool, natural globalization of software companies
 and foundations or even the premise of cheaper cost of
@@
 }
+
+\indraftnote{TODO: cite CIA.vc bot stuff}
 
@@
 \begin{center}
-   \includegraphics[width=0.8\linewidth,keepaspectratio=true]{figs/aa-0_1.png}
+   \includegraphics[width=0.95\linewidth]{figs/aa-0_1.png}
 \end{center}
@@
 
-We would like to also thanks AA: this whole research and even this
-paper was written using AA. The complete log is on-line at
-\url{http://hera.ethymos.com.br:1080/paainel/casca/}.
+We would like to also thank AA: the present research and even this
+manuscript was written using AA. The complete log is on-line at
+\url{http://www.pulapirata.com/skills/aa}.
 

### 2013-02-18 paper.tex
3rd and 4rth par improved

@@
 
-However, dispite the advantages of GCD, we have noticed how
+Despite the advantages of GCD, we have noticed how
 difficult it is to coordinate and fund free software on a larger scale
@@
 not only volunteers and very experienced developers, but also
-contractors from different backgrounds and cultures. Our observations
-found base on the factors suggested by Carmel~\cite{carmel1999} as
+contractors and freelancers from different backgrounds and cultures. Our observations
+are founded on the factors suggested by Carmel~\cite{carmel1999} as
 main difficulties for GCD: distance, time and cultural differences. In
@@
 become the norm to participate in too many meetings with the ``laptop
-open'', which is unproductive at the very least. Software developers
+open'', which can be un-productive. Software developers
 like to code, to be productive, to have their hands on their project,
-to do what they are best at. They dislike to have to stop for meetings
-or to write lengthy reports to justify their funding.
+to do what they are best at. They dislike to forcibly stop for meetings
+or to do other bureaucratic activities such as writing lengthy reports to justify their funding.
 %\todo{ler mais. cacm, etc}
 
-Here we propose AA, a methodology and an associated software system
-for coordinating distributed team work dealing with the disadvantages
+In this paper we propose the AA methodology and an associated software system
+for coordinating distributed team work, tackling the disadvantages
 of GSD. Team members take on an egalitarian role, and stay

### 2013-02-18 paper.tex
5th par improvd

@@
 (e.g.\ 2 hours per day), during which they log a periodical short text
-sentence --- similar to a ``tweet'' in this Twitter era --- as the
-status of their activity using an easy to use command. These
-``microblog sentences'' are publicly aggregated and validated by other
-team members. Through AA, we have a methodology and an associated
-system to validate and enable the activities of a distributed software
-team. It implicitly legitimizes financial support for the expansion of
-the activity of the developer team. The AA methodology is specially
-useful for coordinating distributed and decentralized team work,
-providing effective means to asynchronously update different team
-members without the need for synchronous unproductive meetings.
+sentence --- similar to a ``tweet'' from Twitter --- as the
+status of their activity. Logging is carried out using an easy a series of
+available UI's: unix shell
+commands, native GUI or web page, conventional social network posts, or chat messages to a log bot
+listening to IRC, GTalk, G+, and others.  These ``microblog sentences'' are publicly aggregated and validated by other
+team members. Through AA, we have a methodology and an associated system to help
+implement and validate the activities of a distributed software team. It implicitly
+legitimizes financial support for the expansion of the activity of the developer
+team. The AA methodology is specially useful for coordinating distributed and
+decentralized team work, providing effective means to asynchronously update
+different team members without the need for synchronous unproductive meetings.
 

### 2013-02-18 paper.tex
mais fixes

@@
 to do what they are best at. They dislike to forcibly stop for meetings
-or to do other bureaucratic activities such as writing lengthy reports to justify their funding.
+or to do other bureaucratic activities such as writing lengthy reports to
+justify their funding.~\cite{Thompson:Wired:2012}
 %\todo{ler mais. cacm, etc}
@@
 
-A brief overview of current papers about GSD methodologies related with
+A brief overview of current work on GSD methodologies related to
 AA is presented in Section~\ref{related-work}, while in
@@
 characteristics of the AA methodology followed by the description of
-our experience using AA in a team of 9 developers since July of
-2011. Finally, in Section~\ref{conclusions}, we draw some final
-conclusions and indicate future possibilities to the
-practical use of AA on other teams of software developers or
-individuals working on non-software activities.
+our experience using AA in a team of 9 paid developers 
+% nao era mais de 9?
+since July 2011. In Section~\ref{conclusions} we draw some final
+conclusions and indicate future possibilities for the
+practical use of AA in other types of software developers teams or
+organizations working on non-software distributed activities.
 
@@
 and comprising three distances: geographical, cultural and
-temporal. Geographical distance prejudice \emph{coordination}, being
-the act of integrating all the tasks distributed between
-units~\cite{carmel2001}; \emph{control}, or the process to maintain
-specific goals, policies or quality levels; and
-\emph{communication}. All these factors are correlated, for example, a
-team need to have clearly communication to work on tasks of a specific
-problem. Cultural distance encompass differences on organizational and
+temporal.
+
+First, geographical distance handicaps ($i$) \emph{coordination}, the act of
+integrating all the tasks distributed between units~\cite{carmel2001}; ($ii$)
+\emph{control}, or the process to maintain specific goals, policies or quality
+levels; and ($iii$) \emph{communication}. All those factors are correlated, \eg,
+a team needs to have clear communication to work on tasks of a specific problem.
+
+Second, cultural distance encompasses differences on organizational and
 natural culture. Spoken language, unit and ethnic values are common
-forms of this distance. Some companies prefer to situate development
-units in foreign locations with minimal cultural distance (e.g.\ an
-American firm prefer Ireland, because of spoken language
-similarity~\cite{carmel2001}). And finally the temporal distance that
-prejudices synchronous communications like telephone or
+forms of such distance. Some companies prefer to allocate development
+units in foreign locations with minimal cultural distance (\eg, an
+American company may prefer Ireland due to spoken language
+similarity~\cite{carmel2001}). Third, the temporal distance that
+hampers synchronous communications like telephone or
 videoconferences. Units of developers working on different time-zones

### 2013-02-18 paper.tex
polished sect 2

@@
 Targeting geographical distance, Carmel~\cite{carmel2001} suggests a
-tactic to reduce intensive collaboration. His approach divides the
-whole software life-cycle on levels of complexity. Each level has a
+strategy to reduce intensive collaboration. His approach divides the
+whole software life-cycle into levels of complexity. Each level has a
 degree of collaboration. For example, some developers working on a
@@
 team working on another time-zone. Other tactics are suggested by the
-same author to deal with the three distances, like separating foreign
+same author to deal with the three distances, such as separating foreign
 units of developers in time-zone bands.
 
-Battin et al.~\cite{battin2001} propose and argues about their
+Battin~\etal~\cite{battin2001} propose and discuss their
 experiments using specific methodologies created for the distributed
-development centers of the Motorola Company (which has over 25
+development centers from Motorola (at the time having 25+
 software development centers worldwide). These methodologies included
 constant communication with critical units, incremental integration
-and scheduled based on time-zones distributed to developers on 6
+and schedules based on time-zones distributed to developers on 6
 countries from 3 continents.
 
-While considering free software projects instead of companies, the
-same factors are present and some methodologies
-arises. German~\cite{german2003} gives a concise review of
-methodologies used by the GNOME project, one of the most active free
-software projects and used by companies like Sun Microsystems. It is
-interesting to note a difference on viewpoint present on his paper:
-German focuses the methodology description on code. He start explaining
-that GNOME is separated into modules (76 on version 2.4, to be
-precise) and each module has one maintainer who divide his modules
-into separated parts in which other developers can work on independent
-tasks, along other responsibilities. As like others free software
-projects, all the development was done using a \emph{bugtracker} to
-bugs and issues management, mail lists and Internet Relay Chat (IRC)
-to discussion and communication and a software configuration
-management like SVN or Git. Periodical (commonly yearly) conferences
-like GUADEC is common on free and open source projects to face-to-face
-meeting and is based each time on a different place.
+In considering free software projects instead of companies,
+similar factors are present and specialized methodologies
+arise. German~\cite{german2003} provides a concise review of
+the methodologies used by the GNOME project, one of the most active of all free
+software projects. The manuscript is centered on the software architecture. It
+begins by explaining that GNOME is separated into modules (76 on version 2.4, to be
+precise) and each module has one maintainer who divides his modules into
+separate parts in which other developers can work on independent tasks, along
+other responsibilities. As in other free software projects, all 
+development was done using a \emph{bugtracker} for bug and issue management,
+mail lists and Internet Relay Chat (IRC) for discussion and communication and a
+version control system like Git or Mercurial. Periodical (commonly yearly)
+conferences like GUADEC is common on free and open source projects for
+face-to-face meetings and is based on a different place each time.
 

### 2013-02-18 paper.tex
microlog term

@@
 (e.g.\ 2 hours per day), during which they log a periodical short text
-sentence --- similar to a ``tweet'' from Twitter --- as the
+sentence or microlog --- similar to a ``tweet'' from Twitter --- as the
 status of their activity. Logging is carried out using an easy a series of
@@
 
-As noted, some strategies for GSD is based on complex methodologies
-and many of those were built for a specific company or software
-center. Here we propose an alternative methodology based on a simple
+Some of the strategies for GSD mentioned on the previous section are based on complex methodologies
+and many were built for a specific company or software
+center. We propose an alternative methodology based on a simple
 idea: small working sessions logged by a computational
@@
    \caption{
-   Mindmap of our methodology
+   A mindmap of our methodology
    }
@@
 
-From the developer perspective, the AA methodology is based on
+From the developer's perspective, the AA methodology is based on
 creating pretty small high perspective reports of what they are doing
 in a specific time frame, that can be something between 5 to 15
-minutes, depending in what is more comfortable for the developer. An
-AA Session would be a period of at least 2 hours doing these
-reports. The developer can set reminders to show up when its time to
+minutes, depending on what is most convenient for the developer and the team. An
+AA Session would be a period of at least 2 hours doing these periodical
+micrologs. The developer can set reminders to show up when its time to
 make a new report. The objective of the flexible time frame and

### 2013-02-18 paper.tex
polished heavily until prior to validation section

@@
 (e.g.\ 2 hours per day), during which they log a periodical short text
-sentence or microlog --- similar to a ``tweet'' from Twitter --- as the
+sentence or \emph{microlog} --- similar to a `tweet' from Twitter --- as the
 status of their activity. Logging is carried out using an easy a series of
@@
 minutes, depending on what is most convenient for the developer and the team. An
-AA Session would be a period of at least 2 hours doing these periodical
-micrologs. The developer can set reminders to show up when its time to
-make a new report. The objective of the flexible time frame and
-reminders is to minimize developer overhead during his AA session. In
-this way he can make the reports while staying concentrated in his
-code. Each report can be sent directly to an on-line server, or stored
-locally in a temporarily database for sending later. This make
-possible the use of AA while offline.
+\emph{AA Session} would be a period of at least 2 hours doing these periodical
+\emph{micrologs} also called \emph{AA posts}. The developer can set reminders or alerts to show up when its time to
+\emph{microlog}. The objective of the flexible time frame and
+alert scheme is to \emph{minimize developer overhead} during his AA session. In
+this way, the developer can issue micrologs while staying maximally concentrated in his
+code. Each microlog may be sent directly to an on-line server, or stored
+locally in a temporarily database for sending/pushing later. This enables
+offline micrologging and periodical alerting.
 
-Developers can also record a video screencast in the end of the
+Developers optionally record a \emph{video screencast} at the end of the
 session summarizing what has been done, explaining with his words and
-showing his most important results. This, combined with the textual
-log of his session, makes the whole report more understandable for
-himself or other people searching for information about his
-production.
+showing his most important results. This is very similar to the video logging system
+in the movie Avatar~\cite{Avatar}, although it is proved from our July 2011 git repositories and online
+wiki that we have used this powerful concept in AA prior to the release of
+Avatar. Moreover, screencasting differs from general videologging in that it
+typically captures actual workflow on
+the computer screen. Screencasting, combined with the textual log of the AA
+Session, renders the final report more understandable for the individual developer himself
+or to other people searching for information about his production.
 
@@
 
-All AA reports made by the developers are sent to a Web server and
-become public in a Website, being possible then for a manager or
-another developer to follow very closely the work of a developer,
-nearly real-time, reading each of his small reports of what he is
-doing right now.
+All AA reports made by the developers are ultimatelly sent to a Web server and
+become publically aggregated on a Website, enabling a manager or another
+developer to follow very closely the work of a given developer, nearly real-time,
+reading each of his small reports or micrologs of what he is doing right now.
 
-Another possibility is to check older sessions to see when sometimes
-was done and the comments of the developer about it, since each AA
-post happens in a very small time frame of work, the information about
-what was done become very easy to understand, instead of a long
-report in the end of a session.
+Another use case is to check older sessions to check when certain tasks
+were carried out and the comments of the developer about the \emph{process}.
+Since each AA microlog happens in a very short timeframe of work, the information
+about what was was done -- specially \emph{how} it was done -- becomes very easy to
+understand, instead of a long report in the end of a session.
 
-The site is where the developer can add to his session a link for his
-screencast about that session, a video summary of what was done to
-complement the reports, useful specially on cases where the small
-reports were done in a hurry, because the developer did not want to
+IN the current version of the AA server infrastructure, the aggregating
+website is where the developer can attach a link for his
+screencast about each session. Screencasts are specially useful on cases where
+the small reports were done in a hurry, because the developer did not want to
 lose his focus on something important at that moment.

### 2013-02-18 paper.tex
bombei inicio results and discussion

@@
 
-\subsection{Validation}
+\subsection{Peer Validation}
 
-Each AA session must be validated by another developer, it means that
-all reports are read by someone that will consider then valid or not
-and will even write commentaries about the specific session. The
-developer in charge of validating a session is decided randomly by the
-AA Web server, which send an email to the chosen developer with an
-URL to a validation interface.
+In order to prevent spamming and to improve the overall quality of AA reports,
+each AA session must be validated by another developer. More specifically,
+all reports are read by someone that will mark them as `valid' or `invalid'
+and may optionally write commentaries about the specific session and quality of
+micrologs. The developer in charge of validating any given session is randomly assigned 
+by the AA Web server, which sends an email to the chosen developer with an URL to
+a validation interface.
 
@@
 
-The easy and effective management of teams working on GSD is the main
-purpose of the AA methodology. We applied this methodology to a group
+Easy and effective GSD team management is the main
+purpose of the AA methodology. We applied the proposed methodology to a group
 of 9 developers in July of 2011 on what we called Lab
-Macambira~\footnote{LabMacambira.sf.net:
-  \url{http://labmacambira.sf.net}}. The main objective of the team
-was to work on different free software projects, contributing directly
-to their development, sending bug corrections or proposing new
-features on their source code.
+Macambira~\footnote{LabMacambira.sf.net: \url{http://labmacambira.sf.net}}. The
+main objective of the team was to work on an array of strategic free software
+projects in the broad audiovisual and web categories,
+contributing directly to their development, submitting bug patches or
+committing new features to their source code.
 
-All the team members had different levels of knowledge on software
-development as part of large and distributed free software projects
+The team members had different levels of experience on software
+development for large and distributed free software projects
 like Scilab or Mozilla. In this way, one month of training was
 conducted by two experienced developers, teaching the basics of use of
-development support tools like bugtrackers, programming languages and
-version control systems. After this period, a challenge was proposed
-for the new developers: send a bug correction or a new feature to a
-large free software project and you will work with us, you will be a
-``Macambira'' developer. Table~\ref{tabela:contribuicoes} summarizes
-the contributions of each ``Macambira'' to free software projects.
+development support tools like bugtrackers, programming languages,
+version control systems, and build systems. After this period, a starter project
+was proposed for new developers, namely to have an accepted patch or commit to
+the official repository for a large and widely used free software app.
+Developers passing the starter project would be deemed `initiated' and be
+called a `Macambira' developer.  Table~\ref{tabela:contribuicoes} summarizes the
+effective accepted contributions of each successfully initiated developer to
+free software projects in 2011.
 

### 2013-02-18 paper.tex
tab contribs

@@
 effective accepted contributions of each successfully initiated developer to
-free software projects in 2011.
+free software projects in 2011, which used the AA methodology.
 
@@
   \caption{Free and open software projects that received contributions
-    from ``Macambiras''. On the first column we can see a list of
-    applications of those projects. At right, the pseudonym of the
+    from successfully initiated developers  or `macambiras' that worked under
+    the AA methodology. The first column lists applications to which
+    contributions were officially accepted and whose development process was
+    performed and logged using AA. The second column shows the pseudonym of the
     ``Macambiras'' who sent \textit{commits} to the application. At
-    ``Lab Macambira'', and at free software community in general, is a
-    common practice to use pseudonym as identification.}
+    Lab Macambira, and at the free software community at large, 
+    It is
+    common practice at Lab Macambira, and the free software community at large,
+    to use pseudonyms as identification in order to enhance privacy in practice.}
     \begin{tabular}{|l|l|}

### 2013-02-18 paper.tex
mais mudancas

@@
 
-In one month, each developer contributed to many very large free software
-projects. Many of the developers started the training with no
-knowledge of what is free software and ended that period becoming a
-free software developer.
+In one month, each developer officially contributed to one or many free software
+projects. Many developers started the initiation training with no knowledge of what
+is free software and ended that period becoming a free software developer.
 

### 2013-02-18 paper.tex
foda: AA eh Anarquia Algoritmica

@@
 
-\subsection{Peer Validation}
+\subsection{Peer Validation and Peer Pressure}
 
-In order to prevent spamming and to improve the overall quality of AA reports,
-each AA session must be validated by another developer. More specifically,
-all reports are read by someone that will mark them as `valid' or `invalid'
-and may optionally write commentaries about the specific session and quality of
-micrologs. The developer in charge of validating any given session is randomly assigned 
-by the AA Web server, which sends an email to the chosen developer with an URL to
-a validation interface.
+No set bosses or leaders are required in an ideal aplication the AA methodology;
+in practice the role of bosses basically disappears or is greatly lightened --
+hence the name Algorithmic \emph{Autoregulation} and other implicit
+interpretations to the AA acronym and symbol. The primary mechanism to achieve
+AA is peer coordination and social behavior, be it deliberate or implicit. In
+order to prevent spamming and to improve the overall quality of AA reports, each
+AA session must be validated by another developer. More specifically, all
+reports are read by someone that will mark them as `valid' or `invalid' and may
+optionally write commentaries about the specific session and quality of
+micrologs. The developer in charge of validating any given session is randomly
+assigned by the AA Web server, which sends an email to the chosen developer with
+an URL to a validation interface.
 
@@
 is free software and ended that period becoming a free software developer.
+During that month, the same team of trainees also developed the first version of the AA
+system and used AA to manage their activities, even while developing
+the other aforementioned free software projects.
 
-During that month, the same team developed the first version of the AA
-system and used AA to manage their activities. Even while developing
-the system. All the source code of AA --- the client that sends the
-logs and the AA Web server --- is public available~\footnote{AA source
+The source code of AA --- both the client that sends the
+logs and the AA Web server --- is public available as free software~\footnote{AA source
   code:
   \url{http://labmacambira.git.sourceforge.net/git/gitweb.cgi?p=labmacambira/aa}}
-and all the AA sessions log of the whole team of ``Lab Macambira'' is
-also on-line~\footnote{Logs of AA sessions:
+and all the actual AA log data of the entire team of Lab Macambira from 2011 to
+the present time is also on-line~\footnote{Logs of AA sessions:
   \url{http://labmacambira.git.sourceforge.net/git/gitweb.cgi?p=labmacambira/paainel}}.
 
-After the training period, during more 6 months, the ``Macambiras''
-worked on a large range of free software projects, distributed on work
+After the initial training period of 1 month, the ``Macambiras''
+worked during 6 additional months on a large range of free software projects, distributed on work
 groups --- each work group focusing on a specific theme like video,
-audio and web --- and financed by contracts and support of the
-``Pont\~{a}o N\'{o}s Digitais''. In Table~\ref{tabela:criados} we can
-see a list of the free software created by ``Lab Macambira'' since
-July of 2011.
+audio and web --- and funded by contracts, freelance, and support of the
+Pont\~{a}o N\'{o}s Digitais~\footnote{\url{http://nosdigitais.teia.org.br}}. 
+Table~\ref{tabela:criados} has a list of \emph{new} free software applications created by
+``Lab Macambira'' since July 2011.
 
@@
     \caption{Software projects created by ``Lab Macambira'' since July
-      of 2011 with a short description and the technologies --- like
-      programming languages or frameworks --- involved. It is
-      interesting to note the heterogeneity of projects and its areas
+      2011 with a short description and the technologies involved (\eg,
+      programming languages or frameworks). It is
+      interesting to note the heterogeneity of projects and their areas
       of application.}
@@
 
-As of this writing the ``Lab Macambira'' have many software
-developers, and some of the trained developers continue to work
-voluntarily in the project.
+As of this writing the `Lab Macambira' unites 15 software developers, and 
+key 2011-initiated developers continue to work voluntarily in the project.
 
@@
 %% brief introduction
-In a scenario where GCD is growing as a popular form of software
-development, not just on free software projects but in the whole
-software industry, we need methodologies to deal with its
-disadvantages and at the same time to amplify its advantages.
+In a scenario where Global Software Development is growing as a popular 
+software development scenario both for free software projects and for the entire
+software industry, we need methodologies to deal with its potential disadvantages and at
+the same time to amplify its advantages.
 
-This paper has presented a methodology to GCD, being the development
-conducted on large or small groups of software developers, working on
-different countries or even at the same room. The AA methodology
+This paper has presented a methodology for GSD, \ie, the development
+involving a series of large or small groups of software developers, each either working on
+different countries or at the same room. The AA methodology
 implements a simple system where each developer take notes of his work
-generating a periodical log of small text sentences. The sum of those
-sentences, along an entire session of work, results in a complete
+posting a periodic log of small text sentences or microblogs. The sum of those
+micrologs, along with an entire session of work, results in a complete
 report. The report is made public available through a Website and be
-validated by other developers sorted aleatory by the AA Web server.
+validated by other developers randomly sorted by the AA Web server.
 
-Instead of a merely work-management tool, AA act as a methodology to
-improve the time sense of individuals, dividing their work on small
-sessions, and also reducing the need of extensive
+AA is not limited to a merely work-management tool, but acts as \emph{a
+methodology to improve the time sensibility of individuals}, dividing their
+complex work on small chunks or sessions, and also reducing the need of extensive
 reports or unnecessary meetings. By asking users to write a minimal
-text sentence as a continuously log, AA does not disturb developers
-concentrated on programming: developers just have to type some
-characters, hit \textit{enter} and go back to coding.
+text sentence as a continuous log feed, the proposed methodology avoids
+disturbing the flow of developers which are heavily concentrated on programming: developers
+just have to type a few words and go back to coding.
 
-AA application is not restricted to software development. As of this
-writing there is a comic book studio~\footnote{Pula pirata:
-  \url{http://pulapirata.com}} starting to use AA to manage their
-activities. People with non-software background, like social
-scientists, musicians and activists has also using AA and contributing
-for its improvement.
+The AA methodology is not restricted to software development, even though it was
+built for the latter. As of this
+writing there is an entertainment studio, Pula Pirata~\footnote{
+\url{www.pulapirata.com}}, that has started to use AA to manage their creative activities.
+Other people with no software background, like social scientists, musicians and
+activists also have been using AA and contributing for its improvement.
 
-For developer teams, we have experienced the use of AA to
-auto-regulate the work of ``Lab Macambira'', a group of free software
-developers from Brazil. Since July of 2011 the group have contributed
-and created new free and open source software for a vast number of
-applications.
+For developer teams, we have analyzed data and gained practical experience on
+the use of AA to auto-regulate the work of Lab Macambira, a group of free
+software developers from Brazil. Since July of 2011 the group has contributed
+and created new free and open source software for a vast number of additional applications.
 
-There are many aspects of the work which remain unfinished. New ways
-to report logs --- the ``Twitter like'' messages --- from different
-interfaces like IRC, Internet Messaging services and email can make
-the use of AA easy and widespread, turning AA an ubiquitous system,
-presented on everyday communication channels. Even the work logs
-generated since July of 2011 could be vastly statistically analyzed
-aiming to recognize patterns in the behavior of individuals and their
-productions.
-
-We would like to conclude setting an important role of AA: being a
-free software system and an open methodology, AA could be used to
-auto-manage groups of individuals working on software or other kinds
-of activities. In this way, we are interested to spread AA for those
-groups, to have even more developers contributing in a collaborative
-way.
+There are many aspects of the work which remain unfinished. Other client
+interfaces to report micrologs -- the `Twitter-like' messages -- from different
+interfaces beyond IRC, \eg, internet social services and email, can make the use of
+AA easy and widespread, turning AA an ubiquitous system, presented on everyday
+communication channels. Even the actual work logs generated by the Lab Macambira
+and Pula Pirata collectives since July of 2011 could be
+vastly statistically analyzed aiming to recognize patterns in the behavior of
+individuals and their productions.
 
@@
 manuscript was written using AA. The complete log is on-line at
-\url{http://www.pulapirata.com/skills/aa}.
+\url{www.pulapirata.com/skills/aa}.
 

### 2013-02-18 paper.tex
paper inteiro basicamente pronto pra send. falta todos e detalhes menores.

@@
 The AA methodology is not restricted to software development, even though it was
-built for the latter. As of this
+designed for the latter. As of this
 writing there is an entertainment studio, Pula Pirata~\footnote{
@@
 
-There are many aspects of the work which remain unfinished. Other client
-interfaces to report micrologs -- the `Twitter-like' messages -- from different
-interfaces beyond IRC, \eg, internet social services and email, can make the use of
-AA easy and widespread, turning AA an ubiquitous system, presented on everyday
-communication channels. Even the actual work logs generated by the Lab Macambira
-and Pula Pirata collectives since July of 2011 could be
-vastly statistically analyzed aiming to recognize patterns in the behavior of
-individuals and their productions.
+There are many aspects of the work which remain unfinished. Additional
+ubiquitous client
+interfaces for micrologging -- the `Twitter-like' messages -- from different
+interfaces beyond IRC, \eg, web social services and email, would greatly make the use of
+AA easier and more widespread, turning AA a truly ubiquitous and replicable system, presented on everyday
+communication channels. Another research direction is that the actual work logs
+generated by the Lab Macambira and Pula Pirata collectives since July of 2011
+could be vastly statistically analyzed aiming to recognize patterns in the
+behavior of individuals and their creative production process.
 
 \section*{Acknowledgments}
-%The authors would like to thank NSF and CNPq.
+%TODO The authors would like to thank NSF and CNPq.
 
-We would like to also thank AA: the present research and even this
+We would like to thank AA: the present research and even this
 manuscript was written using AA. The complete log is on-line at

### 2013-02-18 paper.tex
ie,etc,commands
tablesize

@@
 \usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}
+
+\newcommand{\ie}{{\it i.e.}}
+\newcommand{\etc}{{\it etc}}
+\newcommand{\eg}{{\it e.g.}}
+\newcommand{\wrt}{{\it w.r.t. }}
+\newcommand{\etal}{{\it et.~al.}}
 
@@
 \begin{center}
-   \includegraphics[width=0.95\linewidth]{figs/aa-0_1.png}
+   \includegraphics[width=\linewidth]{figs/aa-0_1.png}
 \end{center}
-   \caption{AA Version 0.1}
+   \caption{AA Report Agregator Version 0.1}
 \label{fig:aaserver}
@@
 \begin{table}
-    \caption{Software projects created by ``Lab Macambira'' since July
+    \caption{Software projects created by Lab Macambira since July
       2011 with a short description and the technologies involved (\eg,
@@
       of application.}
+
+    \footnotesize{
     \begin{tabular}{|l|p{5cm}|l|}
@@
         \hline
-        \end{tabular}
+        \end{tabular}}
     \label{tabela:criados}

### 2013-02-18 paper.tex
enderecos bons agora

@@
 %\documentclass[10pt,twocolumn,letterpaper]{article}
-\documentclass[letterpaper]{article}
+\documentclass{article}
 
@@
 \usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}
+%\usepackage{draftwatermark}
 
@@
 %\newcommand{\draftnote}[1]{\marginpar{\tiny\raggedright\textsf{\hspace{0pt}#1}}}
-\newcommand{\draftnote}[1]{\marginpar{\tiny\raggedright\textsf{\hspace{0pt}#1}}}
-%\newcommand{\draftnote}[1]{}
+%\newcommand{\draftnote}[1]{\marginpar{\tiny\raggedright\textsf{\hspace{0pt}#1}}}
+\newcommand{\draftnote}[1]{}
 
 % This one is just for the comments for in-line text.
-\newcommand{\indraftnote}[1]{\textcolor{blue}{\texttt{\footnotesize[#1]}}}
+%\newcommand{\indraftnote}[1]{\textcolor{blue}{\texttt{\footnotesize[#1]}}}
+\newcommand{\indraftnote}[1]{}
 \newcommand{\todo}[1]{\indraftnote{todo: #1}}
-%\newcommand{\indraftnote}[1]{}
 
@@
 or to do other bureaucratic activities such as writing lengthy reports to
-justify their funding.~\cite{Thompson:Wired:2012}
+justify their funding.\cite{Thompson:Wired:2012}
 %\todo{ler mais. cacm, etc}
@@
 showing his most important results. This is very similar to the video logging system
-in the movie Avatar~\cite{Avatar}, although it is proved from our July 2011 git repositories and online
+in the movie Avatar, although it is proved from our July 2011 git repositories and online
 wiki that we have used this powerful concept in AA prior to the release of

### 2013-02-18 front-matter.tex
enderecos bons agora
address improved

@@
 \author{%
-Renato~Fabbri \and Ricardo~Fabbri \and Vilson Vieira \and Alexandre Negrao \and Lucas Zambianchi
-\and Marcos Mendonca \and Danilo Shiga
+Renato~Fabbri \and Ricardo~Fabbri \and Vilson Vieira \and Alexandre Negr\~{a}o \and Lucas Zambianchi
+\and Marcos Mendon\c{c}a \and Danilo Shiga\\[1em]
+\small{
+Instituto de F\'{i}sica de S\~{a}o Carlos (IFSC), Universidade de
+S\~{a}o Paulo (USP), Brazil}\\[1em]
+\small{Instituto Polit\'{e}cnico da Universidade do Estado do Rio de
+Janeiro, Nova Friburgo, RJ, Brazil}\\[1em]
+\url{LabMacambira.sourceforge.net}.
 }

### 2013-02-18 front-matter.tex
address improved

@@
 Instituto de F\'{i}sica de S\~{a}o Carlos (IFSC), Universidade de
-S\~{a}o Paulo (USP), Brazil}\\[1em]
-\small{Instituto Polit\'{e}cnico da Universidade do Estado do Rio de
-Janeiro, Nova Friburgo, RJ, Brazil}\\[1em]
+S\~{a}o Paulo (USP), Brazil}\\[0.5em]
+\small{Instituto Polit\'{e}cnico, Universidade do Estado do Rio de
+Janeiro, Nova Friburgo, RJ, Brazil}\\[0.5em]
 \url{LabMacambira.sourceforge.net}.

### 2013-05-31 front-matter.tex
arrumado abstract e titulo

--- 2012-06-26 front-matter.tex
@@
 \title{
-AA Distributed Software Development Methodology
+    AA: the Algoritmic Autoregulation Software Development Methodology
 }
@@
 \begin{abstract}
-We present a new self-regulating methodology for coordinating
-distributed team work called Algorithmic Auto-regulation (AA), based
-on recent social networking concepts and individual merit. Team
-members take on an egalitarian role, and stay voluntarily logged into
-so-called AA sessions for part of their time (e.g.\ 2 hours per day),
-during which they create periodical logs --- short text sentences ---
-they wish to share about their activity with the team. These logs are
+We present a new methodology for coordinating
+teamwork called Algorithmic Autoregulation (AA).
+This methodology is based on recent social networking concepts and individual merit and exhibits convenient asynchronous, autodocumenting and autoregulating characteristics.
+Team
+members take on an egalitarian and volutary role, by logging into
+periodic ``AA sessions'' for an arbitrary duration (e.g.\ 2 hours per day).
+During each session, a user creates a log composed of short text sentences
+about their activity. These logs are
 publicly aggregated in a Website and are peer-validated, as in code
-review. A video is generally recorded by the members to make their
-work logs more understandable. This methodology is well-suited for
-increasing the efficiency of distributed teams working on what is
-called Global Software Development (GSD) through asynchronous
-on-demand communication, reducing the need for central management,
-unproductive meetings or time-consuming reports. The AA methodology
-also legitimizes the activities of a distributed software team.  It
-thus enables entities to have a solid means to fund these activities,
-allowing for new and concrete business models to emerge for very
-distributed software development. AA as a methodology is also at the
-core of having self-replicating hacker initiatives. These claims are
-discussed in a real case-study of running a distributed free software
+review. A short screencast is ideally recorded at the end of each session to make
+AA logs more understandable. This methodology seems to be well-suited for
+increasing the efficiency of teams working on
+Global Software Development (GSD). Recognized reasons for this are
+the: 1) builtin asynchronous
+on-demand communication, documentation and working hours comprobatives;
+2) reduced need for central management,
+meetings and time-consuming reports. Hence, AA
+legitimizes and eases activities of a distributed software team.  It
+enables groups to have new means to fund activities,
+allowing for business models to emerge from
+distributed software development. AA is proposed,
+at it's core, as a way of having self-replicating hacker initiatives. These claims are
+discussed in a real case-study of a distributed free software
 hacker team called Lab Macambira.

### 2013-05-31 paper.tex
address improved
pequenas melhoras e amadurecimentos, revisao ainda na introducao

--- 2013-02-18 paper.tex
@@
 One of the defining features of modern times is the widening
-geographical distribution of software teams~\cite{last2003} creating
-what is called Global Software Development (henceforth
-GSD)~\cite{german2003}. An example is the free software
+geographical distribution of software teams~\cite{last2003}.
+This is responsible for the so called Global Software Development (GSD)~\cite{german2003}. 
+An example is the free software
 movement. Projects and institutions like Mozilla Foundation has several
@@
 
-Despite the advantages of GCD, we have noticed how
+However, dispite the advantages of GCD, we have noticed how
 difficult it is to coordinate and fund free software on a larger scale
@@
 not only volunteers and very experienced developers, but also
-contractors and freelancers from different backgrounds and cultures. Our observations
-are founded on the factors suggested by Carmel~\cite{carmel1999} as
+contractors from different backgrounds and cultures. Our observations
+found base on the factors suggested by Carmel~\cite{carmel1999} as
 main difficulties for GCD: distance, time and cultural differences. In
@@
 become the norm to participate in too many meetings with the ``laptop
-open'', which can be un-productive. Software developers
+open'', which is unproductive at the very least. Software developers
 like to code, to be productive, to have their hands on their project,
-to do what they are best at. They dislike to forcibly stop for meetings
-or to do other bureaucratic activities such as writing lengthy reports to
-justify their funding.~\cite{Thompson:Wired:2012}
+to do what they are best at. They dislike to have to stop for meetings
+or to write lengthy reports to justify their funding.
 %\todo{ler mais. cacm, etc}
 
-In this paper we propose the AA methodology and an associated software system
-for coordinating distributed team work, tackling the disadvantages
+Here we propose AA, a methodology and an associated software system
+for coordinating distributed team work dealing with the disadvantages
 of GSD. Team members take on an egalitarian role, and stay
@@
 (e.g.\ 2 hours per day), during which they log a periodical short text
-sentence or \emph{microlog} --- similar to a `tweet' from Twitter --- as the
-status of their activity. Logging is carried out using an easy a series of
-available UI's: unix shell
-commands, native GUI or web page, conventional social network posts, or chat messages to a log bot
-listening to IRC, GTalk, G+, and others.  These ``microblog sentences'' are publicly aggregated and validated by other
-team members. Through AA, we have a methodology and an associated system to help
-implement and validate the activities of a distributed software team. It implicitly
-legitimizes financial support for the expansion of the activity of the developer
-team. The AA methodology is specially useful for coordinating distributed and
-decentralized team work, providing effective means to asynchronously update
-different team members without the need for synchronous unproductive meetings.
+sentence --- similar to a ``tweet'' in this Twitter era --- as the
+status of their activity using an easy to use command. These
+``microblog sentences'' are publicly aggregated and validated by other
+team members. Through AA, we have a methodology and an associated
+system to validate and enable the activities of a distributed software
+team. It implicitly legitimizes financial support for the expansion of
+the activity of the developer team. The AA methodology is specially
+useful for coordinating distributed and decentralized team work,
+providing effective means to asynchronously update different team
+members without the need for synchronous unproductive meetings.
 
@@
 
-A brief overview of current work on GSD methodologies related to
+A brief overview of current papers about GSD methodologies related with
 AA is presented in Section~\ref{related-work}, while in
@@
 characteristics of the AA methodology followed by the description of
-our experience using AA in a team of 9 paid developers 
-% nao era mais de 9?
-since July 2011. In Section~\ref{conclusions} we draw some final
-conclusions and indicate future possibilities for the
-practical use of AA in other types of software developers teams or
-organizations working on non-software distributed activities.
+our experience using AA in a team of 9 developers since July of
+2011. Finally, in Section~\ref{conclusions}, we draw some final
+conclusions and indicate future possibilities to the
+practical use of AA on other teams of software developers or
+individuals working on non-software activities.
 
@@
 and comprising three distances: geographical, cultural and
-temporal.
-
-First, geographical distance handicaps ($i$) \emph{coordination}, the act of
-integrating all the tasks distributed between units~\cite{carmel2001}; ($ii$)
-\emph{control}, or the process to maintain specific goals, policies or quality
-levels; and ($iii$) \emph{communication}. All those factors are correlated, \eg,
-a team needs to have clear communication to work on tasks of a specific problem.
-
-Second, cultural distance encompasses differences on organizational and
+temporal. Geographical distance prejudice \emph{coordination}, being
+the act of integrating all the tasks distributed between
+units~\cite{carmel2001}; \emph{control}, or the process to maintain
+specific goals, policies or quality levels; and
+\emph{communication}. All these factors are correlated, for example, a
+team need to have clearly communication to work on tasks of a specific
+problem. Cultural distance encompass differences on organizational and
 natural culture. Spoken language, unit and ethnic values are common
-forms of such distance. Some companies prefer to allocate development
-units in foreign locations with minimal cultural distance (\eg, an
-American company may prefer Ireland due to spoken language
-similarity~\cite{carmel2001}). Third, the temporal distance that
-hampers synchronous communications like telephone or
+forms of this distance. Some companies prefer to situate development
+units in foreign locations with minimal cultural distance (e.g.\ an
+American firm prefer Ireland, because of spoken language
+similarity~\cite{carmel2001}). And finally the temporal distance that
+prejudices synchronous communications like telephone or
 videoconferences. Units of developers working on different time-zones
@@
 Targeting geographical distance, Carmel~\cite{carmel2001} suggests a
-strategy to reduce intensive collaboration. His approach divides the
-whole software life-cycle into levels of complexity. Each level has a
+tactic to reduce intensive collaboration. His approach divides the
+whole software life-cycle on levels of complexity. Each level has a
 degree of collaboration. For example, some developers working on a
@@
 team working on another time-zone. Other tactics are suggested by the
-same author to deal with the three distances, such as separating foreign
+same author to deal with the three distances, like separating foreign
 units of developers in time-zone bands.
 
-Battin~\etal~\cite{battin2001} propose and discuss their
+Battin et al.~\cite{battin2001} propose and argues about their
 experiments using specific methodologies created for the distributed
-development centers from Motorola (at the time having 25+
+development centers of the Motorola Company (which has over 25
 software development centers worldwide). These methodologies included
 constant communication with critical units, incremental integration
-and schedules based on time-zones distributed to developers on 6
+and scheduled based on time-zones distributed to developers on 6
 countries from 3 continents.
 
-In considering free software projects instead of companies,
-similar factors are present and specialized methodologies
-arise. German~\cite{german2003} provides a concise review of
-the methodologies used by the GNOME project, one of the most active of all free
-software projects. The manuscript is centered on the software architecture. It
-begins by explaining that GNOME is separated into modules (76 on version 2.4, to be
-precise) and each module has one maintainer who divides his modules into
-separate parts in which other developers can work on independent tasks, along
-other responsibilities. As in other free software projects, all 
-development was done using a \emph{bugtracker} for bug and issue management,
-mail lists and Internet Relay Chat (IRC) for discussion and communication and a
-version control system like Git or Mercurial. Periodical (commonly yearly)
-conferences like GUADEC is common on free and open source projects for
-face-to-face meetings and is based on a different place each time.
+While considering free software projects instead of companies, the
+same factors are present and some methodologies
+arises. German~\cite{german2003} gives a concise review of
+methodologies used by the GNOME project, one of the most active free
+software projects and used by companies like Sun Microsystems. It is
+interesting to note a difference on viewpoint present on his paper:
+German focuses the methodology description on code. He start explaining
+that GNOME is separated into modules (76 on version 2.4, to be
+precise) and each module has one maintainer who divide his modules
+into separated parts in which other developers can work on independent
+tasks, along other responsibilities. As like others free software
+projects, all the development was done using a \emph{bugtracker} to
+bugs and issues management, mail lists and Internet Relay Chat (IRC)
+to discussion and communication and a software configuration
+management like SVN or Git. Periodical (commonly yearly) conferences
+like GUADEC is common on free and open source projects to face-to-face
+meeting and is based each time on a different place.
 
@@
 
-Some of the strategies for GSD mentioned on the previous section are based on complex methodologies
-and many were built for a specific company or software
-center. We propose an alternative methodology based on a simple
+As noted, some strategies for GSD is based on complex methodologies
+and many of those were built for a specific company or software
+center. Here we propose an alternative methodology based on a simple
 idea: small working sessions logged by a computational
@@
    \caption{
-   A mindmap of our methodology
+   Mindmap of our methodology
    }
@@
 
-From the developer's perspective, the AA methodology is based on
+From the developer perspective, the AA methodology is based on
 creating pretty small high perspective reports of what they are doing
 in a specific time frame, that can be something between 5 to 15
-minutes, depending on what is most convenient for the developer and the team. An
-\emph{AA Session} would be a period of at least 2 hours doing these periodical
-\emph{micrologs} also called \emph{AA posts}. The developer can set reminders or alerts to show up when its time to
-\emph{microlog}. The objective of the flexible time frame and
-alert scheme is to \emph{minimize developer overhead} during his AA session. In
-this way, the developer can issue micrologs while staying maximally concentrated in his
-code. Each microlog may be sent directly to an on-line server, or stored
-locally in a temporarily database for sending/pushing later. This enables
-offline micrologging and periodical alerting.
+minutes, depending in what is more comfortable for the developer. An
+AA Session would be a period of at least 2 hours doing these
+reports. The developer can set reminders to show up when its time to
+make a new report. The objective of the flexible time frame and
+reminders is to minimize developer overhead during his AA session. In
+this way he can make the reports while staying concentrated in his
+code. Each report can be sent directly to an on-line server, or stored
+locally in a temporarily database for sending later. This make
+possible the use of AA while offline.
 
-Developers optionally record a \emph{video screencast} at the end of the
+Developers can also record a video screencast in the end of the
 session summarizing what has been done, explaining with his words and
-showing his most important results. This is very similar to the video logging system
-in the movie Avatar~\cite{Avatar}, although it is proved from our July 2011 git repositories and online
-wiki that we have used this powerful concept in AA prior to the release of
-Avatar. Moreover, screencasting differs from general videologging in that it
-typically captures actual workflow on
-the computer screen. Screencasting, combined with the textual log of the AA
-Session, renders the final report more understandable for the individual developer himself
-or to other people searching for information about his production.
+showing his most important results. This, combined with the textual
+log of his session, makes the whole report more understandable for
+himself or other people searching for information about his
+production.
 
@@
 
-All AA reports made by the developers are ultimatelly sent to a Web server and
-become publically aggregated on a Website, enabling a manager or another
-developer to follow very closely the work of a given developer, nearly real-time,
-reading each of his small reports or micrologs of what he is doing right now.
+All AA reports made by the developers are sent to a Web server and
+become public in a Website, being possible then for a manager or
+another developer to follow very closely the work of a developer,
+nearly real-time, reading each of his small reports of what he is
+doing right now.
 
-Another use case is to check older sessions to check when certain tasks
-were carried out and the comments of the developer about the \emph{process}.
-Since each AA microlog happens in a very short timeframe of work, the information
-about what was was done -- specially \emph{how} it was done -- becomes very easy to
-understand, instead of a long report in the end of a session.
+Another possibility is to check older sessions to see when sometimes
+was done and the comments of the developer about it, since each AA
+post happens in a very small time frame of work, the information about
+what was done become very easy to understand, instead of a long
+report in the end of a session.
 
-IN the current version of the AA server infrastructure, the aggregating
-website is where the developer can attach a link for his
-screencast about each session. Screencasts are specially useful on cases where
-the small reports were done in a hurry, because the developer did not want to
+The site is where the developer can add to his session a link for his
+screencast about that session, a video summary of what was done to
+complement the reports, useful specially on cases where the small
+reports were done in a hurry, because the developer did not want to
 lose his focus on something important at that moment.
 
-\subsection{Peer Validation and Peer Pressure}
+\subsection{Validation}
 
-No set bosses or leaders are required in an ideal aplication the AA methodology;
-in practice the role of bosses basically disappears or is greatly lightened --
-hence the name Algorithmic \emph{Autoregulation} and other implicit
-interpretations to the AA acronym and symbol. The primary mechanism to achieve
-AA is peer coordination and social behavior, be it deliberate or implicit. In
-order to prevent spamming and to improve the overall quality of AA reports, each
-AA session must be validated by another developer. More specifically, all
-reports are read by someone that will mark them as `valid' or `invalid' and may
-optionally write commentaries about the specific session and quality of
-micrologs. The developer in charge of validating any given session is randomly
-assigned by the AA Web server, which sends an email to the chosen developer with
-an URL to a validation interface.
+Each AA session must be validated by another developer, it means that
+all reports are read by someone that will consider then valid or not
+and will even write commentaries about the specific session. The
+developer in charge of validating a session is decided randomly by the
+AA Web server, which send an email to the chosen developer with an
+URL to a validation interface.
 
@@
 
-Easy and effective GSD team management is the main
-purpose of the AA methodology. We applied the proposed methodology to a group
+The easy and effective management of teams working on GSD is the main
+purpose of the AA methodology. We applied this methodology to a group
 of 9 developers in July of 2011 on what we called Lab
-Macambira~\footnote{LabMacambira.sf.net: \url{http://labmacambira.sf.net}}. The
-main objective of the team was to work on an array of strategic free software
-projects in the broad audiovisual and web categories,
-contributing directly to their development, submitting bug patches or
-committing new features to their source code.
+Macambira~\footnote{LabMacambira.sf.net:
+  \url{http://labmacambira.sf.net}}. The main objective of the team
+was to work on different free software projects, contributing directly
+to their development, sending bug corrections or proposing new
+features on their source code.
 
-The team members had different levels of experience on software
-development for large and distributed free software projects
+All the team members had different levels of knowledge on software
+development as part of large and distributed free software projects
 like Scilab or Mozilla. In this way, one month of training was
 conducted by two experienced developers, teaching the basics of use of
-development support tools like bugtrackers, programming languages,
-version control systems, and build systems. After this period, a starter project
-was proposed for new developers, namely to have an accepted patch or commit to
-the official repository for a large and widely used free software app.
-Developers passing the starter project would be deemed `initiated' and be
-called a `Macambira' developer.  Table~\ref{tabela:contribuicoes} summarizes the
-effective accepted contributions of each successfully initiated developer to
-free software projects in 2011, which used the AA methodology.
+development support tools like bugtrackers, programming languages and
+version control systems. After this period, a challenge was proposed
+for the new developers: send a bug correction or a new feature to a
+large free software project and you will work with us, you will be a
+``Macambira'' developer. Table~\ref{tabela:contribuicoes} summarizes
+the contributions of each ``Macambira'' to free software projects.
 
@@
   \caption{Free and open software projects that received contributions
-    from successfully initiated developers  or `macambiras' that worked under
-    the AA methodology. The first column lists applications to which
-    contributions were officially accepted and whose development process was
-    performed and logged using AA. The second column shows the pseudonym of the
+    from ``Macambiras''. On the first column we can see a list of
+    applications of those projects. At right, the pseudonym of the
     ``Macambiras'' who sent \textit{commits} to the application. At
-    Lab Macambira, and at the free software community at large, 
-    It is
-    common practice at Lab Macambira, and the free software community at large,
-    to use pseudonyms as identification in order to enhance privacy in practice.}
+    ``Lab Macambira'', and at free software community in general, is a
+    common practice to use pseudonym as identification.}
     \begin{tabular}{|l|l|}
@@
 
-In one month, each developer officially contributed to one or many free software
-projects. Many developers started the initiation training with no knowledge of what
-is free software and ended that period becoming a free software developer.
-During that month, the same team of trainees also developed the first version of the AA
-system and used AA to manage their activities, even while developing
-the other aforementioned free software projects.
+In one month, each developer contributed to many very large free software
+projects. Many of the developers started the training with no
+knowledge of what is free software and ended that period becoming a
+free software developer.
 
-The source code of AA --- both the client that sends the
-logs and the AA Web server --- is public available as free software~\footnote{AA source
+During that month, the same team developed the first version of the AA
+system and used AA to manage their activities. Even while developing
+the system. All the source code of AA --- the client that sends the
+logs and the AA Web server --- is public available~\footnote{AA source
   code:
   \url{http://labmacambira.git.sourceforge.net/git/gitweb.cgi?p=labmacambira/aa}}
-and all the actual AA log data of the entire team of Lab Macambira from 2011 to
-the present time is also on-line~\footnote{Logs of AA sessions:
+and all the AA sessions log of the whole team of ``Lab Macambira'' is
+also on-line~\footnote{Logs of AA sessions:
   \url{http://labmacambira.git.sourceforge.net/git/gitweb.cgi?p=labmacambira/paainel}}.
 
-After the initial training period of 1 month, the ``Macambiras''
-worked during 6 additional months on a large range of free software projects, distributed on work
+After the training period, during more 6 months, the ``Macambiras''
+worked on a large range of free software projects, distributed on work
 groups --- each work group focusing on a specific theme like video,
-audio and web --- and funded by contracts, freelance, and support of the
-Pont\~{a}o N\'{o}s Digitais~\footnote{\url{http://nosdigitais.teia.org.br}}. 
-Table~\ref{tabela:criados} has a list of \emph{new} free software applications created by
-``Lab Macambira'' since July 2011.
+audio and web --- and financed by contracts and support of the
+``Pont\~{a}o N\'{o}s Digitais''. In Table~\ref{tabela:criados} we can
+see a list of the free software created by ``Lab Macambira'' since
+July of 2011.
 
@@
     \caption{Software projects created by ``Lab Macambira'' since July
-      2011 with a short description and the technologies involved (\eg,
-      programming languages or frameworks). It is
-      interesting to note the heterogeneity of projects and their areas
+      of 2011 with a short description and the technologies --- like
+      programming languages or frameworks --- involved. It is
+      interesting to note the heterogeneity of projects and its areas
       of application.}
@@
 
-As of this writing the `Lab Macambira' unites 15 software developers, and 
-key 2011-initiated developers continue to work voluntarily in the project.
+As of this writing the ``Lab Macambira'' have many software
+developers, and some of the trained developers continue to work
+voluntarily in the project.
 
@@
 %% brief introduction
-In a scenario where Global Software Development is growing as a popular 
-software development scenario both for free software projects and for the entire
-software industry, we need methodologies to deal with its potential disadvantages and at
-the same time to amplify its advantages.
+In a scenario where GCD is growing as a popular form of software
+development, not just on free software projects but in the whole
+software industry, we need methodologies to deal with its
+disadvantages and at the same time to amplify its advantages.
 
-This paper has presented a methodology for GSD, \ie, the development
-involving a series of large or small groups of software developers, each either working on
-different countries or at the same room. The AA methodology
+This paper has presented a methodology to GCD, being the development
+conducted on large or small groups of software developers, working on
+different countries or even at the same room. The AA methodology
 implements a simple system where each developer take notes of his work
-posting a periodic log of small text sentences or microblogs. The sum of those
-micrologs, along with an entire session of work, results in a complete
+generating a periodical log of small text sentences. The sum of those
+sentences, along an entire session of work, results in a complete
 report. The report is made public available through a Website and be
-validated by other developers randomly sorted by the AA Web server.
+validated by other developers sorted aleatory by the AA Web server.
 
-AA is not limited to a merely work-management tool, but acts as \emph{a
-methodology to improve the time sensibility of individuals}, dividing their
-complex work on small chunks or sessions, and also reducing the need of extensive
+Instead of a merely work-management tool, AA act as a methodology to
+improve the time sense of individuals, dividing their work on small
+sessions, and also reducing the need of extensive
 reports or unnecessary meetings. By asking users to write a minimal
-text sentence as a continuous log feed, the proposed methodology avoids
-disturbing the flow of developers which are heavily concentrated on programming: developers
-just have to type a few words and go back to coding.
+text sentence as a continuously log, AA does not disturb developers
+concentrated on programming: developers just have to type some
+characters, hit \textit{enter} and go back to coding.
 
-The AA methodology is not restricted to software development, even though it was
-designed for the latter. As of this
-writing there is an entertainment studio, Pula Pirata~\footnote{
-\url{www.pulapirata.com}}, that has started to use AA to manage their creative activities.
-Other people with no software background, like social scientists, musicians and
-activists also have been using AA and contributing for its improvement.
+AA application is not restricted to software development. As of this
+writing there is a comic book studio~\footnote{Pula pirata:
+  \url{http://pulapirata.com}} starting to use AA to manage their
+activities. People with non-software background, like social
+scientists, musicians and activists has also using AA and contributing
+for its improvement.
 
-For developer teams, we have analyzed data and gained practical experience on
-the use of AA to auto-regulate the work of Lab Macambira, a group of free
-software developers from Brazil. Since July of 2011 the group has contributed
-and created new free and open source software for a vast number of additional applications.
+For developer teams, we have experienced the use of AA to
+auto-regulate the work of ``Lab Macambira'', a group of free software
+developers from Brazil. Since July of 2011 the group have contributed
+and created new free and open source software for a vast number of
+applications.
 
-There are many aspects of the work which remain unfinished. Additional
-ubiquitous client
-interfaces for micrologging -- the `Twitter-like' messages -- from different
-interfaces beyond IRC, \eg, web social services and email, would greatly make the use of
-AA easier and more widespread, turning AA a truly ubiquitous and replicable system, presented on everyday
-communication channels. Another research direction is that the actual work logs
-generated by the Lab Macambira and Pula Pirata collectives since July of 2011
-could be vastly statistically analyzed aiming to recognize patterns in the
-behavior of individuals and their creative production process.
+There are many aspects of the work which remain unfinished. New ways
+to report logs --- the ``Twitter like'' messages --- from different
+interfaces like IRC, Internet Messaging services and email can make
+the use of AA easy and widespread, turning AA an ubiquitous system,
+presented on everyday communication channels. Even the work logs
+generated since July of 2011 could be vastly statistically analyzed
+aiming to recognize patterns in the behavior of individuals and their
+productions.
+
+We would like to conclude setting an important role of AA: being a
+free software system and an open methodology, AA could be used to
+auto-manage groups of individuals working on software or other kinds
+of activities. In this way, we are interested to spread AA for those
+groups, to have even more developers contributing in a collaborative
+way.
 
 \section*{Acknowledgments}
-%TODO The authors would like to thank NSF and CNPq.
+%The authors would like to thank NSF and CNPq.
 
-We would like to thank AA: the present research and even this
+We would like to also thank AA: the present research and even this
 manuscript was written using AA. The complete log is on-line at
-\url{www.pulapirata.com/skills/aa}.
+\url{http://www.pulapirata.com/skills/aa}.
 

### 2013-05-31 paper.tex
feita correcao da intro

@@
 many countries. The same is true for GNOME~\cite{german2003}, OpenBSD,
-MySQL or Apache Software Foundation, to cite just a few of the most active
-projects.\footnote{Ohloh, the open source network, have a more complete
+MySQL or Apache Software Foundation, to cite just a few\footnote{Ohloh, the open source network, have a more complete
   and constantly updated list of the most active projects on-line at
-  \url{www.ohloh.net}} Along the free and open software
-projects, GSD is growing popularity in every niche of the software
+  \url{www.ohloh.net}}. Along the free and open software
+projects, GSD has a growing popularity in every niche of the software
 industry as a whole, even on those distributing their software with
@@
 
-However, dispite the advantages of GCD, we have noticed how
+Dispite the advantages of GSD, it is known how
 difficult it is to coordinate and fund free software on a larger scale
-than currently available, when teams are very heterogeneous containing
-not only volunteers and very experienced developers, but also
-contractors from different backgrounds and cultures. Our observations
-found base on the factors suggested by Carmel~\cite{carmel1999} as
-main difficulties for GCD: distance, time and cultural differences. In
+than currently available. That is, when teams are very heterogeneous, containing
+not only volunteers and experienced developers, but also
+contractors from different backgrounds and cultures. This observations
+are in consonance with dedicated studies, such as~\cite{carmel1999}.
+As a common base, related difficulties for GSD gather around distance, time and cultural differences. In
 the case of free or open software projects, all these factors are
@@
 collectives are frequent ineffective meetings, which are seldom
-focused to the interest to any attendant. The result is that it has
-become the norm to participate in too many meetings with the ``laptop
-open'', which is unproductive at the very least. Software developers
+focused on the interest of any attendant. The result is that it has
+become the norm to participate in these with the ``laptop
+open'', which is unproductive. Software developers
 like to code, to be productive, to have their hands on their project,
@@
 
-Here we propose AA, a methodology and an associated software system
+To address these matters is the purpose of AA, a methodology and an associated software system
 for coordinating distributed team work dealing with the disadvantages
@@
 ``microblog sentences'' are publicly aggregated and validated by other
-team members. Through AA, we have a methodology and an associated
+team members. Through AA, the community has a methodology and an associated
 system to validate and enable the activities of a distributed software
 team. It implicitly legitimizes financial support for the expansion of
-the activity of the developer team. The AA methodology is specially
+the activity of the developer team as the activities are documented and validated.
+The AA methodology is specially
 useful for coordinating distributed and decentralized team work,
 providing effective means to asynchronously update different team
-members without the need for synchronous unproductive meetings.
+members, diminishing the need for harsh synchronous meetings.
 
@@
 A brief overview of current papers about GSD methodologies related with
-AA is presented in Section~\ref{related-work}, while in
-Section~\ref{aa-methodology} we describe the most relevant
-characteristics of the AA methodology followed by the description of
-our experience using AA in a team of 9 developers since July of
-2011. Finally, in Section~\ref{conclusions}, we draw some final
-conclusions and indicate future possibilities to the
+AA is presented in section~\ref{related-work}. In
+section~\ref{aa-methodology}, the most relevant
+characteristics of the AA methodology are outlined. In this same section,
+there is a description of
+an AA use in a team of 9 developers in the second half of
+2011 and a broader a more released use of AA from 2012 until now. Finally, in Section~\ref{conclusions}, there are final
+conclusions and indicatives of future possibilities to the
 practical use of AA on other teams of software developers or
@@
 %% brief introduction
-In a scenario where GCD is growing as a popular form of software
+In a scenario where GSD is growing as a popular form of software
 development, not just on free software projects but in the whole
@@
 
-This paper has presented a methodology to GCD, being the development
+This paper has presented a methodology to GSD, being the development
 conducted on large or small groups of software developers, working on

### 2013-05-31 front-matter.tex
acho que soh o titulo que mudei
minor fix in abstract

@@
 \title{
-    AA: the Algoritmic Autoregulation Software Development Methodology
+    AA: the Algorithmic Auto-regulation (Software Development) Methodology
 }
@@
 \author{%
-Renato~Fabbri \and Ricardo~Fabbri \and Vilson Vieira \and Alexandre Negrao \and Lucas Zambianchi
+Renato~Fabbri \and Ricardo~Fabbri \and Vilson Vieira \and Alexandre Negr\~{a}o \and Lucas Zambianchi
 \and Marcos Mendonca \and Danilo Shiga
@@
 \begin{abstract}
-We present a new methodology for coordinating
-teamwork called Algorithmic Autoregulation (AA).
-This methodology is based on recent social networking concepts and individual merit and exhibits convenient asynchronous, autodocumenting and autoregulating characteristics.
-Team
-members take on an egalitarian and volutary role, by logging into
-periodic ``AA sessions'' for an arbitrary duration (e.g.\ 2 hours per day).
-During each session, a user creates a log composed of short text sentences
-about their activity. These logs are
-publicly aggregated in a Website and are peer-validated, as in code
-review. A short screencast is ideally recorded at the end of each session to make
-AA logs more understandable. This methodology seems to be well-suited for
-increasing the efficiency of teams working on
-Global Software Development (GSD). Recognized reasons for this are
-the: 1) builtin asynchronous
-on-demand communication, documentation and working hours comprobatives;
-2) reduced need for central management,
-meetings and time-consuming reports. Hence, AA
-legitimizes and eases activities of a distributed software team.  It
-enables groups to have new means to fund activities,
-allowing for business models to emerge from
-distributed software development. AA is proposed,
-at it's core, as a way of having self-replicating hacker initiatives. These claims are
-discussed in a real case-study of a distributed free software
-hacker team called Lab Macambira.
+We present a new methodology for coordinating teamwork called Algorithmic
+Auto-regulation (AA). This methodology is based on recent social networking
+concepts and individual merit and exhibits convenient asynchronous,
+auto-documenting and auto-regulating characteristics. Team members take on an
+egalitarian and voluntary role, by logging into periodic ``AA sessions'' for an
+arbitrary duration (e.g.\ 2 hours per day).  During each session, a user creates
+a log composed of short text sentences about their activity. These logs are
+publicly aggregated in a Website and are peer-validated, as in code review. A
+short screen-cast is ideally recorded at the end of each session to make AA logs
+more understandable. This methodology seems to be well-suited for increasing the
+efficiency of teams working on Global Software Development (GSD). Recognized
+reasons for this are the: 1) built-in asynchronous on-demand communication,
+documentation and proven working hours; 2) reduced need for central management,
+meetings and time-consuming reports. Hence, AA legitimizes and eases activities
+of a distributed software team.  It enables groups to have new means to fund
+activities, allowing for business models to emerge from distributed software
+development. AA is proposed, at it's core, as a way of having self-replicating
+hacker initiatives. These claims are discussed in a real case-study of a
+distributed free software hacker team called Lab Macambira.
 \end{abstract}

### 2013-05-31 paper.tex
spell checked all paper and frontmatter

@@
 
-One of the defining features of modern times is the widening
-geographical distribution of software teams~\cite{last2003}.
-This is responsible for the so called Global Software Development (GSD)~\cite{german2003}. 
-An example is the free software
-movement. Projects and institutions like Mozilla Foundation has several
-employees and thousands of voluntary developers distributed across
-many countries. The same is true for GNOME~\cite{german2003}, OpenBSD,
-MySQL or Apache Software Foundation, to cite just a few\footnote{Ohloh, the open source network, have a more complete
-  and constantly updated list of the most active projects on-line at
-  \url{www.ohloh.net}}. Along the free and open software
-projects, GSD has a growing popularity in every niche of the software
-industry as a whole, even on those distributing their software with
-proprietary licenses. This phenomenon is attributed to a variety of factors
-such as a larger labor pool, natural globalization of software companies
-and foundations or even the premise of cheaper cost of
-production~\cite{komi2005}.
+One of the defining features of modern times is the widening geographical
+distribution of software teams~\cite{last2003}. This is responsible for the so
+called Global Software Development (GSD)~\cite{german2003}.  An example is the
+free software movement. Projects and institutions like Mozilla Foundation has
+several employees and thousands of voluntary developers distributed across many
+countries. The same is true for GNOME~\cite{german2003}, OpenBSD, MySQL or
+Apache Software Foundation, to cite just a few\footnote{Ohloh, the open source
+  network, have a more complete and constantly updated list of the most active
+  projects on-line at \url{www.ohloh.net}}. Along the free and open software
+projects, GSD has a growing popularity in every niche of the software industry
+as a whole, even on those distributing their software with proprietary
+licenses. This phenomenon is attributed to a variety of factors such as a larger
+labor pool, natural globalization of software companies and foundations or even
+the premise of cheaper cost of production~\cite{komi2005}.
 
-Dispite the advantages of GSD, it is known how
-difficult it is to coordinate and fund free software on a larger scale
-than currently available. That is, when teams are very heterogeneous, containing
-not only volunteers and experienced developers, but also
-contractors from different backgrounds and cultures. This observations
-are in consonance with dedicated studies, such as~\cite{carmel1999}.
-As a common base, related difficulties for GSD gather around distance, time and cultural differences. In
-the case of free or open software projects, all these factors are
-involved.
+Despite the advantages of GSD, it is known how difficult it is to coordinate and
+fund free software on a larger scale than currently available. That is, when
+teams are very heterogeneous, containing not only volunteers and experienced
+developers, but also contractors from different backgrounds and cultures. This
+observations are in consonance with dedicated studies, such
+as~\cite{carmel1999}.  As a common base, related difficulties for GSD gather
+around distance, time and cultural differences. In the case of free or open
+software projects, all these factors are involved.
 
-Another problem faced by modern software companies and other
-collectives are frequent ineffective meetings, which are seldom
-focused on the interest of any attendant. The result is that it has
-become the norm to participate in these with the ``laptop
-open'', which is unproductive. Software developers
-like to code, to be productive, to have their hands on their project,
-to do what they are best at. They dislike to have to stop for meetings
-or to write lengthy reports to justify their funding.
+Another problem faced by modern software companies and other collectives are
+frequent ineffective meetings, which are seldom focused on the interest of any
+attendant. The result is that it has become the norm to participate in these
+with the ``laptop open'', which is unproductive. Software developers like to
+code, to be productive, to have their hands on their project, to do what they
+are best at. They dislike to have to stop for meetings or to write lengthy
+reports to justify their funding.
 %\todo{ler mais. cacm, etc}
 
-To address these matters is the purpose of AA, a methodology and an associated software system
-for coordinating distributed team work dealing with the disadvantages
-of GSD. Team members take on an egalitarian role, and stay
-voluntarily \textit{logged} in the system for part of their time
-(e.g.\ 2 hours per day), during which they log a periodical short text
-sentence --- similar to a ``tweet'' in this Twitter era --- as the
-status of their activity using an easy to use command. These
-``microblog sentences'' are publicly aggregated and validated by other
-team members. Through AA, the community has a methodology and an associated
-system to validate and enable the activities of a distributed software
-team. It implicitly legitimizes financial support for the expansion of
-the activity of the developer team as the activities are documented and validated.
-The AA methodology is specially
-useful for coordinating distributed and decentralized team work,
-providing effective means to asynchronously update different team
-members, diminishing the need for harsh synchronous meetings.
+To address these matters is the purpose of AA, a methodology and an associated
+software system for coordinating distributed team work dealing with the
+disadvantages of GSD. Team members take on an egalitarian role, and stay
+voluntarily \textit{logged} in the system for part of their time (e.g.\ 2 hours
+per day), during which they log a periodical short text sentence --- similar to
+a ``tweet'' in this Twitter era --- as the status of their activity using an
+easy to use command. These ``micro-blog sentences'' are publicly aggregated and
+validated by other team members. Through AA, the community has a methodology and
+an associated system to validate and enable the activities of a distributed
+software team. It implicitly legitimizes financial support for the expansion of
+the activity of the developer team as the activities are documented and
+validated.  The AA methodology is specially useful for coordinating distributed
+and decentralized team work, providing effective means to asynchronously update
+different team members, diminishing the need for harsh synchronous meetings.
 
@@
 
-A brief overview of current papers about GSD methodologies related with
-AA is presented in section~\ref{related-work}. In
-section~\ref{aa-methodology}, the most relevant
-characteristics of the AA methodology are outlined. In this same section,
-there is a description of
-an AA use in a team of 9 developers in the second half of
-2011 and a broader a more released use of AA from 2012 until now. Finally, in Section~\ref{conclusions}, there are final
-conclusions and indicatives of future possibilities to the
-practical use of AA on other teams of software developers or
-individuals working on non-software activities.
+A brief overview of current papers about GSD methodologies related with AA is
+presented in section~\ref{related-work}. In section~\ref{aa-methodology}, the
+most relevant characteristics of the AA methodology are outlined. In this same
+section, there is a description of an AA use in a team of 9 developers in the
+second half of 2011 and a broader a more released use of AA from 2012 until
+now. Finally, in Section~\ref{conclusions}, there are final conclusions and
+indicatives of future possibilities to the practical use of AA on other teams of
+software developers or individuals working on non-software activities.
 
@@
 on the digital age enables people to be very productive, the article also
-mentions linux as a hallmark example~\cite{Thompson:Wired:2012}
+mentions Linux as a hallmark example~\cite{Thompson:Wired:2012}
 }
@@
 
-There has been a large amount of research done in the area of
-methodologies to deal with distributed teams of developers. We are
-focusing in GSD here, however some principles involved on those
-methodologies could be used on smaller teams of developers working in
-the same place, time and with minor cultural differences. Moreover we
-generally think on ``distributed development'' being global which is
-not totally true. We even applied AA to a team that work at the same
-city but on different times (more details on
-Section~\ref{results}). Even smaller groups of developers working on
-the same building could use GSD methodologies. A thorough survey of
-these methodologies is beyond the scope of this paper. Here we present
-a brief overview.
+There has been a large amount of research done in the area of methodologies to
+deal with distributed teams of developers. We are focusing in GSD here, however
+some principles involved on those methodologies could be used on smaller teams
+of developers working in the same place, time and with minor cultural
+differences. Moreover we generally think on ``distributed development'' being
+global which is not totally true. We even applied AA to a team that work at the
+same city but on different times (more details on Section~\ref{results}). Even
+smaller groups of developers working on the same building could use GSD
+methodologies. A thorough survey of these methodologies is beyond the scope of
+this paper. Here we present a brief overview.
 
-Various methodologies for GSD were built around the factors that
-affect distributed team works, proposed by Carmel~\cite{carmel1999}
-and comprising three distances: geographical, cultural and
-temporal. Geographical distance prejudice \emph{coordination}, being
-the act of integrating all the tasks distributed between
-units~\cite{carmel2001}; \emph{control}, or the process to maintain
-specific goals, policies or quality levels; and
-\emph{communication}. All these factors are correlated, for example, a
-team need to have clearly communication to work on tasks of a specific
-problem. Cultural distance encompass differences on organizational and
-natural culture. Spoken language, unit and ethnic values are common
-forms of this distance. Some companies prefer to situate development
-units in foreign locations with minimal cultural distance (e.g.\ an
-American firm prefer Ireland, because of spoken language
-similarity~\cite{carmel2001}). And finally the temporal distance that
-prejudices synchronous communications like telephone or
-videoconferences. Units of developers working on different time-zones
-are concerned with managing of their agenda guided by this temporal
-distance.
+Various methodologies for GSD were built around the factors that affect
+distributed team works, proposed by Carmel~\cite{carmel1999} and comprising
+three distances: geographical, cultural and temporal. Geographical distance
+prejudice \emph{coordination}, being the act of integrating all the tasks
+distributed between units~\cite{carmel2001}; \emph{control}, or the process to
+maintain specific goals, policies or quality levels; and
+\emph{communication}. All these factors are correlated, for example, a team need
+to have clearly communication to work on tasks of a specific problem. Cultural
+distance encompass differences on organizational and natural culture. Spoken
+language, unit and ethnic values are common forms of this distance. Some
+companies prefer to situate development units in foreign locations with minimal
+cultural distance (e.g.\ an American firm prefer Ireland, because of spoken
+language similarity~\cite{carmel2001}). And finally the temporal distance that
+prejudices synchronous communications like telephone or video-conferences. Units
+of developers working on different time-zones are concerned with managing of
+their agenda guided by this temporal distance.
 
-Targeting geographical distance, Carmel~\cite{carmel2001} suggests a
-tactic to reduce intensive collaboration. His approach divides the
-whole software life-cycle on levels of complexity. Each level has a
-degree of collaboration. For example, some developers working on a
-project with high collaboration level should use the follow-the-sun
-approach: when concluding the work day, they pass their work to the
-team working on another time-zone. Other tactics are suggested by the
-same author to deal with the three distances, like separating foreign
-units of developers in time-zone bands.
+Targeting geographical distance, Carmel~\cite{carmel2001} suggests a tactic to
+reduce intensive collaboration. His approach divides the whole software
+life-cycle on levels of complexity. Each level has a degree of
+collaboration. For example, some developers working on a project with high
+collaboration level should use the follow-the-sun approach: when concluding the
+work day, they pass their work to the team working on another time-zone. Other
+tactics are suggested by the same author to deal with the three distances, like
+separating foreign units of developers in time-zone bands.
 
-Battin et al.~\cite{battin2001} propose and argues about their
-experiments using specific methodologies created for the distributed
-development centers of the Motorola Company (which has over 25
-software development centers worldwide). These methodologies included
-constant communication with critical units, incremental integration
-and scheduled based on time-zones distributed to developers on 6
-countries from 3 continents.
+Battin et al.~\cite{battin2001} propose and argues about their experiments using
+specific methodologies created for the distributed development centers of the
+Motorola Company (which has over 25 software development centers
+worldwide). These methodologies included constant communication with critical
+units, incremental integration and scheduled based on time-zones distributed to
+developers on 6 countries from 3 continents.
 
-While considering free software projects instead of companies, the
-same factors are present and some methodologies
-arises. German~\cite{german2003} gives a concise review of
-methodologies used by the GNOME project, one of the most active free
-software projects and used by companies like Sun Microsystems. It is
-interesting to note a difference on viewpoint present on his paper:
-German focuses the methodology description on code. He start explaining
-that GNOME is separated into modules (76 on version 2.4, to be
-precise) and each module has one maintainer who divide his modules
-into separated parts in which other developers can work on independent
-tasks, along other responsibilities. As like others free software
-projects, all the development was done using a \emph{bugtracker} to
-bugs and issues management, mail lists and Internet Relay Chat (IRC)
-to discussion and communication and a software configuration
-management like SVN or Git. Periodical (commonly yearly) conferences
-like GUADEC is common on free and open source projects to face-to-face
-meeting and is based each time on a different place.
+While considering free software projects instead of companies, the same factors
+are present and some methodologies arises. German~\cite{german2003} gives a
+concise review of methodologies used by the GNOME project, one of the most
+active free software projects and used by companies like Sun Microsystems. It is
+interesting to note a difference on viewpoint present on his paper: German
+focuses the methodology description on code. He start explaining that GNOME is
+separated into modules (76 on version 2.4, to be precise) and each module has
+one maintainer who divide his modules into separated parts in which other
+developers can work on independent tasks, along other responsibilities. As like
+others free software projects, all the development was done using a
+\emph{bugtracker} to bugs and issues management, mail lists and Internet Relay
+Chat (IRC) to discussion and communication and a software configuration
+management like SVN or Git. Periodical (commonly yearly) conferences like GUADEC
+is common on free and open source projects to face-to-face meeting and is based
+each time on a different place.
 
@@
 
-As noted, some strategies for GSD is based on complex methodologies
-and many of those were built for a specific company or software
-center. Here we propose an alternative methodology based on a simple
-idea: small working sessions logged by a computational
-tool. Figure~\ref{fig:mm} summarizes our methodology.
+As noted, some strategies for GSD is based on complex methodologies and many of
+those were built for a specific company or software center. Here we propose an
+alternative methodology based on a simple idea: small working sessions logged by
+a computational tool. Figure~\ref{fig:mm} summarizes our methodology.
 
@@
 
-From the developer perspective, the AA methodology is based on
-creating pretty small high perspective reports of what they are doing
-in a specific time frame, that can be something between 5 to 15
-minutes, depending in what is more comfortable for the developer. An
-AA Session would be a period of at least 2 hours doing these
-reports. The developer can set reminders to show up when its time to
-make a new report. The objective of the flexible time frame and
-reminders is to minimize developer overhead during his AA session. In
-this way he can make the reports while staying concentrated in his
-code. Each report can be sent directly to an on-line server, or stored
-locally in a temporarily database for sending later. This make
-possible the use of AA while offline.
+From the developer perspective, the AA methodology is based on creating pretty
+small high perspective reports of what they are doing in a specific time frame,
+that can be something between 5 to 15 minutes, depending in what is more
+comfortable for the developer. An AA Session would be a period of at least 2
+hours doing these reports. The developer can set reminders to show up when its
+time to make a new report. The objective of the flexible time frame and
+reminders is to minimize developer overhead during his AA session. In this way
+he can make the reports while staying concentrated in his code. Each report can
+be sent directly to an on-line server, or stored locally in a temporarily
+database for sending later. This make possible the use of AA while offline.
 
-Developers can also record a video screencast in the end of the
-session summarizing what has been done, explaining with his words and
-showing his most important results. This, combined with the textual
-log of his session, makes the whole report more understandable for
-himself or other people searching for information about his
-production.
+Developers can also record a video screencast in the end of the session
+summarizing what has been done, explaining with his words and showing his most
+important results. This, combined with the textual log of his session, makes the
+whole report more understandable for himself or other people searching for
+information about his production.
 
@@
 
-All AA reports made by the developers are sent to a Web server and
-become public in a Website, being possible then for a manager or
-another developer to follow very closely the work of a developer,
-nearly real-time, reading each of his small reports of what he is
-doing right now.
+All AA reports made by the developers are sent to a Web server and become public
+in a Website, being possible then for a manager or another developer to follow
+very closely the work of a developer, nearly real-time, reading each of his
+small reports of what he is doing right now.
 
-Another possibility is to check older sessions to see when sometimes
-was done and the comments of the developer about it, since each AA
-post happens in a very small time frame of work, the information about
-what was done become very easy to understand, instead of a long
-report in the end of a session.
+Another possibility is to check older sessions to see when sometimes was done
+and the comments of the developer about it, since each AA post happens in a very
+small time frame of work, the information about what was done become very easy
+to understand, instead of a long report in the end of a session.
 
-The site is where the developer can add to his session a link for his
-screencast about that session, a video summary of what was done to
-complement the reports, useful specially on cases where the small
-reports were done in a hurry, because the developer did not want to
-lose his focus on something important at that moment.
+The site is where the developer can add to his session a link for his screencast
+about that session, a video summary of what was done to complement the reports,
+useful specially on cases where the small reports were done in a hurry, because
+the developer did not want to lose his focus on something important at that
+moment.
 
@@
 
-Each AA session must be validated by another developer, it means that
-all reports are read by someone that will consider then valid or not
-and will even write commentaries about the specific session. The
-developer in charge of validating a session is decided randomly by the
-AA Web server, which send an email to the chosen developer with an
-URL to a validation interface.
+Each AA session must be validated by another developer, it means that all
+reports are read by someone that will consider then valid or not and will even
+write commentaries about the specific session. The developer in charge of
+validating a session is decided randomly by the AA Web server, which send an
+email to the chosen developer with an URL to a validation interface.
 
@@
 
-The easy and effective management of teams working on GSD is the main
-purpose of the AA methodology. We applied this methodology to a group
-of 9 developers in July of 2011 on what we called Lab
-Macambira~\footnote{LabMacambira.sf.net:
-  \url{http://labmacambira.sf.net}}. The main objective of the team
-was to work on different free software projects, contributing directly
-to their development, sending bug corrections or proposing new
-features on their source code.
+The easy and effective management of teams working on GSD is the main purpose of
+the AA methodology. We applied this methodology to a group of 9 developers in
+July of 2011 on what we called Lab Macambira~\footnote{LabMacambira.sf.net:
+  \url{http://labmacambira.sf.net}}. The main objective of the team was to work
+on different free software projects, contributing directly to their development,
+sending bug corrections or proposing new features on their source code.
 
-All the team members had different levels of knowledge on software
-development as part of large and distributed free software projects
-like Scilab or Mozilla. In this way, one month of training was
-conducted by two experienced developers, teaching the basics of use of
-development support tools like bugtrackers, programming languages and
-version control systems. After this period, a challenge was proposed
-for the new developers: send a bug correction or a new feature to a
-large free software project and you will work with us, you will be a
-``Macambira'' developer. Table~\ref{tabela:contribuicoes} summarizes
+All the team members had different levels of knowledge on software development
+as part of large and distributed free software projects like Scilab or
+Mozilla. In this way, one month of training was conducted by two experienced
+developers, teaching the basics of use of development support tools like
+bugtrackers, programming languages and version control systems. After this
+period, a challenge was proposed for the new developers: send a bug correction
+or a new feature to a large free software project and you will work with us, you
+will be a ``Macambira'' developer. Table~\ref{tabela:contribuicoes} summarizes
 the contributions of each ``Macambira'' to free software projects.
@@
     common practice to use pseudonym as identification.}
-    \begin{tabular}{|l|l|}
+    \small\begin{tabular}{|l|l|}
         \hline
@@
 In one month, each developer contributed to many very large free software
-projects. Many of the developers started the training with no
-knowledge of what is free software and ended that period becoming a
-free software developer.
+projects. Many of the developers started the training with no knowledge of what
+is free software and ended that period becoming a free software developer.
 
-During that month, the same team developed the first version of the AA
-system and used AA to manage their activities. Even while developing
-the system. All the source code of AA --- the client that sends the
-logs and the AA Web server --- is public available~\footnote{AA source
-  code:
+During that month, the same team developed the first version of the AA system
+and used AA to manage their activities. Even while developing the system. All
+the source code of AA --- the client that sends the logs and the AA Web server
+--- is public available~\footnote{AA source code:
   \url{http://labmacambira.git.sourceforge.net/git/gitweb.cgi?p=labmacambira/aa}}
-and all the AA sessions log of the whole team of ``Lab Macambira'' is
-also on-line~\footnote{Logs of AA sessions:
+and all the AA sessions log of the whole team of ``Lab Macambira'' is also
+on-line~\footnote{Logs of AA sessions:
   \url{http://labmacambira.git.sourceforge.net/git/gitweb.cgi?p=labmacambira/paainel}}.
 
-After the training period, during more 6 months, the ``Macambiras''
-worked on a large range of free software projects, distributed on work
-groups --- each work group focusing on a specific theme like video,
-audio and web --- and financed by contracts and support of the
-``Pont\~{a}o N\'{o}s Digitais''. In Table~\ref{tabela:criados} we can
-see a list of the free software created by ``Lab Macambira'' since
-July of 2011.
+After the training period, during more 6 months, the ``Macambiras'' worked on a
+large range of free software projects, distributed on work groups --- each work
+group focusing on a specific theme like video, audio and web --- and financed by
+contracts and support of the ``Pont\~{a}o N\'{o}s Digitais''. In
+Table~\ref{tabela:criados} we can see a list of the free software created by
+``Lab Macambira'' since July of 2011.
 
@@
 
-As of this writing the ``Lab Macambira'' have many software
-developers, and some of the trained developers continue to work
-voluntarily in the project.
+As of this writing the ``Lab Macambira'' have many software developers, and some
+of the trained developers continue to work voluntarily in the project.
 
@@
 %% brief introduction
-In a scenario where GSD is growing as a popular form of software
-development, not just on free software projects but in the whole
-software industry, we need methodologies to deal with its
-disadvantages and at the same time to amplify its advantages.
+In a scenario where GSD is growing as a popular form of software development,
+not just on free software projects but in the whole software industry, we need
+methodologies to deal with its disadvantages and at the same time to amplify its
+advantages.
 
-This paper has presented a methodology to GSD, being the development
-conducted on large or small groups of software developers, working on
-different countries or even at the same room. The AA methodology
-implements a simple system where each developer take notes of his work
-generating a periodical log of small text sentences. The sum of those
-sentences, along an entire session of work, results in a complete
-report. The report is made public available through a Website and be
-validated by other developers sorted aleatory by the AA Web server.
+This paper has presented a methodology to GSD, being the development conducted
+on large or small groups of software developers, working on different countries
+or even at the same room. The AA methodology implements a simple system where
+each developer take notes of his work generating a periodical log of small text
+sentences. The sum of those sentences, along an entire session of work, results
+in a complete report. The report is made public available through a Website and
+be validated by other developers sorted aleatory by the AA Web server.
 
-Instead of a merely work-management tool, AA act as a methodology to
-improve the time sense of individuals, dividing their work on small
-sessions, and also reducing the need of extensive
-reports or unnecessary meetings. By asking users to write a minimal
-text sentence as a continuously log, AA does not disturb developers
-concentrated on programming: developers just have to type some
+Instead of a merely work-management tool, AA act as a methodology to improve the
+time sense of individuals, dividing their work on small sessions, and also
+reducing the need of extensive reports or unnecessary meetings. By asking users
+to write a minimal text sentence as a continuously log, AA does not disturb
+developers concentrated on programming: developers just have to type some
 characters, hit \textit{enter} and go back to coding.
 
-AA application is not restricted to software development. As of this
-writing there is a comic book studio~\footnote{Pula pirata:
-  \url{http://pulapirata.com}} starting to use AA to manage their
-activities. People with non-software background, like social
-scientists, musicians and activists has also using AA and contributing
-for its improvement.
+AA application is not restricted to software development. As of this writing
+there is a comic book studio~\footnote{Pula pirata: \url{http://pulapirata.com}}
+starting to use AA to manage their activities. People with non-software
+background, like social scientists, musicians and activists has also using AA
+and contributing for its improvement.
 
-For developer teams, we have experienced the use of AA to
-auto-regulate the work of ``Lab Macambira'', a group of free software
-developers from Brazil. Since July of 2011 the group have contributed
-and created new free and open source software for a vast number of
-applications.
+For developer teams, we have experienced the use of AA to auto-regulate the work
+of ``Lab Macambira'', a group of free software developers from Brazil. Since
+July of 2011 the group have contributed and created new free and open source
+software for a vast number of applications.
 
-There are many aspects of the work which remain unfinished. New ways
-to report logs --- the ``Twitter like'' messages --- from different
-interfaces like IRC, Internet Messaging services and email can make
-the use of AA easy and widespread, turning AA an ubiquitous system,
-presented on everyday communication channels. Even the work logs
-generated since July of 2011 could be vastly statistically analyzed
-aiming to recognize patterns in the behavior of individuals and their
-productions.
+There are many aspects of the work which remain unfinished. New ways to report
+logs --- the ``Twitter like'' messages --- from different interfaces like IRC,
+Internet Messaging services and email can make the use of AA easy and
+widespread, turning AA an ubiquitous system, presented on everyday communication
+channels. Even the work logs generated since July of 2011 could be vastly
+statistically analyzed aiming to recognize patterns in the behavior of
+individuals and their productions.
 
-We would like to conclude setting an important role of AA: being a
-free software system and an open methodology, AA could be used to
-auto-manage groups of individuals working on software or other kinds
-of activities. In this way, we are interested to spread AA for those
-groups, to have even more developers contributing in a collaborative
-way.
+We would like to conclude setting an important role of AA: being a free software
+system and an open methodology, AA could be used to auto-manage groups of
+individuals working on software or other kinds of activities. In this way, we
+are interested to spread AA for those groups, to have even more developers
+contributing in a collaborative way.
 
@@
 
-We would like to also thank AA: the present research and even this
-manuscript was written using AA. The complete log is on-line at
+We would like to also thank AA: the present research and even this manuscript
+was written using AA. The complete log is on-line at
 \url{http://www.pulapirata.com/skills/aa}.

### 2013-06-01 paper.tex
fixed tables size
revisado artigo todo na escrita

@@
 
-\indraftnote{
-A very good article on the value of asynchronous communication for personal
-and group productivity, related to the key necessity of having moments of
-introversion to avoid daily pressures of forced socialization. The way we work
-on the digital age enables people to be very productive, the article also
-mentions Linux as a hallmark example~\cite{Thompson:Wired:2012}
-}
-
-\indraftnote{TODO: cite CIA.vc bot stuff}
-
-\section{Related Work}
+\section{Related work}
 \label{related-work}
@@
 
-There has been a large amount of research done in the area of methodologies to
-deal with distributed teams of developers. We are focusing in GSD here, however
+There has been a large amount of research in methodologies to
+deal with distributed teams of developers. Although this article focuses in GSD,
 some principles involved on those methodologies could be used on smaller teams
-of developers working in the same place, time and with minor cultural
-differences. Moreover we generally think on ``distributed development'' being
-global which is not totally true. We even applied AA to a team that work at the
+of developers working in the same place, time period and with minor cultural
+differences. Moreover, ``distributed development'' is generally tough as being
+global which is not true. AA was applyed to a team that work at the
 same city but on different times (more details on Section~\ref{results}). Even
@@
 methodologies. A thorough survey of these methodologies is beyond the scope of
-this paper. Here we present a brief overview.
+this paper. This section presents a brief overview.
 
 Various methodologies for GSD were built around the factors that affect
-distributed team works, proposed by Carmel~\cite{carmel1999} and comprising
+distributed team works. As proposed by Carmel~\cite{carmel1999}, these comprises
 three distances: geographical, cultural and temporal. Geographical distance
@@
 maintain specific goals, policies or quality levels; and
-\emph{communication}. All these factors are correlated, for example, a team need
-to have clearly communication to work on tasks of a specific problem. Cultural
+\emph{communication}. All these factors are correlated, for example, a team needs
+to have good communication to work on tasks related to a specific problem. Cultural
 distance encompass differences on organizational and natural culture. Spoken
@@
 cultural distance (e.g.\ an American firm prefer Ireland, because of spoken
-language similarity~\cite{carmel2001}). And finally the temporal distance that
-prejudices synchronous communications like telephone or video-conferences. Units
-of developers working on different time-zones are concerned with managing of
+language similarity~\cite{carmel2001}). Finally, the temporal distance
+prejudices synchronous communications like telephone or video-conferences.
+Developers working on different time-zones manage
 their agenda guided by this temporal distance.
@@
 Targeting geographical distance, Carmel~\cite{carmel2001} suggests a tactic to
-reduce intensive collaboration. His approach divides the whole software
+enhance intensive collaboration. His approach divides the whole software
 life-cycle on levels of complexity. Each level has a degree of
@@
 
-Battin et al.~\cite{battin2001} propose and argues about their experiments using
-specific methodologies created for the distributed development centers of the
+Battin et al.~\cite{battin2001} argues about their experiments using
+methodologies created for the development centers of the
 Motorola Company (which has over 25 software development centers
 worldwide). These methodologies included constant communication with critical
-units, incremental integration and scheduled based on time-zones distributed to
+units, incremental integration and schedules based on time-zones distributed to
 developers on 6 countries from 3 continents.
 
-While considering free software projects instead of companies, the same factors
+While considering free software projects instead of companies, similar factors
 are present and some methodologies arises. German~\cite{german2003} gives a
@@
 active free software projects and used by companies like Sun Microsystems. It is
-interesting to note a difference on viewpoint present on his paper: German
+interesting to note a difference on the viewpoint present on his paper: German
 focuses the methodology description on code. He start explaining that GNOME is
@@
 \emph{bugtracker} to bugs and issues management, mail lists and Internet Relay
-Chat (IRC) to discussion and communication and a software configuration
-management like SVN or Git. Periodical (commonly yearly) conferences like GUADEC
-is common on free and open source projects to face-to-face meeting and is based
+Chat (IRC) to discussion and communication and a software version
+management like SVN or Git. He finishes by observing that periodical, e.g. yearly, conferences like GUADEC
+is common on free and open source projects for face-to-face meetings and is based
 each time on a different place.
 
-\section{The AA Methodology}
+\section{The AA methodology}
 \label{aa-methodology}
@@
 As noted, some strategies for GSD is based on complex methodologies and many of
-those were built for a specific company or software center. Here we propose an
-alternative methodology based on a simple idea: small working sessions logged by
-a computational tool. Figure~\ref{fig:mm} summarizes our methodology.
+those were built for a specific company or software center. This section proposes an
+alternative based on a simple idea: small working sessions logged by
+a computational tool. Figure~\ref{fig:mm} summarizes this methodology.
 
@@
 \begin{center}
-   \includegraphics[width=0.8\linewidth,keepaspectratio=true]{figs/aa-mm.png}
+   \includegraphics[width=0.8\linewidth,keepaspectratio=true]{figs/aa-mm2.png}
 \end{center}
    \caption{
-   Mindmap of our methodology
+   Mindmap of AA methodology: 1) developer engagement cycle, i.e. his use of AA; 2) functionality as main goals of the system; 3) potentialities as enhancements AA delivers to the developer's medium. 
    }
@@
 
-\subsection{AA Session}
+\subsection{The AA session}
 
-From the developer perspective, the AA methodology is based on creating pretty
-small high perspective reports of what they are doing in a specific time frame,
-that can be something between 5 to 15 minutes, depending in what is more
-comfortable for the developer. An AA Session would be a period of at least 2
-hours doing these reports. The developer can set reminders to show up when its
+From a developer's perspective, the AA methodology is based on creating
+small, high perspective, reports of what is being done in a specific time frame. This frame
+can be something between 5 to 15 minutes, depending on what is 
+comfortable for the developer. An AA Session is a period of about 2
+hours, in which these short reports (called ``shouts'') are periodically written. The developer can set reminders to show up when it is
 time to make a new report. The objective of the flexible time frame and
-reminders is to minimize developer overhead during his AA session. In this way
-he can make the reports while staying concentrated in his code. Each report can
-be sent directly to an on-line server, or stored locally in a temporarily
-database for sending later. This make possible the use of AA while offline.
+reminders is to minimize developer overhead during his AA session. In this way,
+he can make the reports with little or no loss of concentration on his coding activity. Each report can
+be sent directly to an online server, or stored locally in a temporary
+database for sending later on. This makes it possible to use AA while offline.
 
-Developers can also record a video screencast in the end of the session
-summarizing what has been done, explaining with his words and showing his most
+Developers can also record a video screencast at the end of the session,
+summarizing what has been done, explaining with his words and showing on the screen his most
 important results. This, combined with the textual log of his session, makes the
-whole report more understandable for himself or other people searching for
+whole report more significant as documentation and understandable for himself or other people searching for
 information about his production.
 
-\subsection{AA Website Report}
+\subsection{The AA website report}
 
@@
 \begin{center}
-   \includegraphics[width=0.95\linewidth]{figs/aa-0_1.png}
+   \includegraphics[width=0.95\linewidth]{figs/aa-0_1_.png}
 \end{center}
-   \caption{AA Version 0.1}
+   \caption{AA Version 0.1.
+Messages of users: hybrid, filter0, v1z e aut0mata about activities of interest for labMacambira.sf.net and other parties (IPRJ/UERJ, IFT/UNESP, IFSC/USP, OPW/Mozilla, PulaPirataComics).
+Each message is a ``shout'', which, grouped, forms an AA session.}
 \label{fig:aaserver}
@@
 All AA reports made by the developers are sent to a Web server and become public
-in a Website, being possible then for a manager or another developer to follow
-very closely the work of a developer, nearly real-time, reading each of his
-small reports of what he is doing right now.
+in a Website. It is then possible for a manager or another developer to follow
+closely the work of a developer, nearly real-time, reading each of his
+small reports of what he is doing.
 
-Another possibility is to check older sessions to see when sometimes was done
-and the comments of the developer about it, since each AA post happens in a very
-small time frame of work, the information about what was done become very easy
-to understand, instead of a long report in the end of a session.
+Another possibility is to check older sessions to see what was done
+and the comments of the developer about it. Since each AA post (or ``shout'') happens in a
+small time frame of work, the information about what was done becomes easy
+to understand, which seems better than a long report in the end of a session.
 
-The site is where the developer can add to his session a link for his screencast
-about that session, a video summary of what was done to complement the reports,
-useful specially on cases where the small reports were done in a hurry, because
-the developer did not want to lose his focus on something important at that
-moment.
+The developer can add to his session a link for his screencast
+about that session. This screencast is a video summary of what was done and complements the report.
 
 \subsection{Validation}
-
-Each AA session must be validated by another developer, it means that all
-reports are read by someone that will consider then valid or not and will even
-write commentaries about the specific session. The developer in charge of
-validating a session is decided randomly by the AA Web server, which send an
-email to the chosen developer with an URL to a validation interface.
+Each AA session must be validated by another developer. This means that all
+reports are read by someone that will consider valid or not and will even
+write comments about the specific session. The developer in charge of
+validating a session is decided randomly by the AA Web server, which sends an
+email to the chosen developer with the URL to a validation interface.
 
@@
 % TODO: adicionar figura com arquitetura do AA cliente + AA web + validação
-% TODO: melhorar screenshot do AA web
+% TODO: adicionar figura do pAAinel
 
-\section{Results and Discussion}
+\section{Results and discussion}
 \label{results}
@@
 The easy and effective management of teams working on GSD is the main purpose of
-the AA methodology. We applied this methodology to a group of 9 developers in
-July of 2011 on what we called Lab Macambira~\footnote{LabMacambira.sf.net:
+the AA methodology. This methodology was applied to a group of 9 developers in the second half of 2011 on what was named ``Lab Macambira''\footnote{LabMacambira.sf.net:
   \url{http://labmacambira.sf.net}}. The main objective of the team was to work
@@
 
-All the team members had different levels of knowledge on software development
-as part of large and distributed free software projects like Scilab or
-Mozilla. In this way, one month of training was conducted by two experienced
-developers, teaching the basics of use of development support tools like
+Team members earned knowledge on software development
+as part of large and distributed free software projects like Scilab and  
+Mozilla. In this way, one month of training was conducted by three experienced
+developers, teaching the basics of development support tools like
 bugtrackers, programming languages and version control systems. After this
@@
   \caption{Free and open software projects that received contributions
-    from ``Macambiras''. On the first column we can see a list of
-    applications of those projects. At right, the pseudonym of the
+    from ``Macambiras''. On the first column there is a list of
+    applications. At right, the pseudonym of the
     ``Macambiras'' who sent \textit{commits} to the application. At
     ``Lab Macambira'', and at free software community in general, is a
-    common practice to use pseudonym as identification.}
+    common practice to use a pseudonym as identification.}
+    \begin{center}
     \small\begin{tabular}{|l|l|}
@@
 \end{tabular}
+\end{center}
     \label{tabela:contribuicoes}
@@
 
-In one month, each developer contributed to many very large free software
-projects. Many of the developers started the training with no knowledge of what
+In one month, ``Macambira'' developers contributed to many large free software
+projects. Some of the developers started the training with no knowledge of what
 is free software and ended that period becoming a free software developer.
@@
 During that month, the same team developed the first version of the AA system
-and used AA to manage their activities. Even while developing the system. All
+and used AA to manage their activities, even while developing the system. All
 the source code of AA --- the client that sends the logs and the AA Web server
---- is public available~\footnote{AA source code:
+--- is publicly available\footnote{AA source code:
   \url{http://labmacambira.git.sourceforge.net/git/gitweb.cgi?p=labmacambira/aa}}
-and all the AA sessions log of the whole team of ``Lab Macambira'' is also
-on-line~\footnote{Logs of AA sessions:
+and all the AA session logs of the ``Lab Macambira'' team is also
+online\footnote{Logs of AA sessions:
   \url{http://labmacambira.git.sourceforge.net/git/gitweb.cgi?p=labmacambira/paainel}}.
 
-After the training period, during more 6 months, the ``Macambiras'' worked on a
-large range of free software projects, distributed on work groups --- each work
-group focusing on a specific theme like video, audio and web --- and financed by
-contracts and support of the ``Pont\~{a}o N\'{o}s Digitais''. In
+After the training period, during 6 more months, the ``Macambiras'' worked on a
+large range of free software projects, distributed on work groups. Each work
+group focused on a specific theme like video, audio and web. They were financed by
+contracts and support of the ``Pont\~{a}o de Cultura Digital: N\'{o}s Digitais''. In
 Table~\ref{tabela:criados} we can see a list of the free software created by
@@
 \begin{table}
-    \caption{Software projects created by ``Lab Macambira'' since July
+    \caption{Software projects, created by ``Lab Macambira'', with the use of AA, since July
       of 2011 with a short description and the technologies --- like
@@
       of application.}
-    \begin{tabular}{|l|p{5cm}|l|}
+    \small\begin{tabular}{|l|p{5cm}|l|}
         \hline
@@
 In a scenario where GSD is growing as a popular form of software development,
-not just on free software projects but in the whole software industry, we need
-methodologies to deal with its disadvantages and at the same time to amplify its
+not just on free software projects but in the whole software industry, there is an increased need for
+methodologies to deal with it's disadvantages and at the same time to amplify it's
 advantages.
 
-This paper has presented a methodology to GSD, being the development conducted
+This paper has presented a simple methodology to GSD, being the development conducted
 on large or small groups of software developers, working on different countries
 or even at the same room. The AA methodology implements a simple system where
-each developer take notes of his work generating a periodical log of small text
+each developer takes notes of his work, generating a periodical log of small text
 sentences. The sum of those sentences, along an entire session of work, results
 in a complete report. The report is made public available through a Website and
-be validated by other developers sorted aleatory by the AA Web server.
+are validated by other developers sorted aleatory by the AA Web server.
 
-Instead of a merely work-management tool, AA act as a methodology to improve the
+Instead of a merely work-management tool, AA acts as a methodology to improve the
 time sense of individuals, dividing their work on small sessions, and also
 reducing the need of extensive reports or unnecessary meetings. By asking users
-to write a minimal text sentence as a continuously log, AA does not disturb
-developers concentrated on programming: developers just have to type some
-characters, hit \textit{enter} and go back to coding.
+to write a minimal text sentence as a continuous log, AA minimizes
+disturbances: developers just type some
+characters, hit \textit{enter} and go back to coding, which can help mental organization about the work being done.
 
 AA application is not restricted to software development. As of this writing
-there is a comic book studio~\footnote{Pula pirata: \url{http://pulapirata.com}}
-starting to use AA to manage their activities. People with non-software
-background, like social scientists, musicians and activists has also using AA
+there is a comic book studio\footnote{Pula Pirata Comics: \url{http://pulapirata.com}}
+using AA to manage their activities. People with non-software
+background, like social scientists, musicians and activists has also been using AA
 and contributing for its improvement.
 
-For developer teams, we have experienced the use of AA to auto-regulate the work
+For software developer teams, the use of AA was used successfully to auto-regulate the work
 of ``Lab Macambira'', a group of free software developers from Brazil. Since
@@
 There are many aspects of the work which remain unfinished. New ways to report
-logs --- the ``Twitter like'' messages --- from different interfaces like IRC,
+logs --- the ``Twitter like'' messages, sometimes called ``shouts'' --- from different interfaces like IRC,
 Internet Messaging services and email can make the use of AA easy and
 widespread, turning AA an ubiquitous system, presented on everyday communication
-channels. Even the work logs generated since July of 2011 could be vastly
-statistically analyzed aiming to recognize patterns in the behavior of
+channels. Even the work logs generated since July of 2011 could be 
+statistically analyzed to recognize patterns in the behavior of
 individuals and their productions.
 
-We would like to conclude setting an important role of AA: being a free software
-system and an open methodology, AA could be used to auto-manage groups of
-individuals working on software or other kinds of activities. In this way, we
-are interested to spread AA for those groups, to have even more developers
+Being a free software
+system and an open methodology, AA could be used to auto-manage groups
+working on software or other kinds of activities. In this way,
+there is interest on spreading AA to have even more developers
 contributing in a collaborative way.
@@
 
-We would like to also thank AA: the present research and even this manuscript
-was written using AA. The complete log is on-line at
+Authors thank AA: the present research and even this manuscript was written using AA. The complete log is on-line at
 \url{http://www.pulapirata.com/skills/aa}.
+Authors are also grateful to IFSC/USP, IPRJ/UERJ, IFT/UNESP and all AA users and collaborators, spacially those who coded AA hacks for shell and bot uses and those who coded the different Web interfaces in current use.
 

### 2013-06-01 paper.tex
adicionado Penalva nos autores e 2 bibs
some minor grammatical corrections

@@
 distribution of software teams~\cite{last2003}. This is responsible for the so
-called Global Software Development (GSD)~\cite{german2003}.  An example is the
+called Global Software Development (GSD)~\cite{german2003,Fryer,Begel}.  An example is the
 free software movement. Projects and institutions like Mozilla Foundation has
@@
 of developers working in the same place, time period and with minor cultural
-differences. Moreover, ``distributed development'' is generally tough as being
+differences. Moreover, ``distributed development'' is generally tought as being
 global which is not true. AA was applyed to a team that work at the
-same city but on different times (more details on Section~\ref{results}). Even
+same city but on different timings (more details on Section~\ref{results}). Even
 smaller groups of developers working on the same building could use GSD
@@
 to have good communication to work on tasks related to a specific problem. Cultural
-distance encompass differences on organizational and natural culture. Spoken
+distance encompass differences in organizational and natural culture. Spoken
 language, unit and ethnic values are common forms of this distance. Some
 companies prefer to situate development units in foreign locations with minimal
-cultural distance (e.g.\ an American firm prefer Ireland, because of spoken
+cultural distance (e.g. an American firm prefer Ireland, because of spoken
 language similarity~\cite{carmel2001}). Finally, the temporal distance
@@
 active free software projects and used by companies like Sun Microsystems. It is
-interesting to note a difference on the viewpoint present on his paper: German
+interesting to note a difference in the viewpoint present on his paper: German
 focuses the methodology description on code. He start explaining that GNOME is
@@
 one maintainer who divide his modules into separated parts in which other
-developers can work on independent tasks, along other responsibilities. As like
+developers can work on independent tasks, along with other responsibilities. As like
 others free software projects, all the development was done using a
@@
 management like SVN or Git. He finishes by observing that periodical, e.g. yearly, conferences like GUADEC
-is common on free and open source projects for face-to-face meetings and is based
+are common in free and open source projects for face-to-face meetings and are based
 each time on a different place.
@@
 This paper has presented a simple methodology to GSD, being the development conducted
-on large or small groups of software developers, working on different countries
+on large or small groups of software developers, working in different countries
 or even at the same room. The AA methodology implements a simple system where

### 2013-06-02 front-matter.tex
adicionado Penalva nos autores e 2 bibs
merged hybrid abstract changes

--- 2013-02-18 front-matter.tex
@@
 \begin{abstract}
-We present a new self-regulating methodology for coordinating
-distributed team work called Algorithmic Auto-regulation (AA), based
-on recent social networking concepts and individual merit. Team
-members take on an egalitarian role, and stay voluntarily logged into
-so-called AA sessions for part of their time (e.g.\ 2 hours per day),
-during which they create periodical logs --- short text sentences ---
-they wish to share about their activity with the team. These logs are
-publicly aggregated in a Website and are peer-validated, as in code
-review. A video is generally recorded by the members to make their
-work logs more understandable. This methodology is well-suited for
-increasing the efficiency of distributed teams working on what is
-called Global Software Development (GSD) through asynchronous
-on-demand communication, reducing the need for central management,
-unproductive meetings or time-consuming reports. The AA methodology
-also legitimizes the activities of a distributed software team.  It
-thus enables entities to have a solid means to fund these activities,
-allowing for new and concrete business models to emerge for very
-distributed software development. AA as a methodology is also at the
-core of having self-replicating hacker initiatives. These claims are
-discussed in a real case-study of running a distributed free software
-hacker team called Lab Macambira.
+We present a new self-regulating methodology for coordinating distributed team
+work called Algorithmic Auto-regulation (AA), based on recent social networking
+concepts and individual merit. Team members take on an egalitarian role, and
+stay voluntarily logged into so-called AA sessions for part of their time (e.g.\
+2 hours per day), during which they create periodical logs --- short text
+sentences --- they wish to share about their activity with the team. These logs
+are publicly aggregated in a Website and are peer-validated after the end of a
+session, as in code review.  A short screencast is ideally recorded at the end
+of each session to make AA logs more understandable.  This methodology has shown
+to be well-suited for increasing the efficiency of distributed teams working on
+what is called Global Software Development (GSD), as observed in our experience
+in actual real-world situations.  This efficiency boost is mainly achieved through 1)
+built-in asyncrhonous on-demand communication, documentation of work products
+and processes, and 2) reduced need for central management, meetings or
+time-consuming reports. Hence, the AA methodology legitimizes and facilitates
+the activities of a distributed software team.  It thus enables other entities
+to have a solid means to fund these activities, allowing for new and concrete
+business models to emerge for very distributed software development. AA has been
+proposed, at its core, as a way of sustaining self-replicating hacker
+initiatives. These claims are discussed in a real case-study of running a
+distributed free software hacker team called Lab Macambira.  
 \end{abstract}

### 2013-06-02 paper.tex
merged hybrid commit a45db4aa47012

--- 2013-05-31 paper.tex
@@
 \usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}
+%\usepackage{draftwatermark}
+
+\newcommand{\ie}{{\it i.e.}}
+\newcommand{\etc}{{\it etc}}
+\newcommand{\eg}{{\it e.g.}}
+\newcommand{\wrt}{{\it w.r.t. }}
+\newcommand{\etal}{{\it et.~al.}}
 
@@
 
-One of the defining features of modern times is the widening geographical
-distribution of software teams~\cite{last2003}. This is responsible for the so
-called Global Software Development (GSD)~\cite{german2003}.  An example is the
-free software movement. Projects and institutions like Mozilla Foundation has
-several employees and thousands of voluntary developers distributed across many
-countries. The same is true for GNOME~\cite{german2003}, OpenBSD, MySQL or
-Apache Software Foundation, to cite just a few\footnote{Ohloh, the open source
-  network, have a more complete and constantly updated list of the most active
-  projects on-line at \url{www.ohloh.net}}. Along the free and open software
-projects, GSD has a growing popularity in every niche of the software industry
-as a whole, even on those distributing their software with proprietary
-licenses. This phenomenon is attributed to a variety of factors such as a larger
-labor pool, natural globalization of software companies and foundations or even
-the premise of cheaper cost of production~\cite{komi2005}.
+One of the defining features of modern times is the widening
+geographical distribution of software teams~\cite{last2003} creating
+what is called Global Software Development (GSD)~\cite{german2003,Fryer,Begel}.
+An example is the free software movement. Projects and institutions like Mozilla
+Foundation has several employees and thousands of voluntary developers
+distributed across many countries. The same is true for GNOME~\cite{german2003},
+OpenBSD, MySQL or Apache Software Foundation, to cite just a few of the most
+active projects.\footnote{Ohloh, the open source network, have a more complete
+and constantly updated list of the most active projects on-line at
+\url{www.ohloh.net}} Along the free and open software projects, GSD is growing
+popularity in every niche of the software industry as a whole, even on those
+distributing their software with proprietary licenses. This phenomenon is
+attributed to a variety of factors such as a larger labor pool, natural
+globalization of software companies and foundations or even the premise of
+cheaper cost of production~\cite{komi2005}.
 
-Despite the advantages of GSD, it is known how difficult it is to coordinate and
-fund free software on a larger scale than currently available. That is, when
-teams are very heterogeneous, containing not only volunteers and experienced
-developers, but also contractors from different backgrounds and cultures. This
-observations are in consonance with dedicated studies, such
-as~\cite{carmel1999}.  As a common base, related difficulties for GSD gather
-around distance, time and cultural differences. In the case of free or open
-software projects, all these factors are involved.
+Despite the advantages of GCD, we have noticed how
+difficult it is to coordinate and fund free software on a larger scale
+than currently available, when teams are very heterogeneous containing
+not only volunteers and very experienced developers, but also
+contractors and freelancers from different backgrounds and cultures. Our observations
+are founded on the factors suggested by Carmel~\cite{carmel1999} as
+main difficulties for GCD: distance, time and cultural differences. In
+the case of free or open software projects, all these factors are
+involved.
 
-Another problem faced by modern software companies and other collectives are
-frequent ineffective meetings, which are seldom focused on the interest of any
-attendant. The result is that it has become the norm to participate in these
-with the ``laptop open'', which is unproductive. Software developers like to
-code, to be productive, to have their hands on their project, to do what they
-are best at. They dislike to have to stop for meetings or to write lengthy
-reports to justify their funding.
+Another problem faced by modern software companies and other
+collectives are frequent ineffective meetings, which are seldom
+focused to the interest to any attendant. The result is that it has
+become the norm to participate in too many meetings with the ``laptop
+open'', which can be un-productive. Software developers
+like to code, to be productive, to have their hands on their project,
+to do what they are best at. They dislike to forcibly stop for meetings
+or to do other bureaucratic activities such as writing lengthy reports to
+justify their funding.\cite{Thompson:Wired:2012}
 %\todo{ler mais. cacm, etc}
 
-To address these matters is the purpose of AA, a methodology and an associated
-software system for coordinating distributed team work dealing with the
-disadvantages of GSD. Team members take on an egalitarian role, and stay
-voluntarily \textit{logged} in the system for part of their time (e.g.\ 2 hours
-per day), during which they log a periodical short text sentence --- similar to
-a ``tweet'' in this Twitter era --- as the status of their activity using an
-easy to use command. These ``micro-blog sentences'' are publicly aggregated and
-validated by other team members. Through AA, the community has a methodology and
-an associated system to validate and enable the activities of a distributed
-software team. It implicitly legitimizes financial support for the expansion of
-the activity of the developer team as the activities are documented and
-validated.  The AA methodology is specially useful for coordinating distributed
-and decentralized team work, providing effective means to asynchronously update
-different team members, diminishing the need for harsh synchronous meetings.
+In this paper we propose the AA methodology and an associated software system
+for coordinating distributed team work, tackling the disadvantages
+of GSD. Team members take on an egalitarian role, and stay
+voluntarily \textit{logged} in the system for part of their time
+(e.g.\ 2 hours per day), during which they log a periodical short text
+sentence or \emph{microlog} --- similar to a `tweet' from Twitter --- as the
+status of their activity. Logging is carried out using an easy a series of
+available UI's: unix shell
+commands, native GUI or web page, conventional social network posts, or chat messages to a log bot
+listening to IRC, GTalk, G+, and others.  These ``microblog sentences'' are publicly aggregated and validated by other
+team members. Through AA, we have a methodology and an associated system to help
+implement and validate the activities of a distributed software team. It implicitly
+legitimizes financial support for the expansion of the activity of the developer
+team. The AA methodology is specially useful for coordinating distributed and
+decentralized team work, providing effective means to asynchronously update
+different team members without the need for synchronous unproductive meetings.
 
@@
 
-A brief overview of current papers about GSD methodologies related with AA is
-presented in section~\ref{related-work}. In section~\ref{aa-methodology}, the
-most relevant characteristics of the AA methodology are outlined. In this same
-section, there is a description of an AA use in a team of 9 developers in the
-second half of 2011 and a broader a more released use of AA from 2012 until
-now. Finally, in Section~\ref{conclusions}, there are final conclusions and
-indicatives of future possibilities to the practical use of AA on other teams of
-software developers or individuals working on non-software activities.
+A brief overview of current work on GSD methodologies related to
+AA is presented in Section~\ref{related-work}, while in
+Section~\ref{aa-methodology} we describe the most relevant
+characteristics of the AA methodology followed by the description of
+our experience using AA in a team of 9 paid developers 
+% nao era mais de 9?
+since July 2011. In Section~\ref{conclusions} we draw some final
+conclusions and indicate future possibilities for the
+practical use of AA in other types of software developers teams or
+organizations working on non-software distributed activities.
 
@@
 on the digital age enables people to be very productive, the article also
-mentions Linux as a hallmark example~\cite{Thompson:Wired:2012}
+mentions linux as a hallmark example~\cite{Thompson:Wired:2012}
 }
@@
 \label{related-work}
+\nocite{gobbo:APSEEP2008}
+\nocite{Reis:PhDThesis:2003}
 
@@
 
-There has been a large amount of research done in the area of methodologies to
-deal with distributed teams of developers. We are focusing in GSD here, however
-some principles involved on those methodologies could be used on smaller teams
-of developers working in the same place, time and with minor cultural
-differences. Moreover we generally think on ``distributed development'' being
-global which is not totally true. We even applied AA to a team that work at the
-same city but on different times (more details on Section~\ref{results}). Even
-smaller groups of developers working on the same building could use GSD
-methodologies. A thorough survey of these methodologies is beyond the scope of
-this paper. Here we present a brief overview.
+There has been a large amount of research done in the area of
+methodologies to deal with distributed teams of developers. We are
+focusing in GSD here, however some principles involved on those
+methodologies could be used on smaller teams of developers working in
+the same place, time and with minor cultural differences. Moreover we
+generally think on ``distributed development'' being global which is
+not totally true. We even applied AA to a team that work at the same
+city but on different times (more details on
+Section~\ref{results}). Even smaller groups of developers working on
+the same building could use GSD methodologies. A thorough survey of
+these methodologies is beyond the scope of this paper. Here we present
+a brief overview.
 
-Various methodologies for GSD were built around the factors that affect
-distributed team works, proposed by Carmel~\cite{carmel1999} and comprising
-three distances: geographical, cultural and temporal. Geographical distance
-prejudice \emph{coordination}, being the act of integrating all the tasks
-distributed between units~\cite{carmel2001}; \emph{control}, or the process to
-maintain specific goals, policies or quality levels; and
-\emph{communication}. All these factors are correlated, for example, a team need
-to have clearly communication to work on tasks of a specific problem. Cultural
-distance encompass differences on organizational and natural culture. Spoken
-language, unit and ethnic values are common forms of this distance. Some
-companies prefer to situate development units in foreign locations with minimal
-cultural distance (e.g.\ an American firm prefer Ireland, because of spoken
-language similarity~\cite{carmel2001}). And finally the temporal distance that
-prejudices synchronous communications like telephone or video-conferences. Units
-of developers working on different time-zones are concerned with managing of
-their agenda guided by this temporal distance.
+Various methodologies for GSD were built around the factors that
+affect distributed team works, proposed by Carmel~\cite{carmel1999}
+and comprising three distances: geographical, cultural and
+temporal.
 
-Targeting geographical distance, Carmel~\cite{carmel2001} suggests a tactic to
-reduce intensive collaboration. His approach divides the whole software
-life-cycle on levels of complexity. Each level has a degree of
-collaboration. For example, some developers working on a project with high
-collaboration level should use the follow-the-sun approach: when concluding the
-work day, they pass their work to the team working on another time-zone. Other
-tactics are suggested by the same author to deal with the three distances, like
-separating foreign units of developers in time-zone bands.
+First, geographical distance handicaps ($i$) \emph{coordination}, the act of
+integrating all the tasks distributed between units~\cite{carmel2001}; ($ii$)
+\emph{control}, or the process to maintain specific goals, policies or quality
+levels; and ($iii$) \emph{communication}. All those factors are correlated, \eg,
+a team needs to have clear communication to work on tasks of a specific problem.
 
-Battin et al.~\cite{battin2001} propose and argues about their experiments using
-specific methodologies created for the distributed development centers of the
-Motorola Company (which has over 25 software development centers
-worldwide). These methodologies included constant communication with critical
-units, incremental integration and scheduled based on time-zones distributed to
-developers on 6 countries from 3 continents.
+Second, cultural distance encompasses differences on organizational and
+natural culture. Spoken language, unit and ethnic values are common
+forms of such distance. Some companies prefer to allocate development
+units in foreign locations with minimal cultural distance (\eg, an
+American company may prefer Ireland due to spoken language
+similarity~\cite{carmel2001}). Third, the temporal distance that
+hampers synchronous communications like telephone or
+videoconferences. Units of developers working on different time-zones
+are concerned with managing of their agenda guided by this temporal
+distance.
 
-While considering free software projects instead of companies, the same factors
-are present and some methodologies arises. German~\cite{german2003} gives a
-concise review of methodologies used by the GNOME project, one of the most
-active free software projects and used by companies like Sun Microsystems. It is
-interesting to note a difference on viewpoint present on his paper: German
-focuses the methodology description on code. He start explaining that GNOME is
-separated into modules (76 on version 2.4, to be precise) and each module has
-one maintainer who divide his modules into separated parts in which other
-developers can work on independent tasks, along other responsibilities. As like
-others free software projects, all the development was done using a
-\emph{bugtracker} to bugs and issues management, mail lists and Internet Relay
-Chat (IRC) to discussion and communication and a software configuration
-management like SVN or Git. Periodical (commonly yearly) conferences like GUADEC
-is common on free and open source projects to face-to-face meeting and is based
-each time on a different place.
+Targeting geographical distance, Carmel~\cite{carmel2001} suggests a
+strategy to reduce intensive collaboration. His approach divides the
+whole software life-cycle into levels of complexity. Each level has a
+degree of collaboration. For example, some developers working on a
+project with high collaboration level should use the follow-the-sun
+approach: when concluding the work day, they pass their work to the
+team working on another time-zone. Other tactics are suggested by the
+same author to deal with the three distances, such as separating foreign
+units of developers in time-zone bands.
+
+Battin~\etal~\cite{battin2001} propose and discuss their
+experiments using specific methodologies created for the distributed
+development centers from Motorola (at the time having 25+
+software development centers worldwide). These methodologies included
+constant communication with critical units, incremental integration
+and schedules based on time-zones distributed to developers on 6
+countries from 3 continents.
+
+In considering free software projects instead of companies,
+similar factors are present and specialized methodologies
+arise. German~\cite{german2003} provides a concise review of
+the methodologies used by the GNOME project, one of the most active of all free
+software projects. The manuscript is centered on the software architecture. It
+begins by explaining that GNOME is separated into modules (76 on version 2.4, to be
+precise) and each module has one maintainer who divides his modules into
+separate parts in which other developers can work on independent tasks, along
+other responsibilities. As in other free software projects, all 
+development was done using a \emph{bugtracker} for bug and issue management,
+mail lists and Internet Relay Chat (IRC) for discussion and communication and a
+version control system like Git or Mercurial. Periodical (commonly yearly)
+conferences like GUADEC is common on free and open source projects for
+face-to-face meetings and is based on a different place each time.
 
@@
 
-As noted, some strategies for GSD is based on complex methodologies and many of
-those were built for a specific company or software center. Here we propose an
-alternative methodology based on a simple idea: small working sessions logged by
-a computational tool. Figure~\ref{fig:mm} summarizes our methodology.
+Some of the strategies for GSD mentioned on the previous section are based on complex methodologies
+and many were built for a specific company or software
+center. We propose an alternative methodology based on a simple
+idea: small working sessions logged by a computational
+tool. Figure~\ref{fig:mm} summarizes our methodology.
 
@@
 
-From the developer perspective, the AA methodology is based on creating pretty
-small high perspective reports of what they are doing in a specific time frame,
-that can be something between 5 to 15 minutes, depending in what is more
-comfortable for the developer. An AA Session would be a period of at least 2
-hours doing these reports. The developer can set reminders to show up when its
-time to make a new report. The objective of the flexible time frame and
-reminders is to minimize developer overhead during his AA session. In this way
-he can make the reports while staying concentrated in his code. Each report can
-be sent directly to an on-line server, or stored locally in a temporarily
-database for sending later. This make possible the use of AA while offline.
+From the developer's perspective, the AA methodology is based on
+creating pretty small high perspective reports of what they are doing
+in a specific time frame, that can be something between 5 to 15
+minutes, depending on what is most convenient for the developer and the team. An
+\emph{AA Session} would be a period of at least 2 hours doing these periodical
+\emph{micrologs} also called \emph{AA posts}. The developer can set reminders or alerts to show up when its time to
+\emph{microlog}. The objective of the flexible time frame and
+alert scheme is to \emph{minimize developer overhead} during his AA session. In
+this way, the developer can issue micrologs while staying maximally concentrated in his
+code. Each microlog may be sent directly to an on-line server, or stored
+locally in a temporarily database for sending/pushing later. This enables
+offline micrologging and periodical alerting.
 
-Developers can also record a video screencast in the end of the session
-summarizing what has been done, explaining with his words and showing his most
-important results. This, combined with the textual log of his session, makes the
-whole report more understandable for himself or other people searching for
-information about his production.
+Developers optionally record a \emph{video screencast} at the end of the
+session summarizing what has been done, explaining with his words and
+showing his most important results. This is very similar to the video logging system
+in the movie Avatar, although it is proved from our July 2011 git repositories and online
+wiki that we have used this powerful concept in AA prior to the release of
+Avatar. Moreover, screencasting differs from general videologging in that it
+typically captures actual workflow on
+the computer screen. Screencasting, combined with the textual log of the AA
+Session, renders the final report more understandable for the individual developer himself
+or to other people searching for information about his production.
 
@@
 
-All AA reports made by the developers are sent to a Web server and become public
-in a Website, being possible then for a manager or another developer to follow
-very closely the work of a developer, nearly real-time, reading each of his
-small reports of what he is doing right now.
+All AA reports made by the developers are ultimatelly sent to a Web server and
+become publically aggregated on a Website, enabling a manager or another
+developer to follow very closely the work of a given developer, nearly real-time,
+reading each of his small reports or micrologs of what he is doing right now.
 
-Another possibility is to check older sessions to see when sometimes was done
-and the comments of the developer about it, since each AA post happens in a very
-small time frame of work, the information about what was done become very easy
-to understand, instead of a long report in the end of a session.
+Another use case is to check older sessions to check when certain tasks
+were carried out and the comments of the developer about the \emph{process}.
+Since each AA microlog happens in a very short timeframe of work, the information
+about what was was done -- specially \emph{how} it was done -- becomes very easy to
+understand, instead of a long report in the end of a session.
 
-The site is where the developer can add to his session a link for his screencast
-about that session, a video summary of what was done to complement the reports,
-useful specially on cases where the small reports were done in a hurry, because
-the developer did not want to lose his focus on something important at that
-moment.
+IN the current version of the AA server infrastructure, the aggregating
+website is where the developer can attach a link for his
+screencast about each session. Screencasts are specially useful on cases where
+the small reports were done in a hurry, because the developer did not want to
+lose his focus on something important at that moment.
 
-\subsection{Validation}
+\subsection{Peer Validation and Peer Pressure}
 
-Each AA session must be validated by another developer, it means that all
-reports are read by someone that will consider then valid or not and will even
-write commentaries about the specific session. The developer in charge of
-validating a session is decided randomly by the AA Web server, which send an
-email to the chosen developer with an URL to a validation interface.
+No set bosses or leaders are required in an ideal aplication the AA methodology;
+in practice the role of bosses basically disappears or is greatly lightened --
+hence the name Algorithmic \emph{Autoregulation} and other implicit
+interpretations to the AA acronym and symbol. The primary mechanism to achieve
+AA is peer coordination and social behavior, be it deliberate or implicit. In
+order to prevent spamming and to improve the overall quality of AA reports, each
+AA session must be validated by another developer. More specifically, all
+reports are read by someone that will mark them as `valid' or `invalid' and may
+optionally write commentaries about the specific session and quality of
+micrologs. The developer in charge of validating any given session is randomly
+assigned by the AA Web server, which sends an email to the chosen developer with
+an URL to a validation interface.
 
@@
 
-The easy and effective management of teams working on GSD is the main purpose of
-the AA methodology. We applied this methodology to a group of 9 developers in
-July of 2011 on what we called Lab Macambira~\footnote{LabMacambira.sf.net:
-  \url{http://labmacambira.sf.net}}. The main objective of the team was to work
-on different free software projects, contributing directly to their development,
-sending bug corrections or proposing new features on their source code.
+Easy and effective GSD team management is the main
+purpose of the AA methodology. We applied the proposed methodology to a group
+of 9 developers in July of 2011 on what we called Lab
+Macambira~\footnote{LabMacambira.sf.net: \url{http://labmacambira.sf.net}}. The
+main objective of the team was to work on an array of strategic free software
+projects in the broad audiovisual and web categories,
+contributing directly to their development, submitting bug patches or
+committing new features to their source code.
 
-All the team members had different levels of knowledge on software development
-as part of large and distributed free software projects like Scilab or
-Mozilla. In this way, one month of training was conducted by two experienced
-developers, teaching the basics of use of development support tools like
-bugtrackers, programming languages and version control systems. After this
-period, a challenge was proposed for the new developers: send a bug correction
-or a new feature to a large free software project and you will work with us, you
-will be a ``Macambira'' developer. Table~\ref{tabela:contribuicoes} summarizes
-the contributions of each ``Macambira'' to free software projects.
+The team members had different levels of experience on software
+development for large and distributed free software projects
+like Scilab or Mozilla. In this way, one month of training was
+conducted by two experienced developers, teaching the basics of use of
+development support tools like bugtrackers, programming languages,
+version control systems, and build systems. After this period, a starter project
+was proposed for new developers, namely to have an accepted patch or commit to
+the official repository for a large and widely used free software app.
+Developers passing the starter project would be deemed `initiated' and be
+called a `Macambira' developer.  Table~\ref{tabela:contribuicoes} summarizes the
+effective accepted contributions of each successfully initiated developer to
+free software projects in 2011, which used the AA methodology.
 
@@
   \caption{Free and open software projects that received contributions
-    from ``Macambiras''. On the first column we can see a list of
-    applications of those projects. At right, the pseudonym of the
+    from successfully initiated developers  or `macambiras' that worked under
+    the AA methodology. The first column lists applications to which
+    contributions were officially accepted and whose development process was
+    performed and logged using AA. The second column shows the pseudonym of the
     ``Macambiras'' who sent \textit{commits} to the application. At
-    ``Lab Macambira'', and at free software community in general, is a
-    common practice to use pseudonym as identification.}
-    \small\begin{tabular}{|l|l|}
+    Lab Macambira, and at the free software community at large, 
+    It is
+    common practice at Lab Macambira, and the free software community at large,
+    to use pseudonyms as identification in order to enhance privacy in practice.}
+    \begin{tabular}{|l|l|}
         \hline
@@
 
-In one month, each developer contributed to many very large free software
-projects. Many of the developers started the training with no knowledge of what
+In one month, each developer officially contributed to one or many free software
+projects. Many developers started the initiation training with no knowledge of what
 is free software and ended that period becoming a free software developer.
+During that month, the same team of trainees also developed the first version of the AA
+system and used AA to manage their activities, even while developing
+the other aforementioned free software projects.
 
-During that month, the same team developed the first version of the AA system
-and used AA to manage their activities. Even while developing the system. All
-the source code of AA --- the client that sends the logs and the AA Web server
---- is public available~\footnote{AA source code:
+The source code of AA --- both the client that sends the
+logs and the AA Web server --- is public available as free software~\footnote{AA source
+  code:
   \url{http://labmacambira.git.sourceforge.net/git/gitweb.cgi?p=labmacambira/aa}}
-and all the AA sessions log of the whole team of ``Lab Macambira'' is also
-on-line~\footnote{Logs of AA sessions:
+and all the actual AA log data of the entire team of Lab Macambira from 2011 to
+the present time is also on-line~\footnote{Logs of AA sessions:
   \url{http://labmacambira.git.sourceforge.net/git/gitweb.cgi?p=labmacambira/paainel}}.
 
-After the training period, during more 6 months, the ``Macambiras'' worked on a
-large range of free software projects, distributed on work groups --- each work
-group focusing on a specific theme like video, audio and web --- and financed by
-contracts and support of the ``Pont\~{a}o N\'{o}s Digitais''. In
-Table~\ref{tabela:criados} we can see a list of the free software created by
-``Lab Macambira'' since July of 2011.
+After the initial training period of 1 month, the ``Macambiras''
+worked during 6 additional months on a large range of free software projects, distributed on work
+groups --- each work group focusing on a specific theme like video,
+audio and web --- and funded by contracts, freelance, and support of the
+Pont\~{a}o N\'{o}s Digitais~\footnote{\url{http://nosdigitais.teia.org.br}}. 
+Table~\ref{tabela:criados} has a list of \emph{new} free software applications created by
+``Lab Macambira'' since July 2011.
 
 \begin{table}
-    \caption{Software projects created by ``Lab Macambira'' since July
-      of 2011 with a short description and the technologies --- like
-      programming languages or frameworks --- involved. It is
-      interesting to note the heterogeneity of projects and its areas
+    \caption{Software projects created by Lab Macambira since July
+      2011 with a short description and the technologies involved (\eg,
+      programming languages or frameworks). It is
+      interesting to note the heterogeneity of projects and their areas
       of application.}
+
+    \footnotesize{
     \begin{tabular}{|l|p{5cm}|l|}
@@
 
-As of this writing the ``Lab Macambira'' have many software developers, and some
-of the trained developers continue to work voluntarily in the project.
+As of this writing the `Lab Macambira' unites 15 software developers, and 
+key 2011-initiated developers continue to work voluntarily in the project.
 
@@
 %% brief introduction
-In a scenario where GSD is growing as a popular form of software development,
-not just on free software projects but in the whole software industry, we need
-methodologies to deal with its disadvantages and at the same time to amplify its
-advantages.
+In a scenario where Global Software Development is growing as a popular 
+software development scenario both for free software projects and for the entire
+software industry, we need methodologies to deal with its potential disadvantages and at
+the same time to amplify its advantages.
 
-This paper has presented a methodology to GSD, being the development conducted
-on large or small groups of software developers, working on different countries
-or even at the same room. The AA methodology implements a simple system where
-each developer take notes of his work generating a periodical log of small text
-sentences. The sum of those sentences, along an entire session of work, results
-in a complete report. The report is made public available through a Website and
-be validated by other developers sorted aleatory by the AA Web server.
+This paper has presented a methodology for GSD, \ie, the development
+involving a series of large or small groups of software developers, each either working on
+different countries or at the same room. The AA methodology
+implements a simple system where each developer take notes of his work
+posting a periodic log of small text sentences or microblogs. The sum of those
+micrologs, along with an entire session of work, results in a complete
+report. The report is made public available through a Website and be
+validated by other developers randomly sorted by the AA Web server.
 
-Instead of a merely work-management tool, AA act as a methodology to improve the
-time sense of individuals, dividing their work on small sessions, and also
-reducing the need of extensive reports or unnecessary meetings. By asking users
-to write a minimal text sentence as a continuously log, AA does not disturb
-developers concentrated on programming: developers just have to type some
-characters, hit \textit{enter} and go back to coding.
+AA is not limited to a merely work-management tool, but acts as \emph{a
+methodology to improve the time sensibility of individuals}, dividing their
+complex work on small chunks or sessions, and also reducing the need of extensive
+reports or unnecessary meetings. By asking users to write a minimal
+text sentence as a continuous log feed, the proposed methodology avoids
+disturbing the flow of developers which are heavily concentrated on programming: developers
+just have to type a few words and go back to coding.
 
-AA application is not restricted to software development. As of this writing
-there is a comic book studio~\footnote{Pula pirata: \url{http://pulapirata.com}}
-starting to use AA to manage their activities. People with non-software
-background, like social scientists, musicians and activists has also using AA
-and contributing for its improvement.
+The AA methodology is not restricted to software development, even though it was
+designed for the latter. As of this
+writing there is an entertainment studio, Pula Pirata~\footnote{
+\url{www.pulapirata.com}}, that has started to use AA to manage their creative activities.
+Other people with no software background, like social scientists, musicians and
+activists also have been using AA and contributing for its improvement.
 
-For developer teams, we have experienced the use of AA to auto-regulate the work
-of ``Lab Macambira'', a group of free software developers from Brazil. Since
-July of 2011 the group have contributed and created new free and open source
-software for a vast number of applications.
+For developer teams, we have analyzed data and gained practical experience on
+the use of AA to auto-regulate the work of Lab Macambira, a group of free
+software developers from Brazil. Since July of 2011 the group has contributed
+and created new free and open source software for a vast number of additional applications.
 
-There are many aspects of the work which remain unfinished. New ways to report
-logs --- the ``Twitter like'' messages --- from different interfaces like IRC,
-Internet Messaging services and email can make the use of AA easy and
-widespread, turning AA an ubiquitous system, presented on everyday communication
-channels. Even the work logs generated since July of 2011 could be vastly
-statistically analyzed aiming to recognize patterns in the behavior of
-individuals and their productions.
-
-We would like to conclude setting an important role of AA: being a free software
-system and an open methodology, AA could be used to auto-manage groups of
-individuals working on software or other kinds of activities. In this way, we
-are interested to spread AA for those groups, to have even more developers
-contributing in a collaborative way.
+There are many aspects of the work which remain unfinished. Additional
+ubiquitous client
+interfaces for micrologging -- the `Twitter-like' messages -- from different
+interfaces beyond IRC, \eg, web social services and email, would greatly make the use of
+AA easier and more widespread, turning AA a truly ubiquitous and replicable system, presented on everyday
+communication channels. Another research direction is that the actual work logs
+generated by the Lab Macambira and Pula Pirata collectives since July of 2011
+could be vastly statistically analyzed aiming to recognize patterns in the
+behavior of individuals and their creative production process.
 
 \section*{Acknowledgments}
-%The authors would like to thank NSF and CNPq.
+%TODO The authors would like to thank NSF and CNPq.
 
-We would like to also thank AA: the present research and even this manuscript
-was written using AA. The complete log is on-line at
-\url{http://www.pulapirata.com/skills/aa}.
+We would like to thank AA: the present research and even this
+manuscript was written using AA. The complete log is on-line at
+\url{www.pulapirata.com/skills/aa}.
 
@@
 \bibliographystyle{acm}
-\bibliography{aa.bib}
+\bibliography{aa,personal}
 %}

### 2013-06-02 paper.tex
footnote spacing and part of 1st paragraph of intro fixed

@@
 
-One of the defining features of modern times is the widening
-geographical distribution of software teams~\cite{last2003} creating
-what is called Global Software Development (GSD)~\cite{german2003,Fryer,Begel}.
+One of the defining features of modern times is the widening geographical
+distribution of software teams~\cite{last2003} creating what is called Global
+Software Development (GSD)~\cite{german2003,Fryer,Begel}.  
 An example is the free software movement. Projects and institutions like Mozilla
@@
 distributed across many countries. The same is true for GNOME~\cite{german2003},
-OpenBSD, MySQL or Apache Software Foundation, to cite just a few of the most
-active projects.\footnote{Ohloh, the open source network, have a more complete
+OpenBSD, MySQL or Apache Software Foundation, to cite just a 
+few of the most active projects.\footnote{Ohloh, the open source network, has a more complete
 and constantly updated list of the most active projects on-line at
@@
 of 9 developers in July of 2011 on what we called Lab
-Macambira~\footnote{LabMacambira.sf.net: \url{http://labmacambira.sf.net}}. The
+Macambira.\footnote{LabMacambira.sf.net: \url{http://labmacambira.sf.net}} The
 main objective of the team was to work on an array of strategic free software
@@
 The source code of AA --- both the client that sends the
-logs and the AA Web server --- is public available as free software~\footnote{AA source
+logs and the AA Web server --- is public available as free software\footnote{AA source
   code:
@@
 and all the actual AA log data of the entire team of Lab Macambira from 2011 to
-the present time is also on-line~\footnote{Logs of AA sessions:
+the present time is also on-line\footnote{Logs of AA sessions:
   \url{http://labmacambira.git.sourceforge.net/git/gitweb.cgi?p=labmacambira/paainel}}.
@@
 audio and web --- and funded by contracts, freelance, and support of the
-Pont\~{a}o N\'{o}s Digitais~\footnote{\url{http://nosdigitais.teia.org.br}}. 
+Pont\~{a}o N\'{o}s Digitais\footnote{\url{http://nosdigitais.teia.org.br}}. 
 Table~\ref{tabela:criados} has a list of \emph{new} free software applications created by
@@
 designed for the latter. As of this
-writing there is an entertainment studio, Pula Pirata~\footnote{
+writing there is an entertainment studio, Pula Pirata\footnote{
 \url{www.pulapirata.com}}, that has started to use AA to manage their creative activities.

### 2013-06-02 paper.tex
integrado ate linha 83 com hybrid

@@
 and constantly updated list of the most active projects on-line at
-\url{www.ohloh.net}} Along the free and open software projects, GSD is growing
-popularity in every niche of the software industry as a whole, even on those
+\url{www.ohloh.net}} Beyond the free and open software community, GSD has a growing
+popularity in every niche of the software industry as a whole, even among those
 distributing their software with proprietary licenses. This phenomenon is
@@
 
-Despite the advantages of GCD, we have noticed how
-difficult it is to coordinate and fund free software on a larger scale
-than currently available, when teams are very heterogeneous containing
-not only volunteers and very experienced developers, but also
-contractors and freelancers from different backgrounds and cultures. Our observations
-are founded on the factors suggested by Carmel~\cite{carmel1999} as
-main difficulties for GCD: distance, time and cultural differences. In
-the case of free or open software projects, all these factors are
-involved.
+Despite the advantages of GSD, we have noticed how
+difficult it is to coordinate and fund free software on a significantly larger scale
+than currently practiced, as series of qualitatively new situations arise. Distributed
+teams are very heterogeneous containing not only volunteers and very experienced
+developers, but also contractors and freelancers from different backgrounds and
+cultures. Our observations are founded on the factors suggested by
+Carmel~\cite{carmel1999} as main difficulties for GSD: distance, time and
+cultural differences. In the case of free or open software projects, all these
+factors are involved.
 
@@
 collectives are frequent ineffective meetings, which are seldom
-focused to the interest to any attendant. The result is that it has
+focused on the particular interest of any attendant. The result is that it has
 become the norm to participate in too many meetings with the ``laptop

### 2013-06-02 paper.tex
headers of sections lowcaps

@@
 sentence or \emph{microlog} --- similar to a `tweet' from Twitter --- as the
-status of their activity. Logging is carried out using an easy a series of
-available UI's: unix shell
+status of their activity. Logging is carried out using a series of
+easy UI alternatives: unix shell
 commands, native GUI or web page, conventional social network posts, or chat messages to a log bot
-listening to IRC, GTalk, G+, and others.  These ``microblog sentences'' are publicly aggregated and validated by other
-team members. Through AA, we have a methodology and an associated system to help
+listening to IRC, GTalk, G+, and others.  These ``microblog sentences'' are
+publicly aggregated and validated by other team members. Through AA, the
+community has a
+methodology and an associated system to help
 implement and validate the activities of a distributed software team. It implicitly
-legitimizes financial support for the expansion of the activity of the developer
-team. The AA methodology is specially useful for coordinating distributed and
-decentralized team work, providing effective means to asynchronously update
-different team members without the need for synchronous unproductive meetings.
+legitimizes financial support for the expansion of the activity of a
+distributed development team. The AA methodology is specially useful for
+coordinating distributed and decentralized team work, providing effective means
+to asynchronously update different team members without the need for synchronous
+unproductive meetings.
 
@@
 
-A brief overview of current work on GSD methodologies related to
-AA is presented in Section~\ref{related-work}, while in
-Section~\ref{aa-methodology} we describe the most relevant
-characteristics of the AA methodology followed by the description of
-our experience using AA in a team of 9 paid developers 
+A brief overview of current work on GSD methodologies related to AA is presented
+in Section~\ref{related-work}. In Section~\ref{aa-methodology} the most relevant
+characteristics of the AA methodology are outlined. In the same section, we
+report an actual use case of AA for coordinating a team of 9 paid developers 
 % nao era mais de 9?
-since July 2011. In Section~\ref{conclusions} we draw some final
-conclusions and indicate future possibilities for the
-practical use of AA in other types of software developers teams or
-organizations working on non-software distributed activities.
+during the second half of 2011, as well as a broader use case of AA from 2012 to
+the present time.  In Section~\ref{conclusions} we summarize the overall conclusions
+and indicate future possibilities for the practical use of AA in other types of
+teams of software developers or organizations working on non-software distributed
+activities.
 
@@
 
-\section{Related Work}
+\section{Related work}
 \label{related-work}
@@
 
-\section{The AA Methodology}
+\section{The AA methodology}
 \label{aa-methodology}
@@
 
-\subsection{AA Session}
+\subsection{AA session}
 
@@
 
-\subsection{AA Website Report}
+\subsection{AA website report}
 
@@
 
-\subsection{Peer Validation and Peer Pressure}
+\subsection{Peer validation and peer pressure}
 
@@
 
-\section{Results and Discussion}
+\section{Results and discussion}
 \label{results}

### 2013-06-02 paper.tex
ate l 155 consolidado com hybrid

@@
 
-There has been a large amount of research done in the area of
-methodologies to deal with distributed teams of developers. We are
-focusing in GSD here, however some principles involved on those
-methodologies could be used on smaller teams of developers working in
-the same place, time and with minor cultural differences. Moreover we
-generally think on ``distributed development'' being global which is
-not totally true. We even applied AA to a team that work at the same
-city but on different times (more details on
-Section~\ref{results}). Even smaller groups of developers working on
-the same building could use GSD methodologies. A thorough survey of
-these methodologies is beyond the scope of this paper. Here we present
-a brief overview.
+There has been a large amount of research in methodologies to deal with
+distributed teams of developers. Although this paper focuses on GSD, some of the
+same
+principles could also be adapted for smaller teams of
+developers working at the same place, timezone and with minor cultural
+differences, depending on the specific context and demands.
+Moreover, `distributed development' is generally tought of as being
+global which is not true. For instance, AA has been effectively applied to a team whose
+members live in the same city but work at different timeframes and locations, see 
+Section~\ref{results}.  Even smaller groups of developers working on the same
+building could use GSD methodologies (or an adapted subset) to their benefit,
+\eg, to account for different work habits, minimize formal meetings, document work
+process and history, and so on. A thorough survey of these methodologies is beyond the scope
+of this paper; this section presents but a brief overview.
 

### 2013-06-02 paper.tex
working on mindmap fig

@@
 There has been a large amount of research in methodologies to deal with
-distributed teams of developers. Although this paper focuses on GSD, some of the
-same
-principles could also be adapted for smaller teams of
-developers working at the same place, timezone and with minor cultural
-differences, depending on the specific context and demands.
-Moreover, `distributed development' is generally tought of as being
-global which is not true. For instance, AA has been effectively applied to a team whose
-members live in the same city but work at different timeframes and locations, see 
-Section~\ref{results}.  Even smaller groups of developers working on the same
-building could use GSD methodologies (or an adapted subset) to their benefit,
-\eg, to account for different work habits, minimize formal meetings, document work
-process and history, and so on. A thorough survey of these methodologies is beyond the scope
-of this paper; this section presents but a brief overview.
+distributed teams of developers. Although this paper focuses on GSD, some of
+its principles could also be adapted for smaller teams of developers working at
+the same place, timezone and with minor cultural differences, depending on the
+specific context and demands.  Moreover, `distributed development' is generally
+tought of as being global which is not true. For instance, AA has been
+effectively applied to a team whose members live in the same city but work at
+different timeframes and locations, see Section~\ref{results}.  Even smaller
+groups of developers working on the same building could use GSD methodologies
+(or an adapted subset) to their benefit, \eg, to account for different work
+habits, minimize formal meetings, document work process and history, and so on.
+A thorough survey of these methodologies is beyond the scope of this paper; this
+section presents but a brief overview.
 
 Various methodologies for GSD were built around the factors that
-affect distributed team works, proposed by Carmel~\cite{carmel1999}
-and comprising three distances: geographical, cultural and
-temporal.
-
+affect distributed teamwork. As proposed by Carmel~\cite{carmel1999}, these
+comprise three distances: geographical, cultural and temporal. 
 First, geographical distance handicaps ($i$) \emph{coordination}, the act of
@@
 
-Battin~\etal~\cite{battin2001} propose and discuss their
-experiments using specific methodologies created for the distributed
-development centers from Motorola (at the time having 25+
-software development centers worldwide). These methodologies included
-constant communication with critical units, incremental integration
-and schedules based on time-zones distributed to developers on 6
+Battin~\etal~\cite{battin2001} propose and discuss their experiments using
+specific methodologies created for the distributed development centers from
+Motorola (at the time having 25+ software development centers worldwide). These
+methodologies included constant communication with critical units, incremental
+integration and schedules based on time-zones distributed to developers on 6
 countries from 3 continents.
@@
 begins by explaining that GNOME is separated into modules (76 on version 2.4, to be
-precise) and each module has one maintainer who divides his modules into
+precise) and each module has one maintainer who divides her modules into
 separate parts in which other developers can work on independent tasks, along
@@
 
-Some of the strategies for GSD mentioned on the previous section are based on complex methodologies
-and many were built for a specific company or software
-center. We propose an alternative methodology based on a simple
-idea: small working sessions logged by a computational
-tool. Figure~\ref{fig:mm} summarizes our methodology.
+Some of the strategies for GSD mentioned on the previous section are based on
+complex methodologies and many were built for a specific company or software
+center. We propose an alternative methodology based on a simple idea: small
+working sessions logged by a computational tool. Figure~\ref{fig:mm} summarizes
+this methodology.
 

### 2013-06-02 paper.tex
synced w hybrid up until sec AA session
ate l 262 consolidado, agora vendo a fig aa-0_1_

@@
    \caption{
-   A mindmap of our methodology
+   A mindmap of the AA methodology: 1) `developer engagement cycle', \ie, his use
+   of AA; 2) `functionality', \ie, main goals of the system; 3) `potentialities',
+   \ie, enhancements AA delivers to the developer's context. 
    }
@@
 
-\subsection{AA session}
+\subsection{The AA session}
 
-From the developer's perspective, the AA methodology is based on
-creating pretty small high perspective reports of what they are doing
-in a specific time frame, that can be something between 5 to 15
-minutes, depending on what is most convenient for the developer and the team. An
-\emph{AA Session} would be a period of at least 2 hours doing these periodical
-\emph{micrologs} also called \emph{AA posts}. The developer can set reminders or alerts to show up when its time to
-\emph{microlog}. The objective of the flexible time frame and
-alert scheme is to \emph{minimize developer overhead} during his AA session. In
-this way, the developer can issue micrologs while staying maximally concentrated in his
-code. Each microlog may be sent directly to an on-line server, or stored
-locally in a temporarily database for sending/pushing later. This enables
+From the developer's perspective, the AA methodology is based on generating
+small, high level reports called \emph{micrologs} or \emph{AA shouts} of what
+they are doing in a specific period of time.  This timeframe can be
+between 5 to 15 minutes in our proposed practice, depending on what is most
+convenient for the developer and the team. An \emph{AA Session} is a focused
+period of continuous work, lasting about 2 hours in our proposed regime, in
+which the developer issues a collection of these short periodical
+\emph{micrologs}. The developer can set reminders or alerts to show up when it
+is time to \emph{microlog}. The objective of the flexible timeframe and alert
+scheme is to \emph{minimize developer overhead} during his AA session.  In this
+way, the developer can issue micrologs while staying maximally focused on
+his code. Each microlog may be sent directly to an on-line server, or stored
+locally in a temporarily database for sending/pushing later on.  This enables
 offline micrologging and periodical alerting.
 
-Developers optionally record a \emph{video screencast} at the end of the
-session summarizing what has been done, explaining with his words and
-showing his most important results. This is very similar to the video logging system
-in the movie Avatar, although it is proved from our July 2011 git repositories and online
-wiki that we have used this powerful concept in AA prior to the release of
-Avatar. Moreover, screencasting differs from general videologging in that it
-typically captures actual workflow on
-the computer screen. Screencasting, combined with the textual log of the AA
-Session, renders the final report more understandable for the individual developer himself
-or to other people searching for information about his production.
+Developers optionally record a brief \emph{video screencast} at the end of the
+session summarizing what has been done, explaining his goals and challenges
+in his own words and showing his most important results. This is very similar to
+the video logging system in the movie Avatar, although it is clear from our
+July 2011 git repositories and online wiki that we have used this powerful
+concept in AA prior to the release of Avatar. Moreover, screencasting differs
+from general videologging in that it typically captures actual workflow on the
+computer screen. Screencasting, combined with the textual log of the AA Session,
+renders the final report more understandable for the individual developer
+himself or to other people searching for information about his production.
+% TODO: talvez figura com diagraminha explicando uma semana tipica,
+% e cada sessao AA com um bloco dividido nos 15 min, e no inicio do bloco tb
+% podemos listar no diagrama ``developer reads felow AA logs and lists his
+% assigned tickets before he starts''
 
-\subsection{AA website report}
+\subsection{The AA website report}
 
@@
 \begin{center}
-   \includegraphics[width=\linewidth]{figs/aa-0_1.png}
+   \includegraphics[width=0.95\linewidth]{figs/aa-0_1.png}
 \end{center}
-   \caption{AA Report Agregator Version 0.1}
+   \caption{AA Report Agregator Version 0.1. Messages of users hybrid, filter0,
+   v1z e aut0mata about activities of interest for labMacambira.sf.net and
+   related entities (IPRJ/UERJ, IFT/UNESP, IFSC/USP, OPW/Mozilla,
+   PulaPirataComics). Each message is a ``shout'', which, grouped, forms an AA
+   session.}
 \label{fig:aaserver}

### 2013-06-02 paper.tex
290

@@
 \begin{center}
-   \includegraphics[width=0.95\linewidth]{figs/aa-0_1.png}
+   \includegraphics[width=0.95\linewidth]{figs/aa-0_1_.png}
 \end{center}
@@
    v1z e aut0mata about activities of interest for labMacambira.sf.net and
-   related entities (IPRJ/UERJ, IFT/UNESP, IFSC/USP, OPW/Mozilla,
+   a diversity of collaborating entities (IPRJ/UERJ, IFT/UNESP, IFSC/USP, OPW/Mozilla,
    PulaPirataComics). Each message is a ``shout'', which, grouped, forms an AA
@@
 All AA reports made by the developers are ultimatelly sent to a Web server and
-become publically aggregated on a Website, enabling a manager or another
+become publically aggregated on a Website. It is then possible for a manager or another
 developer to follow very closely the work of a given developer, nearly real-time,
-reading each of his small reports or micrologs of what he is doing right now.
+reading each of his small reports or micrologs of what he is working on.
 
-Another use case is to check older sessions to check when certain tasks
+Another possibility is to check older sessions to check when certain tasks
 were carried out and the comments of the developer about the \emph{process}.
@@
 about what was was done -- specially \emph{how} it was done -- becomes very easy to
-understand, instead of a long report in the end of a session.
+understand, as opposed to having a long report in the end of a session.
 
-IN the current version of the AA server infrastructure, the aggregating
-website is where the developer can attach a link for his
+In the current version of the AA server infrastructure, the aggregating
+website allows the developer to attach a link for his
 screencast about each session. Screencasts are specially useful on cases where
 the small reports were done in a hurry, because the developer did not want to
-lose his focus on something important at that moment.
+lose his focus on something critical at that moment.
 

### 2013-06-02 paper.tex
at results and discussion

@@
 
-\subsection{Peer validation and peer pressure}
+\subsection{Peer validation}
 
 No set bosses or leaders are required in an ideal aplication the AA methodology;
-in practice the role of bosses basically disappears or is greatly lightened --
+in practice, the role of bosses basically disappears or is greatly alleviated --
 hence the name Algorithmic \emph{Autoregulation} and other implicit
@@
 AA session must be validated by another developer. More specifically, all
-reports are read by someone that will mark them as `valid' or `invalid' and may
-optionally write commentaries about the specific session and quality of
-micrologs. The developer in charge of validating any given session is randomly
-assigned by the AA Web server, which sends an email to the chosen developer with
-an URL to a validation interface.
+reports are read by someone that will mark them collectively as `valid' or
+`invalid' and may optionally write commentaries about the specific session and
+quality of micrologs. The developer in charge of validating any given session is
+randomly assigned by the AA Web server, which sends an email to the chosen
+developer with an URL to a validation interface.
+
+Peer validation also helps in making decentralized collaboration more cohesive
+by encouraging members to be minimally aware of peer activities, even when these
+are not immediately useful for accomplishing the task at hand.  We have observed
+that decentralized teamwork can get so efficient at actual production that 
+the team gets short-sighted in terms of coordination: non-communicating
+subteams can get formed in practice if care is not taken, causing a
+fragmentation of the collective. Peer validation is one way to help avoid
+fragmentation and is an essential mechanism of decentralized team autoregulation.
 

### 2013-06-02 paper.tex
moving on to conclusions

@@
 
-Easy and effective GSD team management is the main
-purpose of the AA methodology. We applied the proposed methodology to a group
-of 9 developers in July of 2011 on what we called Lab
+Easy and effective GSD team management is the main purpose of the AA
+methodology. We applied the proposed methodology to a group of 9 developers in
+July of 2011, 3 of which are coauthors of the present work, on what we called Lab
 Macambira.\footnote{LabMacambira.sf.net: \url{http://labmacambira.sf.net}} The
 main objective of the team was to work on an array of strategic free software
-projects in the broad audiovisual and web categories,
-contributing directly to their development, submitting bug patches or
-committing new features to their source code.
+projects in the broad audiovisual and web categories, contributing directly to
+their development, submitting bug patches or committing new features to their
+source code.
 
-The team members had different levels of experience on software
-development for large and distributed free software projects
-like Scilab or Mozilla. In this way, one month of training was
-conducted by two experienced developers, teaching the basics of use of
-development support tools like bugtrackers, programming languages,
-version control systems, and build systems. After this period, a starter project
-was proposed for new developers, namely to have an accepted patch or commit to
-the official repository for a large and widely used free software app.
-Developers passing the starter project would be deemed `initiated' and be
-called a `Macambira' developer.  Table~\ref{tabela:contribuicoes} summarizes the
-effective accepted contributions of each successfully initiated developer to
-free software projects in 2011, which used the AA methodology.
+The team members had different levels of experience on software development for
+large and distributed free software projects like Scilab or Mozilla. In this
+way, one month of training was conducted by three experienced developers (the
+first authors of the present work), teaching
+the basics of use of development infrastructure tools like bugtrackers, programming
+languages, version control systems, and build systems. After this period, a
+starter project was proposed for new developers: to submit a bugfix or implement
+a new feature and have an accepted
+patch or commit to the official repository for a large free
+software project.  Developers passing the starter project would be deemed
+`initiated' and be called a `Macambira' developer, and be hired for the
+remainder of the semester.
+Table~\ref{tabela:contribuicoes} summarizes the effective accepted contributions
+of each successfully initiated developer to free software projects in 2011,
+which used the AA methodology.
 
@@
 \begin{table}
-  \caption{Free and open software projects that received contributions
-    from successfully initiated developers  or `macambiras' that worked under
-    the AA methodology. The first column lists applications to which
-    contributions were officially accepted and whose development process was
-    performed and logged using AA. The second column shows the pseudonym of the
-    ``Macambiras'' who sent \textit{commits} to the application. At
-    Lab Macambira, and at the free software community at large, 
-    It is
-    common practice at Lab Macambira, and the free software community at large,
-    to use pseudonyms as identification in order to enhance privacy in practice.}
-    \begin{tabular}{|l|l|}
+  \caption{Free and open software projects that received contributions from
+  successfully initiated developers  or `Macambiras' that worked under the AA
+  methodology. The first column lists applications to which contributions were
+  officially accepted and whose development process was tracked using AA. The
+  second column shows the pseudonym of the committers. At Lab Macambira it is
+  common practice to use pseudonyms in AA as identification in order to enhance
+  privacy.}
+  \begin{center}
+  \small\begin{tabular}{|l|l|}
         \hline
@@
     \end{tabular}
+    \end{center}
     \label{tabela:contribuicoes}
@@
 system and used AA to manage their activities, even while developing
-the other aforementioned free software projects.
-
-The source code of AA --- both the client that sends the
+the other aforementioned free software projects. Thus, AA and the associated
+software system was tested, prototyped, and developed in close contact with
+actual practice.  The source code of AA --- both the client that sends the
 logs and the AA Web server --- is public available as free software\footnote{AA source
@@
 
-After the initial training period of 1 month, the ``Macambiras''
-worked during 6 additional months on a large range of free software projects, distributed on work
-groups --- each work group focusing on a specific theme like video,
-audio and web --- and funded by contracts, freelance, and support of the
+After the initial training period of 1 month, the initiated ``Macambiras''
+worked during 6 additional months on a large range of free software projects,
+divided into work groups --- each work group focusing on a specific theme like
+video, audio and web --- and funded by contracts, freelance, and support of the
 Pont\~{a}o N\'{o}s Digitais\footnote{\url{http://nosdigitais.teia.org.br}}.
-Table~\ref{tabela:criados} has a list of \emph{new} free software applications created by
-``Lab Macambira'' since July 2011.
+Table~\ref{tabela:criados} has a list of \emph{new} free software applications
+created by ``Lab Macambira'' since July 2011.
 
@@
       of application.}
-
-    \footnotesize{
-    \begin{tabular}{|l|p{5cm}|l|}
+    \footnotesize{\begin{tabular}{|l|p{5cm}|l|}
         \hline
@@
 
-As of this writing the `Lab Macambira' unites 15 software developers, and 
-key 2011-initiated developers continue to work voluntarily in the project.
+As of this writing, `Lab Macambira' unites over 15 software developers, and 
+key developers trained in 2011 continue to work voluntarily in the project.
 

### 2013-06-02 paper.tex
integrado com hybrid

@@
 %% brief introduction
-In a scenario where Global Software Development is growing as a popular 
-software development scenario both for free software projects and for the entire
-software industry, we need methodologies to deal with its potential disadvantages and at
-the same time to amplify its advantages.
+In a scenario where Global Software Development is growing as a popular form of
+software development in the entire software industry, there is an increasing
+need for methodologies to deal with its potential disadvantages and at the same
+time to amplify its advantages.
 
-This paper has presented a methodology for GSD, \ie, the development
-involving a series of large or small groups of software developers, each either working on
+This paper has presented a new methodology for GSD, in scenarios
+involving a series of large or small groups of software developers, either working on
 different countries or at the same room. The AA methodology
 implements a simple system where each developer take notes of his work
-posting a periodic log of small text sentences or microblogs. The sum of those
-micrologs, along with an entire session of work, results in a complete
-report. The report is made public available through a Website and be
-validated by other developers randomly sorted by the AA Web server.
+posting a periodic log of small text sentences or micrologs. The sum of those
+activity logs, along with an entire session of work, results in a complete
+unit of report. The report is made public available through a Website and is
+validated by peers that are randomly selected by the AA Web server.
 
-AA is not limited to a merely work-management tool, but acts as \emph{a
-methodology to improve the time sensibility of individuals}, dividing their
-complex work on small chunks or sessions, and also reducing the need of extensive
-reports or unnecessary meetings. By asking users to write a minimal
+AA is not limited to a work-management tool, but acts as \emph{a
+methodology to improve the time sensibility of individuals}, helping divide their
+complex tasks in time into small chunks or sessions, and also reducing the need
+of extensive reports or unnecessary meetings. By asking users to write a minimal
 text sentence as a continuous log feed, the proposed methodology avoids
-disturbing the flow of developers which are heavily concentrated on programming: developers
-just have to type a few words and go back to coding.
+disturbing the flow of developers which are heavily concentrated on programming:
+developers just have to type a few words and go back to coding.
 
-The AA methodology is not restricted to software development, even though it was
-designed for the latter. As of this
-writing there is an entertainment studio, Pula Pirata\footnote{
-\url{www.pulapirata.com}}, that has started to use AA to manage their creative activities.
-Other people with no software background, like social scientists, musicians and
-activists also have been using AA and contributing for its improvement.
-
-For developer teams, we have analyzed data and gained practical experience on
+We have analyzed data and gained practical experience on
 the use of AA to auto-regulate the work of Lab Macambira, a group of free
@@
 
+The AA methodology is not restricted to software development, even though it was
+designed for the latter. As of this writing there is an entertainment studio,
+Pula Pirata\footnote{ \url{www.pulapirata.com}}, that has been using AA to
+manage their creative activities.  Other people with no software background,
+like social scientists, musicians and activists also have been using AA and
+contributing for its broader improvement.
+
 There are many aspects of the work which remain unfinished. Additional
-ubiquitous client
-interfaces for micrologging -- the `Twitter-like' messages -- from different
-interfaces beyond IRC, \eg, web social services and email, would greatly make the use of
-AA easier and more widespread, turning AA a truly ubiquitous and replicable system, presented on everyday
-communication channels. Another research direction is that the actual work logs
-generated by the Lab Macambira and Pula Pirata collectives since July of 2011
-could be vastly statistically analyzed aiming to recognize patterns in the
-behavior of individuals and their creative production process.
+ubiquitous client interfaces for micrologging from different interfaces beyond
+IRC, \eg, web social services and email, would greatly make the use of AA easier
+and more widespread, turning AA a truly ubiquitous and replicable system,
+presented on everyday communication channels. Another research direction is that
+the actual work logs generated by the Lab Macambira and Pula Pirata collectives
+since July of 2011 could be statistically analyzed aiming to recognize
+patterns in the behavior of individuals and their creative production process.
+It would also be desirable to provide more extensive experiments and
+psychological studies focusing on further backing specific claims made in this
+paper, in order to refine the methodology, its mechanisms and parameters.
 
 \section*{Acknowledgments}
-%TODO The authors would like to thank NSF and CNPq.
-
-We would like to thank AA: the present research and even this
+The authors acknowledge the financial support from Pontao Nos Digitais, and
+Ricardo Fabbri acknowledges support from FAPERJ/Brazil 111.852/2012.
+We would also like to thank AA: the present research and even this
 manuscript was written using AA. The complete log is on-line at

### 2013-06-02 paper.tex
paragrafo penalvim integrado

@@
 All AA reports made by the developers are ultimatelly sent to a Web server and
-become publically aggregated on a Website. It is then possible for a manager or another
+become publically aggregated on a Website called pAAnel. It is then possible for a manager or another
 developer to follow very closely the work of a given developer, nearly real-time,
@@
 
-After the initial training period of 1 month, the initiated ``Macambiras''
+After the initial training period of 1 month, the initiated `Macambiras'
 worked during 6 additional months on a large range of free software projects,
@@
 Table~\ref{tabela:criados} has a list of \emph{new} free software applications
-created by ``Lab Macambira'' since July 2011.
+created by `Lab Macambira' since July 2011.
+
+While using the AA system, developers learned to work asynchronously with others
+and got used to the habit of periodically updating their status on their
+projects. Each programmer was given the chance to work with considerable
+freedom, in any place and time of preference. The strictest required
+responsibility was that of using AA for at least one sesssion of 2h per day,
+while working on the agreed upon tasks.  The online pAAnel allowed each
+developer to quickly grasp activities from others while avoiding interrupting them, a
+process further aided by the screencasts. Adjustments to the task deadlines and
+milestones (which were managed in Trac) were done based on verified progress of
+individuals and labMacambira.sf.net team as a whole. The various newcommers
+benefitted from a fast learning team by the use of AA as a flexible and simple
+transparency system. Updates from the team were not transmitted on a
+person-to-person basis, but rather on a person-to-team basis based on the
+available online progress information.
 

### 2013-06-02 paper.tex
tentando author bulk package

@@
         \hline
-        Application & Description & Technologies involved \\ 
+        Application & Description & Technologies \\ 
         \hline \hline
@@
         \hline
-        animal        & An Imaging Library              & C \\
+        Animal        & An Imaging Library              & C \\
         \hline
@@
         \hline
-        ``Confer\^{e}ncia Permanente'' & Platform for the permanent
+        Confer\^{e}ncia Permanente & Platform for the permanent
         conference of the rights of minors & PHP, JavaScript \\
@@
         sound from flying origami tracked by webcams & Puredata,
-        C/C++, Scilab \\
+        C++, Scilab \\
         \hline

### 2013-06-02 front-matter.tex
merged hybrid commit a45db4aa47012
tentando author bulk package
lista de autores ficou baoa

@@
 
-\author{%
-Renato~Fabbri \and Ricardo~Fabbri \and Vilson Vieira \and Alexandre Negr\~{a}o \and Lucas Zambianchi
-\and Marcos Mendon\c{c}a \and Danilo Shiga\\[1em]
-\small{
-Instituto de F\'{i}sica de S\~{a}o Carlos (IFSC), Universidade de
-S\~{a}o Paulo (USP), Brazil}\\[0.5em]
-\small{Instituto Polit\'{e}cnico, Universidade do Estado do Rio de
-Janeiro, Nova Friburgo, RJ, Brazil}\\[0.5em]
-\url{LabMacambira.sourceforge.net}.
-}
+%\author{%
+%Renato~Fabbri \and Ricardo~Fabbri \and Vilson~Vieira \and Alexandre~Negr\~{a}o \and Lucas~Zambianchi
+%\and Marcos~Mendon\c{c}a \and Daniel~Penalva \and Danilo~Shiga\\[1em]
+%\small{
+%Instituto de F\'{i}sica de S\~{a}o Carlos (IFSC), Universidade de
+%S\~{a}o Paulo (USP), Brazil}\\[0.5em]
+%\small{Instituto Polit\'{e}cnico, Universidade do Estado do Rio de
+%Janeiro, Nova Friburgo, RJ, Brazil}\\[0.5em]
+%\url{LabMacambira.sourceforge.net}.
+%}
+
+\author[1,3]{Renato Fabbri\thanks{renato.fabbri@gmail.com}}
+\author[2,3]{Ricardo Fabbri\thanks{rfabbri@iprj.uerj.br}}
+\author[1,3]{Vilson Vieira\thanks{vilson@void.cc}}
+\author[3]{Alexandre Negr\~{a}o}
+\author[3]{Lucas Zambianchi}
+\author[3]{Marcos~Mendon\c{c}a}
+\author[1,3]{Daniel Penalva}
+\author[3]{Danilo Shiga}
+\affil[1]{\small Physics Insitute at S\~{a}o Carlos (IFSC), University of S\~{a}o Paulo (USP), Brazil}
+\affil[2]{\small Polytechnic Institute (IPRJ), State University of Rio de Janeiro (UERJ), Nova Friburgo, Brazil}
+\affil[3]{\small \url{LabMacambira.sourceforge.net}}
+
+\renewcommand\Authands{ and }
 

### 2013-06-02 paper.tex
lista de autores ficou baoa
colocado diagrama sem corte brusco aa direita
retiradas incidencias de primeiras pessoas sing e pl

@@
 \usepackage{multirow}
+\usepackage{authblk}
 \usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}
@@
 
-Despite the advantages of GSD, we have noticed how
+Despite the advantages of GSD, it is known how
 difficult it is to coordinate and fund free software on a significantly larger scale
@@
 or to do other bureaucratic activities such as writing lengthy reports to
-justify their funding.\cite{Thompson:Wired:2012}
+justify their funding~\cite{Thompson:Wired:2012}.
 %\todo{ler mais. cacm, etc}
 
-In this paper we propose the AA methodology and an associated software system
+To address these matters is the purpose of
+AA methodology and an associated software system
 for coordinating distributed team work, tackling the disadvantages
@@
 easy UI alternatives: unix shell
-commands, native GUI or web page, conventional social network posts, or chat messages to a log bot
+commands, native GUI or Web page, conventional social network posts, or chat messages to a log bot
 listening to IRC, GTalk, G+, and others.  These ``microblog sentences'' are
@@
 in Section~\ref{related-work}. In Section~\ref{aa-methodology} the most relevant
-characteristics of the AA methodology are outlined. In the same section, we
-report an actual use case of AA for coordinating a team of 9 paid developers 
+characteristics of the AA methodology are outlined. In the same section, it
+is reported an actual use case of AA for coordinating a team of 9 paid developers 
 % nao era mais de 9?
 during the second half of 2011, as well as a broader use case of AA from 2012 to
-the present time.  In Section~\ref{conclusions} we summarize the overall conclusions
-and indicate future possibilities for the practical use of AA in other types of
+the present time.  In Section~\ref{conclusions}, there are overall conclusions
+and indicative of future possibilities for the practical use of AA in other types of
 teams of software developers or organizations working on non-software distributed
@@
 complex methodologies and many were built for a specific company or software
-center. We propose an alternative methodology based on a simple idea: small
+center. This section describes an alternative methodology based on a simple idea: small
 working sessions logged by a computational tool. Figure~\ref{fig:mm} summarizes
@@
 \begin{center}
-   \includegraphics[width=0.8\linewidth,keepaspectratio=true]{figs/aa-mm.png}
+   \includegraphics[width=0.8\linewidth,keepaspectratio=true]{figs/aa-mm2.png}
 \end{center}
@@
 in his own words and showing his most important results. This is very similar to
-the video logging system in the movie Avatar, although it is clear from our
-July 2011 git repositories and online wiki that we have used this powerful
-concept in AA prior to the release of Avatar. Moreover, screencasting differs
+the video logging system in the movie Avatar, although it is clear, from
+July 2011 git repositories and online wiki, that we have used this powerful concept in AA independently of any major mass midia incident. Moreover, screencasting differs
 from general videologging in that it typically captures actual workflow on the
@@
 renders the final report more understandable for the individual developer
-himself or to other people searching for information about his production.
+himself and to other people searching for information about his production.
 % TODO: talvez figura com diagraminha explicando uma semana tipica,
@@
 by encouraging members to be minimally aware of peer activities, even when these
-are not immediately useful for accomplishing the task at hand.  We have observed
+are not immediately useful for accomplishing the task at hand.  Authors have observed
 that decentralized teamwork can get so efficient at actual production that 
@@
 Easy and effective GSD team management is the main purpose of the AA
-methodology. We applied the proposed methodology to a group of 9 developers in
-July of 2011, 3 of which are coauthors of the present work, on what we called Lab
+methodology. The proposed methodology was applyed to a group of 9 developers in
+July of 2011, 3 of which are coauthors of the present work, on what was named Lab
 Macambira.\footnote{LabMacambira.sf.net: \url{http://labmacambira.sf.net}} The
@@
 
-We have analyzed data and gained practical experience on
+Authors have analyzed data and gained practical experience on
 the use of AA to auto-regulate the work of Lab Macambira, a group of free
@@
 \section*{Acknowledgments}
-The authors acknowledge the financial support from Pontao Nos Digitais, and
+Authors acknowledge the financial support from Pontao Nos Digitais, and
 Ricardo Fabbri acknowledges support from FAPERJ/Brazil 111.852/2012.
-We would also like to thank AA: the present research and even this
+Authors also thank AA: the present research and even this
 manuscript was written using AA. The complete log is on-line at
 \url{www.pulapirata.com/skills/aa}.
+Finally, authors are also grateful to IFSC/USP, IPRJ/UERJ, IFT/UNESP and all AA users and collaborators, especially those who coded AA hacks for use through shell and bot, and those who coded different Web interfaces in use.
 

### 2013-06-02 front-matter.tex
mudado titulo
arrumados vinculos da autoria

@@
 \title{
-AA Distributed Software Development Methodology
+    AA: The Algorithmic Autoregulation (Distributed Software Development) Methodology
 }
@@
 
-\author[1,3]{Renato Fabbri\thanks{renato.fabbri@gmail.com}}
-\author[2,3]{Ricardo Fabbri\thanks{rfabbri@iprj.uerj.br}}
-\author[1,3]{Vilson Vieira\thanks{vilson@void.cc}}
-\author[3]{Alexandre Negr\~{a}o}
-\author[3]{Lucas Zambianchi}
-\author[3]{Marcos~Mendon\c{c}a}
+\author[1,4]{Renato Fabbri\thanks{renato.fabbri@gmail.com}}
+\author[2,4]{Ricardo Fabbri\thanks{rfabbri@iprj.uerj.br}}
+\author[1,4]{Vilson Vieira\thanks{vilson@void.cc}}
+\author[4]{Alexandre Negr\~{a}o}
+\author[4]{Lucas Zambianchi}
+\author[4]{Marcos~Mendon\c{c}a}
 \author[1,3]{Daniel Penalva}
-\author[3]{Danilo Shiga}
-\affil[1]{\small Physics Insitute at S\~{a}o Carlos (IFSC), University of S\~{a}o Paulo (USP), Brazil}
+\author[4]{Danilo Shiga}
+\affil[1]{\small Physics Insitute at S\~{a}o Carlos (IFSC), University of S\~{a}o Paulo (USP), S\~ao Carlos, Brazil}
 \affil[2]{\small Polytechnic Institute (IPRJ), State University of Rio de Janeiro (UERJ), Nova Friburgo, Brazil}
-\affil[3]{\small \url{LabMacambira.sourceforge.net}}
+\affil[3]{\small  Institute for Theoretical Physics (IFT), Universidade Estadual Paulista (UNESP), S\~ao Paulo, Brazil}
+\affil[4]{\small \url{LabMacambira.sourceforge.net}}
 

### 2013-06-03 front-matter.tex
retirados hifens de auto-regulationS, mudado meu email, arrumado unfold de IFSC, colocados pts finais nos rodapes

@@
 
-\author[1,4]{Renato Fabbri\thanks{renato.fabbri@gmail.com}}
+\author[1,4]{Renato Fabbri\thanks{fabbri@usp.br}}
 \author[2,4]{Ricardo Fabbri\thanks{rfabbri@iprj.uerj.br}}
@@
 \author[4]{Danilo Shiga}
-\affil[1]{\small Physics Insitute at S\~{a}o Carlos (IFSC), University of S\~{a}o Paulo (USP), S\~ao Carlos, Brazil}
+\affil[1]{\small S\~ao Carlos Institute of Physics (IFSC), University of S\~{a}o Paulo (USP), S\~ao Carlos, Brazil}
 \affil[2]{\small Polytechnic Institute (IPRJ), State University of Rio de Janeiro (UERJ), Nova Friburgo, Brazil}
@@
 We present a new self-regulating methodology for coordinating distributed team
-work called Algorithmic Auto-regulation (AA), based on recent social networking
+work called Algorithmic Autoregulation (AA), based on recent social networking
 concepts and individual merit. Team members take on an egalitarian role, and

### 2013-06-03 paper.tex
todo discussion about epistemological shift of academy with the AA methodology productive to processual
retirados hifens de auto-regulationS, mudado meu email, arrumado unfold de IFSC, colocados pts finais nos rodapes

@@
 and constantly updated list of the most active projects on-line at
-\url{www.ohloh.net}} Beyond the free and open software community, GSD has a growing
+\url{www.ohloh.net}.} Beyond the free and open software community, GSD has a growing
 popularity in every niche of the software industry as a whole, even among those
@@
 July of 2011, 3 of which are coauthors of the present work, on what was named Lab
-Macambira.\footnote{LabMacambira.sf.net: \url{http://labmacambira.sf.net}} The
+Macambira.\footnote{LabMacambira.sf.net: \url{http://labmacambira.sf.net}.} The
 main objective of the team was to work on an array of strategic free software
@@
   code:
-  \url{http://labmacambira.git.sourceforge.net/git/gitweb.cgi?p=labmacambira/aa}}
+  \url{http://labmacambira.git.sourceforge.net/git/gitweb.cgi?p=labmacambira/aa}.}
 and all the actual AA log data of the entire team of Lab Macambira from 2011 to
 the present time is also on-line\footnote{Logs of AA sessions:
-  \url{http://labmacambira.git.sourceforge.net/git/gitweb.cgi?p=labmacambira/paainel}}.
+  \url{http://labmacambira.git.sourceforge.net/git/gitweb.cgi?p=labmacambira/paainel}.}.
 
@@
 video, audio and web --- and funded by contracts, freelance, and support of the
-Pont\~{a}o N\'{o}s Digitais\footnote{\url{http://nosdigitais.teia.org.br}}.
+Pont\~{a}o N\'{o}s Digitais\footnote{\url{http://nosdigitais.teia.org.br}.}.
 Table~\ref{tabela:criados} has a list of \emph{new} free software applications
@@
         \hline \hline
-        AA            & Algorithmic Auto-regulation      & Python, PHP \\
+        AA            & Algorithmic Autoregulation      & Python, PHP \\
         \hline
@@
 Authors have analyzed data and gained practical experience on
-the use of AA to auto-regulate the work of Lab Macambira, a group of free
+the use of AA to autoregulate the work of Lab Macambira, a group of free
 software developers from Brazil. Since July of 2011 the group has contributed
@@
 designed for the latter. As of this writing there is an entertainment studio,
-Pula Pirata\footnote{ \url{www.pulapirata.com}}, that has been using AA to
+Pula Pirata\footnote{ \url{www.pulapirata.com}.}, that has been using AA to
 manage their creative activities.  Other people with no software background,
@@
 paper, in order to refine the methodology, its mechanisms and parameters.
+
+%\todo{ Discussao : AA como mudanca de epistemologia academica, vinculacao do valor processual atraves do AA, citacao nos shouts (durante trabalho academico), validacao de citacoes atraves de validacao do log de AA + revisao do processo academico presento no log em algum nivel de compreensao midiatica}
 

### 2013-06-05 paper.tex
final version for wsl2013

@@
 status of their activity. Logging is carried out using a series of
-easy UI alternatives: unix shell
+easy UI alternatives: UNIX shell
 commands, native GUI or Web page, conventional social network posts, or chat messages to a log bot
@@
 specific context and demands.  Moreover, `distributed development' is generally
-tought of as being global which is not true. For instance, AA has been
+thought of as being global which is not true. For instance, AA has been
 effectively applied to a team whose members live in the same city but work at
@@
 A thorough survey of these methodologies is beyond the scope of this paper; this
-section presents but a brief overview.
+section presents a brief overview.
 
@@
 hampers synchronous communications like telephone or
-videoconferences. Units of developers working on different time-zones
+video conferences. Units of developers working on different time-zones
 are concerned with managing of their agenda guided by this temporal
@@
 the video logging system in the movie Avatar, although it is clear, from
-July 2011 git repositories and online wiki, that we have used this powerful concept in AA independently of any major mass midia incident. Moreover, screencasting differs
+July 2011 git repositories and online wiki, that we have used this powerful concept in AA independently of any major mass media incident. Moreover, screencasting differs
 from general videologging in that it typically captures actual workflow on the
@@
 
-All AA reports made by the developers are ultimatelly sent to a Web server and
-become publically aggregated on a Website called pAAnel. It is then possible for a manager or another
+All AA reports made by the developers are ultimately sent to a Web server and
+become publicly aggregated on a Website called pAAnel. It is then possible for a manager or another
 developer to follow very closely the work of a given developer, nearly real-time,
@@
 Since each AA microlog happens in a very short timeframe of work, the information
-about what was was done -- specially \emph{how} it was done -- becomes very easy to
+about what was done -- specially \emph{how} it was done -- becomes very easy to
 understand, as opposed to having a long report in the end of a session.
@@
 
-No set bosses or leaders are required in an ideal aplication the AA methodology;
+No set bosses or leaders are required in an ideal application the AA methodology;
 in practice, the role of bosses basically disappears or is greatly alleviated --
@@
 Easy and effective GSD team management is the main purpose of the AA
-methodology. The proposed methodology was applyed to a group of 9 developers in
+methodology. The proposed methodology was applied to a group of 9 developers in
 July of 2011, 3 of which are coauthors of the present work, on what was named Lab
@@
   officially accepted and whose development process was tracked using AA. The
-  second column shows the pseudonym of the committers. At Lab Macambira it is
+  second column shows the pseudonym of the ``committers''. At Lab Macambira it is
   common practice to use pseudonyms in AA as identification in order to enhance
@@
         \hline
-        Application           & ``Commiters''                       \\
+        Application           & ``Committers''                       \\
         \hline \hline
@@
 freedom, in any place and time of preference. The strictest required
-responsibility was that of using AA for at least one sesssion of 2h per day,
+responsibility was that of using AA for at least one session of 2h per day,
 while working on the agreed upon tasks.  The online pAAnel allowed each
@@
 milestones (which were managed in Trac) were done based on verified progress of
-individuals and labMacambira.sf.net team as a whole. The various newcommers
-benefitted from a fast learning team by the use of AA as a flexible and simple
+individuals and labMacambira.sf.net team as a whole. The various newcomers
+benefited from a fast learning team by the use of AA as a flexible and simple
 transparency system. Updates from the team were not transmitted on a
@@
         \hline
-        pAAinel & Dashboard for the real-time visualization of Lab
+        pAAnel & Dashboard for the real-time visualization of Lab
         Macambira activity & Python, Django \\
@@
 \section*{Acknowledgments}
-Authors acknowledge the financial support from Pontao Nos Digitais, and
+Authors acknowledge the financial support from Pont\~{a}o N\'{o}s Digitais, and
 Ricardo Fabbri acknowledges support from FAPERJ/Brazil 111.852/2012.

### 2013-08-07 paper.tex
small changes in the point of the reviewer0

@@
 OpenBSD, MySQL or Apache Software Foundation, to cite just a 
-few of the most active projects.\footnote{Ohloh, the open source network, has a more complete
-and constantly updated list of the most active projects on-line at
-\url{www.ohloh.net}.} Beyond the free and open software community, GSD has a growing
+few of the most active projects.Beyond the free and open software community, GSD has a growing
 popularity in every niche of the software industry as a whole, even among those
@@
 globalization of software companies and foundations or even the premise of
-cheaper cost of production~\cite{komi2005}.
+cheaper cost of production~\cite{komi2005}.\footnote{The open source network Ohloh has a more complete
+and constantly updated list of the most active projects on-line at
+\url{www.ohloh.net}.}
 
@@
 
-A brief overview of current work on GSD methodologies related to AA is presented
+A brief overview of current work in GSD methodologies related to AA is presented
 in Section~\ref{related-work}. In Section~\ref{aa-methodology} the most relevant
@@
 
-Second, cultural distance encompasses differences on organizational and
+Second, cultural distance encompasses differences in organizational and
 natural culture. Spoken language, unit and ethnic values are common
@@
 approach: when concluding the work day, they pass their work to the
-team working on another time-zone. Other tactics are suggested by the
+team working in another time-zone. Other tactics are suggested by the
 same author to deal with the three distances, such as separating foreign
@@
 methodologies included constant communication with critical units, incremental
-integration and schedules based on time-zones distributed to developers on 6
+integration and schedules based in time-zones distributed to developers on 6
 countries from 3 continents.
@@
 conferences like GUADEC is common on free and open source projects for
-face-to-face meetings and is based on a different place each time.
+face-to-face meetings and is based in a different place each time.
 
@@
 
-Some of the strategies for GSD mentioned on the previous section are based on
+Some of the strategies for GSD mentioned in the previous section are based on
 complex methodologies and many were built for a specific company or software
@@
 scheme is to \emph{minimize developer overhead} during his AA session.  In this
-way, the developer can issue micrologs while staying maximally focused on
+way, the developer can issue micrologs while staying maximally focused in
 his code. Each microlog may be sent directly to an on-line server, or stored
@@
 July of 2011, 3 of which are coauthors of the present work, on what was named Lab
-Macambira.\footnote{LabMacambira.sf.net: \url{http://labmacambira.sf.net}.} The
-main objective of the team was to work on an array of strategic free software
+Macambira.The main objective of the team was to work on an array of strategic free software
 projects in the broad audiovisual and web categories, contributing directly to
 their development, submitting bug patches or committing new features to their
-source code.
+source code.\footnote{LabMacambira.sf.net: \url{http://labmacambira.sf.net}.}
 
@@
 actual practice.  The source code of AA --- both the client that sends the
-logs and the AA Web server --- is public available as free software\footnote{AA source
+logs and the AA Web server --- is public available as free software and all the actual AA log data of the entire team of Lab Macambira from 2011 to
+the present time is also on-line.\footnote{AA source
   code:
-  \url{http://labmacambira.git.sourceforge.net/git/gitweb.cgi?p=labmacambira/aa}.}
-and all the actual AA log data of the entire team of Lab Macambira from 2011 to
-the present time is also on-line\footnote{Logs of AA sessions:
-  \url{http://labmacambira.git.sourceforge.net/git/gitweb.cgi?p=labmacambira/paainel}.}.
+  \url{http://labmacambira.git.sourceforge.net/git/gitweb.cgi?p=labmacambira/aa}.} \footnote{Logs of AA sessions:
+  \url{http://labmacambira.git.sourceforge.net/git/gitweb.cgi?p=labmacambira/paainel}.}
 
@@
 video, audio and web --- and funded by contracts, freelance, and support of the
-Pont\~{a}o N\'{o}s Digitais\footnote{\url{http://nosdigitais.teia.org.br}.}.
+Pont\~{a}o N\'{o}s Digitais.
 Table~\ref{tabela:criados} has a list of \emph{new} free software applications
-created by `Lab Macambira' since July 2011.
+created by `Lab Macambira' since July 2011.\footnote{\url{http://nosdigitais.teia.org.br}.}
 
@@
 responsibility was that of using AA for at least one session of 2h per day,
-while working on the agreed upon tasks.  The online pAAnel allowed each
+while working on the agreed tasks.  The online pAAnel allowed each
 developer to quickly grasp activities from others while avoiding interrupting them, a
@@
 benefited from a fast learning team by the use of AA as a flexible and simple
-transparency system. Updates from the team were not transmitted on a
-person-to-person basis, but rather on a person-to-team basis based on the
+transparency system. Updates from the team were not transmitted in a
+person-to-person basis, but rather in a person-to-team basis based on the
 available online progress information.
@@
 text sentence as a continuous log feed, the proposed methodology avoids
-disturbing the flow of developers which are heavily concentrated on programming:
+disturbing the flow of developers which are heavily concentrated in programming:
 developers just have to type a few words and go back to coding.
@@
 designed for the latter. As of this writing there is an entertainment studio,
-Pula Pirata\footnote{ \url{www.pulapirata.com}.}, that has been using AA to
-manage their creative activities.  Other people with no software background,
+Pula Pirata, that has been using AA to manage their creative activities.  Other people with no software background,
 like social scientists, musicians and activists also have been using AA and
-contributing for its broader improvement.
+contributing for its broader improvement.\footnote{ \url{www.pulapirata.com}.}
 

### 2013-08-07 paper.tex
revisao feita, listada em https://etherpad.mozilla.org/paaper. parece ok para enviar a revista

@@
 remainder of the semester.
-Table~\ref{tabela:contribuicoes} summarizes the effective accepted contributions
+To illustrate the coverage of overall contributions by `Macambira' developers, Table~\ref{tabela:contribuicoes} summarizes the effective accepted commits
 of each successfully initiated developer to free software projects in 2011,
@@
 As of this writing, `Lab Macambira' unites over 15 software developers, and 
-key developers trained in 2011 continue to work voluntarily in the project.
+key developers trained in 2011 continue to work voluntarily in the project and
+as voluntaries and contracted developers on fundations like Mozilla.
 
